### 2014-11-30 paper/paper.tex
upload paper

\documentclass[a4paper, twocolumn, 11pt]{article}

\usepackage{graphicx, covington, times, booktabs, multirow, color, colortbl}
\definecolor{Gray}{gray}{0.9}
\usepackage[utf8]{inputenc}
\author{Folgert Karsdorp\\\texttt{fbkarsdorp@fastmail.com}}
\title{Animacy Detection in Folktales}
\date{}

\begin{document}
\maketitle

\section{Introduction}

\section{Description of the Data}

\section{Experimental Setup}

\section{Results}
\begin{table*}
\centering
\begin{tabular}{llrrrr}
\toprule
features                                                 &           & precision & recall & $F1$ \\ \midrule
\multirow{2}{*}{word}                                    & animate   & 0.93      & 0.80   & 0.86 \\
                                                         & inanimate & 0.96      & 0.99   & 0.98 \\
\multirow{2}{*}{word + PoS}                              & animate   & 0.92      & 0.88   & 0.90 \\
                                                         & inanimate & 0.98      & 0.99   & 0.98 \\
\multirow{2}{*}{word + PoS + lemma}                      & animate   & 0.92      & 0.88   & 0.90 \\
                                                         & inanimate & 0.98      & 0.99   & 0.98 \\
\multirow{2}{*}{word + PoS + lemma + dep}                & animate   & 0.93      & 0.88   & 0.90 \\
                                                         & inanimate & 0.98      & 0.99   & 0.98 \\
\multirow{2}{*}{word + embeddings}                       & animate   & 0.93      & 0.89   & 0.91 \\
                                                         & inanimate & 0.98      & 0.99   & 0.98 \\
\rowcolor{Gray}                                          & animate   & 0.93      & 0.93   & 0.93 \\
\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings} & inanimate & 0.99      & 0.99   & 0.99 \\
\multirow{2}{*}{word + PoS + lemma + embeddings}         & animate   & 0.93      & 0.93   & 0.93 \\
                                                         & inanimate & 0.99      & 0.99   & 0.99 \\
\multirow{2}{*}{word + PoS + lemma + dep + embeddings}   & animate   & 0.92      & 0.93   & 0.93 \\
                                                         & inanimate & 0.99      & 0.99   & 0.99 \\
\multirow{2}{*}{embeddings}                              & animate   & 0.90      & 0.76   & 0.83 \\
                                                         & inanimate & 0.96      & 0.98   & 0.97 \\

\bottomrule
\end{tabular}
\caption{Precision, recall and $F1$ scores for animate and inanimate classes per feature settings for all words.}
\label{tab:results-all}
\end{table*}

\begin{table*}
\centering
\begin{tabular}{llrrrr}
\toprule
features                                               &           & precision & recall & $F1$ \\ \midrule
\multirow{2}{*}{word}                                  & animate   & 0.64      & 0.96   & 0.77 \\
                                                       & inanimate & 0.98      & 0.81   & 0.89 \\
\multirow{2}{*}{word + PoS}                            & animate   & 0.80      & 0.92   & 0.86 \\
                                                       & inanimate & 0.95      & 0.89   & 0.92 \\
\multirow{2}{*}{word + PoS + lemma}                    & animate   & 0.80      & 0.92   & 0.86 \\
                                                       & inanimate & 0.95      & 0.89   & 0.92 \\
\multirow{2}{*}{word + PoS + lemma + dep}              & animate   & 0.80      & 0.93   & 0.86 \\
                                                       & inanimate & 0.96      & 0.88   & 0.92 \\
\multirow{2}{*}{word + embeddings}                     & animate   & 0.83      & 0.93   & 0.88 \\
                                                       & inanimate & 0.96      & 0.90   & 0.93 \\
\rowcolor{Gray}                                        & animate   & 0.92      & 0.92   & 0.92 \\
\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings} & inanimate & 0.95    & 0.95   & 0.95 \\
\multirow{2}{*}{word + PoS + lemma + embeddings}       & animate   & 0.92      & 0.92   & 0.92 \\
                                                       & inanimate & 0.95      & 0.95   & 0.95 \\
\multirow{2}{*}{word + PoS + lemma + dep + embeddings} & animate   & 0.91      & 0.92   & 0.92 \\
                                                       & inanimate & 0.95      & 0.95   & 0.95 \\
\multirow{2}{*}{embeddings}                            & animate   & 0.66      & 0.88   & 0.75 \\
                                                       & inanimate & 0.94      & 0.81   & 0.87 \\

\bottomrule
\end{tabular}
\caption{Precision, recall and $F1$ scores for animate and inanimate classes per feature settings for all words tagged
as noun.}
\label{tab:results-noun}
\end{table*}
\section{Concluding Remarks}

\section*{Acknowledgments}

\bibliographystyle{plainnat}
\bibliography{paper}
\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2014-12-01 paper/paper.tex
updates

@@
-\documentclass[a4paper, twocolumn, 11pt]{article}
+\documentclass[a4paper,UKenglish]{oasics}
 
-\usepackage{graphicx, covington, times, booktabs, multirow, color, colortbl}
+\usepackage{microtype, graphicx, booktabs, multirow, color, colortbl}
+
+\usepackage{blindtext}
+\bibliographystyle{plain}
+
 \definecolor{Gray}{gray}{0.9}
 \usepackage[utf8]{inputenc}
-\author{Folgert Karsdorp\\\texttt{fbkarsdorp@fastmail.com}}
+
 \title{Animacy Detection in Folktales}
-\date{}
+\author[1]{Folgert Karsdorp}
+\author[2]{Antal van den Bosch}
+\affil[1]{Meertens Institute\\
+  Amsterdam, The Netherlands\\
+  \texttt{folgert.karsdorp@meertens.knaw.nl}}
+\affil[2]{Radboud University\\
+  Nijmegen, The Netherlands\\
+  \texttt{a.vandenbosch@let.ru.nl}}
+\authorrunning{F. Karsdorp and A. van den Bosch}
+\Copyright{Folgert Karsdorp and Antal van den Bosch}
+\subjclass{machine learning}
+\keywords{machine learning}
+
+%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\serieslogo{}%please provide filename (without suffix)
+\volumeinfo%(easychair interface)
+  {Billy Editor and Bill Editors}% editors
+  {2}% number of editors: 1, 2, ....
+  {Conference/workshop/symposium title on which this volume is based on}% event
+  {1}% volume
+  {1}% issue
+  {1}% starting page number
+\EventShortName{}
+\DOI{10.4230/OASIcs.xxx.yyy.p}% to be completed by the volume editor
+% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+\begin{abstract}
+\blindtext[1]
+\end{abstract}
+
 \section{Introduction}
 
-\section{Description of the Data}
+\section{Previous Work}
+
+\subsection{Definitions of Animacy}
+In most of the literature on automatic animacy detection, people make
+a binary distinction between animate and inanimate. \cite{bowman:12}
+for example, define objects to be animate if they are alive and have
+the ability to move under their own will. \cite{orasan:07} define animacy
+in the context of anaphora resolution: ``if its referent can also be
+referred to using one of the pronouns he, she, him, her, his, hers,
+himself, herself, or a combination of such pro- nouns (e.g. his/her
+)'' Not only are some parts of these definitions somewhat circular, we
+believe that they are both essentialistic and not particularly suited
+for the rich and wondrous entities that live in the realm of
+folktales. A more sophisticated way to handle the distinction between
+animate and inanimate objects is given by Dennett CITE who argues that
+animacy and aliveness should be treated as epistemological stances
+rather than fixed states in the world.
+
+\subsection{Methods for Automatic Animacy Detection}
+
+\section{The Data, Annotation and Preprocessing}
+
+\subsection{The Data}
+Our corpus consists of 74 Dutch folktales from the collection
+\textit{Volkssprookjes uit Nederland en Vlaanderen} compiled by
+\cite{sinninghe:78}. The collection contains retellings of popular
+folktales from the Netherlands and Flanders, such as \textit{The
+  Bremen Town Musicians} (ATU 130) and \textit{The Table, the Ass, and
+  the Stick } (ATU 563). Some of the less well-know stories are
+\textit{Cock, Hen, Duck, Pin, and Needle on a Journey} (AT 0210) and
+(ATU 780) \textit{The Singing Bone}. A digital version of the
+collection available in the Dutch Folktale Database from the Meertens
+Institute (corpus SINVSUNV.20E).\footnote{See
+  \url{http://www.verhalenbank.nl}} We chose this collection because
+of its homogeneity while it still displays enough diversity: (1) all
+tales are written in standard Dutch, (2) the stories are edited by the
+same editor, (3) the tales have a comparable length and (4) the
+collection contains tales from a number of different genres, such as
+fairy tales, legends, riddles etc.
+
+\subsection{Annotation}
+We labeled the corpus for animacy using the definition provided by
+%\cite{bowman:12}:
+
+\subsection{Preprocessing}
 
 \section{Experimental Setup}
+This section describes our experimental setup including the features
+used, the machine learning models we applied and our methods of
+evaluation.
+
+\subsection{Task description}
+
+\subsection{Features}
+
+We explore a range of different features and feature combinations
+including lexical features, morphological features, syntactic features
+and semantic features.
+
+\begin{description}
+\item[Lexical features]
+We take a sliding window approach where for each focus word (i.e.\ the
+word for which we want to predict whether it is animate or not) we
+extract $n$ words to the left, $n$ words to the right and the focus
+word itself. In all experiments we set $n$ to 3. In addition to the
+word forms, we extract for each word in a window the lemmata
+corresponding to the words.
+
+\item[Morphological Features]
+For each word we extract its Part-of-Speech tag. For reasons of
+comparability for choose to use the tags as provided by the
+output of the syntactic parser, instead of a more specialized
+Part-of-Speech tagger. Again, we take a sliding window approach and
+extract the Part-of-Speech tags for $n$ word left of the focus word
+and $n$ word right of the focus word as well as the tag of the focus
+word itself.
+
+\item[Syntactic Features]
+We extract the dependency tag for each word and its $n$ neighbors to
+the right and to the left as provided by the syntactic parser
+Alpino. Animate entities tend to take subject or object positions in a
+sentences which is why this feature is expected and has proven to
+perform rather well.
+
+\item[Semantic Features]
+For each focus word we add a 300 dimension vector representation that
+was learned by applying word2vec on the Dutch version of CoW. Missing
+words were given a vector were all dimensions are set to zero.
+
+\end{description}
+
+\subsection{Models}
+We make use of a Logistic Regression classifier also know as a Maximum
+Entropy classifier with L2 regularization as implemented in
+\cite{sklearn}. In all experiments, we set the regularization strength
+parameter $C$ to 1.
+
+We compare nine models in which we make use of different feature
+combinations: (1) words, (2) words and Part-of-Speech tags, (3)
+words, Part-of-Speech tags and lemmata, (4) words, Part-of-Speech
+tags, lemmata and dependency tags, (5) word embeddings and (6-9) the
+features in model 1 to 4 plus word embeddings.
+
+\subsection{Evaluation}
+
+Animate words are far outnumbered by the inanimate words in our
+collection. Reporting accuracy scores would therefore provide
+distorted results to much in favor of the majority category. The
+rarity of animate words make evaluation measures such as the
+well-known F1-score more appropriate. For all experiments we report on
+the precision, recall and F1-score \cite{rijsbergen:79}. In most of
+the literature on animacy detection results are only given for the
+classification of nouns. We will report the results for all words in a
+text and on those specifically for nouns separately.
 
 \section{Results}
+
 \begin{table*}
@@
 \end{tabular}
-\caption{Precision, recall and $F1$ scores for animate and inanimate classes per feature settings for all words tagged
-as noun.}
+\caption{Precision, recall and $F1$ scores for animate and inanimate classes
+  per feature settings for all words tagged as noun.}
 \label{tab:results-noun}
 \end{table*}
+
 \section{Concluding Remarks}
@@
 
-\bibliographystyle{plainnat}
-\bibliography{paper}
+The work on which this paper is based has been supported by the
+Computational Humanities Programme of the Royal Netherlands Academy of
+Arts and Sciences, under the auspices of the Tunes \& Tales
+project. For further information, see \url{http://ehumanities.nl}.
+
+Constantin Oraˇsan and Richard J. Evans. 2007. NP an- imacy identification for anaphora resolution. Journal of Artificial Intelligence Research (JAIR), 29:79–103.
+\begin{thebibliography}{50}
+\bibitem{sinninghe:78} Sinninghe J. \textit{Volkssprookjes uit
+    Nederland en Vlaanderen}. Kruseman, Den Haag, 1978.
+\bibitem{sklearn} Pedregosa, F. and Varoquaux, G. and Gramfort, A. and Michel, V.
+         and Thirion, B. and Grisel, O. and Blondel, M. and Prettenhofer, P.
+         and Weiss, R. and Dubourg, V. and Vanderplas, J. and Passos, A. and
+         Cournapeau, D. and Brucher, M. and Perrot, M. and Duchesnay,
+         E. `Scikit-Learn: Machine Learning in {P}ython'. In:
+         \textit{Journal of Machine Learning Research}, 12,
+         2825--2830, 2011.
+\bibitem{rijsbergen:79} {Van Rijsbergen}, C. \textit{Information
+    Retrieval}. Butterworths.
+\bibitem{bowman:12} Bowman S. and Chopra, H. `Automatic animacy
+  classification'. In \textit{Proceedings of the NAACL - HLT 2012
+    Student Research Workshop}, 2012.
+\bibitem{orasan:07} Or\u{a}san, C. and Evans, R. `NP animacy
+  identification for anaphora resolution'. In \textit{Journal of
+    Artificial Intelligence Research}, 29, 79--103, 2007.
+\end{thebibliography}
+
 \end{document}

### 2014-12-02 paper/paper.tex
updates

@@
 
-\title{Animacy Detection in Folktales}
+\title{Animacy Detection in Stories}
 \author[1]{Folgert Karsdorp}
-\author[2]{Antal van den Bosch}
+\author[2]{Marten van der Meulen}
+\author[3]{Antal van den Bosch}
 \affil[1]{Meertens Institute\\
@@
   \texttt{folgert.karsdorp@meertens.knaw.nl}}
+\affil[1]{Meertens Institute\\
+  Amsterdam, The Netherlands\\
+  \texttt{marten.van.der.meulen@meertens.knaw.nl}}
 \affil[2]{Radboud University\\
@@
   \texttt{a.vandenbosch@let.ru.nl}}
-\authorrunning{F. Karsdorp and A. van den Bosch}
-\Copyright{Folgert Karsdorp and Antal van den Bosch}
+\authorrunning{F. Karsdorp and M. van der Meulen and A. van den Bosch}
+\Copyright{Folgert Karsdorp and Marten van der Meulen and Antal van den Bosch}
 \subjclass{machine learning}
@@
 
-\section{Previous Work}
+For almost all species in the world, the capacity to distinguish
+animate objects from inanimate objects is essential to their
+survival. Those objects could be prey, for example, or predators or
+mates. The fundamental nature of the distinction between animate and
+inanimate appears in the early acquisition of the division. Children
+of less than 6 months old are well able to make the division. Recent
+brain research shows that the distinction appears in the organization
+of the ventral temporal cortex in both humans and other primates. For
+some researchers this provides evidence for the idea that the division
+between animate and inanimate is an innate part of how we see the
+world.
 
-\subsection{Definitions of Animacy}
+A common view supported by philosophers such as Daniel Dennett is that
+animacy and aliveness are better to be treated as epistemological
+stances rather than fixed states in the world. Not ineffable qualia
+but behavioral capacity defines our stance towards objects. Depending
+on whether people \textit{think} that an object is animate, they
+utilize different cognitive strategies to explain and predict the
+actions of those objects. The following opening of a well-known story
+will make this point more clear:
+
+\begin{quotation} {\it A farmer bought a pancake on the market. Once
+    he got home, the farmer was hungry and began to bake the
+    pancake. The farmer tried one of his skillful flipping techniques,
+    but he failed and the pancake fell on the ground. Coincidentally,
+    the door of the kitchen was open and the pancake rolled out to the
+    field, as hard as he could\ldots}
+\end{quotation}
+
+As readers (or listeners) of a story, we choose to view participating
+objects at varying levels of abstraction in order to predict their
+behavior. Dennett \cite{dennett:96} defines three levels of
+abstraction: (1) the \textit{physical stance}, (2) the \textit{design
+  stance} and (3) the \textit{intentional stance}. The physical stance
+deals with predictions about objects given their physical
+properties. The design stance is slightly more abstract and deals with
+concepts like purpose, function or design. Finally the intentional
+stance is concerned with belief, thinking and intentions. These are
+all cognitive strategies we use to predict and explain the actions of
+objects in our environment. Interestingly, in the process of reading
+the opening of the story about the fleeing pancake, readers and
+listeners experience the transition from one strategy to the next
+quite clearly. Initially, the pancake is interpreted from a physical
+stance or perhaps the more abstract design stance in terms of the
+purpose (i.e. to stave off hunger). It is only at the last adverbial
+phrase `as hard as he could' that we start to wonder whether we should
+adopt to the yet more abstract intentional stance and consider the
+pancake to be a rational agent.
+
+Given the fundamental nature of the distinction between animate and
+inanimate, it is perhaps not too surprising that it has proven to be
+useful in a variety of natural language processing tasks dealing with
+syntax, semantics, parsing, information retrieval, etc. Existing
+methods to automatically label natural language for animacy are
+usually rule based or apply machine learning techniques or make a
+combination of both. Common to most approaches is that they make use
+of semantic lexicons with information about animacy and syntactic cues
+in a text. Both feature types are relatively costly to obtain as they
+require large vocabularies or syntactic parsing systems, which are
+besides for a handful of western languages, often not available.
+
+In this paper we present a new method to automatically label texts for
+animacy. We will show that we can mostly do away with features that
+require syntactic parsing or semantic lexicons while still yielding
+competitive performance. We will focus on labeling animacy in stories
+because stories pose some interesting problems to automatic systems of
+animacy recognition. As the example of the fleeing pancake already
+made clear, quite some entities in stories, though inanimate in the
+`real' world, exhibit all behavioral capacities to consider them to be
+animate and rational agents. Such examples make a clear case for
+developing dynamic, data driven systems that do not rely too much on
+static and fixed world knowledge.
+
+The rest of this paper is structured as follows. We will start with an
+short overview of existing techniques to automatically label texts for
+animacy (\S\ref{sec:previous-work}). After a description a the corpus
+used in our study and the annotations of the corpus have been
+established (\S\ref{sec:data}), we will give an account of the
+computational models used in the study in Section \ref{sec:models}. We
+report on the empirical results in Section \ref{sec:results}. The last
+section offers our conclusions and possible directions for future
+research.
+
+\section{Previous Work}\label{sec:previous-work}
+
 In most of the literature on automatic animacy detection, people make
@@
 for example, define objects to be animate if they are alive and have
-the ability to move under their own will. \cite{orasan:07} define animacy
-in the context of anaphora resolution: ``if its referent can also be
-referred to using one of the pronouns he, she, him, her, his, hers,
-himself, herself, or a combination of such pro- nouns (e.g. his/her
-)'' Not only are some parts of these definitions somewhat circular, we
-believe that they are both essentialistic and not particularly suited
-for the rich and wondrous entities that live in the realm of
-folktales. A more sophisticated way to handle the distinction between
-animate and inanimate objects is given by Dennett CITE who argues that
-animacy and aliveness should be treated as epistemological stances
-rather than fixed states in the world.
+the ability to move under their own will. \cite{orasan:07} define
+animacy in the context of anaphora resolution: ``if its referent can
+also be referred to using one of the pronouns he, she, him, her, his,
+hers, himself, herself, or a combination of such pronouns
+(e.g. his/her)'' Not only are some parts of these definitions somewhat
+circular, we believe that they are both essentialistic and are not
+particularly suited for the rich and wondrous entities that live in
+the realm of stories. We choose to define animacy in terms of
+Dennett's intentional stance which comes down to the question whether
+``you decide to treat the object whose behavior is to be predicted as
+a rational agent'' \cite[17]{dennett:96}.
 
-\subsection{Methods for Automatic Animacy Detection}
+A handful of papers deal with automatic animacy detection. Most
+approaches make use of rule-based systems or machine learning systems
+with morphological and syntactic features. \cite{evans:00} present a
+rule-based system that makes use of the lexical-semantic database
+WordNet. They label each synset in WordNet for animacy. Using a
+variety of rules to detect the head of an NP, the use the fraction of
+synsets in which a particular noun occurs to arrive at a
+classification for animacy. \cite{orasan:01} extend their previous
+algorithm by first determining the animacy of senses from WordNet on
+the basis of an annotated corpus. They then apply a $k$-nearest
+neighbor classifier using a number of lexical and syntactic features
+and features derived from WordNet to arrive at a final animacy
+classification.
 
-\section{The Data, Annotation and Preprocessing}
+\cite{ovrelid:04,ovrelid:05,ovrelid:06,ovrelid:08,ovrelid:09} present
+a number of animacy classifiers that make use of syntactic and
+morphological features. Some of the features used are: how often has a
+noun been analyzed as `subject' or `object'? How often does a noun
+occur in a passive by-phrase? How often does a noun occur as a
+subject, immediately followed by animate personal pronouns or
+inanimate personal pronouns?  These features are then aggregated for
+each lemma after which a machine learning system (decision tree or
+$k$-nearest neighbor classifier) is trained. In \cite{bowman:12}
+similar approach is presented. In this study a maximum entropy
+classifier is trained on the basis of three feature types: (1)
+bag-of-words with and without their corresponding Part-of-Speech tags,
+(2) internal syntactic features such as the syntactic head and (3)
+external syntactic features that describe the dependency relation of a
+noun to a verb (i.e.\ subject relation, object relation etc.)  This is
+the only study that makes use of a fully labeled corpus for animacy.
+Not so much concerned with animacy detection, \cite{karsdorp:12}
+attempt to extract the cast (i.e.\ all characters) from a
+story. Similar to \cite{bowman:12} they rely on dependency tags to
+extract the subjects of direct and indirect speech.
 
-\subsection{The Data}
-Our corpus consists of 74 Dutch folktales from the collection
+\cite{bloem:13} present a model that attempts to generalize the
+animacy information in a lexical-semantic database of Dutch by
+augmenting `non-ambiguous' animate entries with contextual information
+from a large treebank of Dutch. They apply a $k$-nearest neighbor
+algorithm with distributional lexical features that aim to capture the
+association using measures like mutual information between a verb or
+adjective and a particular noun. The idea is that nouns that occur in
+similar contexts as animate nouns are more likely to be animate than
+nouns that occur more frequently in contexts similar to inanimate nouns.
+
+\cite{moore:13} present an interesting approach that combines a number
+of animacy classifiers in a voting scheme and aims at an interpretable
+and correctable model of animacy classification. They combine a
+variety of classifiers such as the WordNet-based approach of
+\cite{evans:00}, dictionary sources and systems for named entity
+recognition.
+
+% commentary on above approaches
+There are a number of problems with the approaches mentioned
+above. First, almost all of them heavily rely on costly features derived from
+lexical-semantic databases or syntactic parsing. For most languages in
+the world, however, we cannot rely on these resources. Second, animacy
+detection is often seen as a useful feature for a range of natural
+language processing techniques, such as anaphora resolution and
+syntactic parsing. The mutual dependence of these techniques and
+animacy detection, however, leads to a chicken and egg
+situation. Another problem with the approaches above is that they are
+lemma based which means that the models are generally insensitive to
+different usages of a particular word. A pancake is typically not an
+animate entity, but, as we have seen, whether it is animate or not is
+only to be determined in relation to the story in which it appears. We
+therefore need to develop a dynamic, data driven system that is token
+based and does not rely too much on static and fixed world knowledge.
+
+\section{The Data, Annotation and Preprocessing}\label{sec:data}
+
+Our corpus consists of 74 Dutch stories from the collection
 \textit{Volkssprookjes uit Nederland en Vlaanderen} compiled by
 \cite{sinninghe:78}. The collection contains retellings of popular
-folktales from the Netherlands and Flanders, such as \textit{The
+stories from the Netherlands and Flanders, such as \textit{The
   Bremen Town Musicians} (ATU 130) and \textit{The Table, the Ass, and
@@
 
-\subsection{Annotation}
-We labeled the corpus for animacy using the definition provided by
-%\cite{bowman:12}:
+We assigned to each word in the collection a binary label `animate' or
+`inanimate'. 11542 words were labeled as animate and 62926 words as
+inanimate. The vocabulary size is 5549 with 743 unique animate words
+and 5011 unique inanimate words. The annotations consist of 4627
+animate words that have been tagged as noun or name against 6878
+inanimate words.
 
-\subsection{Preprocessing}
+After tokenization with the tokenization module of the Python software
+package Pattern \cite{smedt:12} we fed all stories to the
+state of the art syntactic parser for Dutch, Alpino \cite{bouma:01}.
 
-\section{Experimental Setup}
+\section{Experimental Setup}\label{sec:models}
 This section describes our experimental setup including the features
 used, the machine learning models we applied and our methods of
-evaluation.
+evaluation.\footnote{The data set and the code to perform the
+  experiments are made available online at
+  \url{https://fbkarsdorp.github.io/animacy-detection}}.
 
@@
 
-\begin{description}
-\item[Lexical features]
+\subparagraph*{Lexical features}
 We take a sliding window approach where for each focus word (i.e.\ the
@@
 
-\item[Morphological Features]
+\subparagraph*{Morphological Features}
 For each word we extract its Part-of-Speech tag. For reasons of
@@
 
-\item[Syntactic Features]
+\subparagraph*{Syntactic Features}
 We extract the dependency tag for each word and its $n$ neighbors to
@@
 
-\item[Semantic Features]
+\subparagraph*{Semantic Features}
 For each focus word we add a 300 dimension vector representation that
@@
 words were given a vector were all dimensions are set to zero.
-
-\end{description}
 
@@
 
-\section{Results}
+Story-based training!
+
+\section{Results}\label{sec:results}
 
@@
 
-\section*{Acknowledgments}
+\subparagraph*{Acknowledgments}
 
@@
 
-Constantin Oraˇsan and Richard J. Evans. 2007. NP an- imacy identification for anaphora resolution. Journal of Artificial Intelligence Research (JAIR), 29:79–103.
-\begin{thebibliography}{50}
-\bibitem{sinninghe:78} Sinninghe J. \textit{Volkssprookjes uit
-    Nederland en Vlaanderen}. Kruseman, Den Haag, 1978.
-\bibitem{sklearn} Pedregosa, F. and Varoquaux, G. and Gramfort, A. and Michel, V.
-         and Thirion, B. and Grisel, O. and Blondel, M. and Prettenhofer, P.
-         and Weiss, R. and Dubourg, V. and Vanderplas, J. and Passos, A. and
-         Cournapeau, D. and Brucher, M. and Perrot, M. and Duchesnay,
-         E. `Scikit-Learn: Machine Learning in {P}ython'. In:
-         \textit{Journal of Machine Learning Research}, 12,
-         2825--2830, 2011.
-\bibitem{rijsbergen:79} {Van Rijsbergen}, C. \textit{Information
-    Retrieval}. Butterworths.
-\bibitem{bowman:12} Bowman S. and Chopra, H. `Automatic animacy
-  classification'. In \textit{Proceedings of the NAACL - HLT 2012
-    Student Research Workshop}, 2012.
-\bibitem{orasan:07} Or\u{a}san, C. and Evans, R. `NP animacy
-  identification for anaphora resolution'. In \textit{Journal of
-    Artificial Intelligence Research}, 29, 79--103, 2007.
-\end{thebibliography}
+\bibliography{paper}
 

### 2014-12-04 experiment.py
flexible fileinput

@@ -17,7 +17,7 @@ from gensim.models.word2vec import Word2Vec
 
 def load_data(limit=None):
     X, y = [[]], [[]]
-    with codecs.open("data/sinninghe-tagged.tsv", encoding="utf-8") as infile:
+    with codecs.open(sys.argv[2], encoding="utf-8") as infile:
         for i, line in enumerate(infile):
             if limit is not None and i >= limit:
                 break

### 2014-12-04 experiment.py
class weights

@@ -211,7 +211,7 @@ for experiment in experiments:
     y_train = le.fit_transform(y_train_docs)
     y_test = le.transform(y_test_docs)
     # initialize a classifier
-    clf = LogisticRegression(C=1.0)
+    clf = LogisticRegression(C=1.0, class_weight='auto')
     clf.fit(X_train, y_train)
     preds = clf.predict(X_test)
     print classification_report(y_test, preds)

### 2014-12-04 experiment.py
fix lemma, option for class weight

@@ -187,8 +187,8 @@ y_test_docs = [label for i in y_test_idx for label in y[i]]
 model = Word2Vec.load(sys.argv[1])
 
 # set up a number of experimental settings
-experiments = [('word',), ('word', 'pos'), ('word', 'pos', 'lemma'),
-               ('word', 'pos', 'lemma', 'rel'), tuple(FIELDNAMES)]
+experiments = [('word',), ('word', 'pos'), ('word', 'pos', 'root'),
+               ('word', 'pos', 'root', 'rel'), tuple(FIELDNAMES)]
 experiments = experiments + [experiment + ('embeddings', )
                              for experiment in experiments]
 experiments += [('embeddings', )]
@@ -211,7 +211,10 @@ for experiment in experiments:
     y_train = le.fit_transform(y_train_docs)
     y_test = le.transform(y_test_docs)
     # initialize a classifier
-    clf = LogisticRegression(C=1.0, class_weight='auto')
+    if sys.argv[3] == 'auto':
+        clf = LogisticRegression(C=1.0, class_weight='auto')
+    else:
+        clf = LogisticRegression(C=1.0)
     clf.fit(X_train, y_train)
     preds = clf.predict(X_test)
     print classification_report(y_test, preds)

### 2014-12-04 experiment.py
multiple classifiers

@@ -8,7 +8,8 @@ from sklearn.base import BaseEstimator
 from sklearn.cross_validation import train_test_split
 from sklearn.dummy import DummyClassifier
 from sklearn.feature_extraction import DictVectorizer
-from sklearn.linear_model import LogisticRegression
+from sklearn.linear_model import LogisticRegression, SGDClassifier
+from sklearn.svm import LinearSVC
 from sklearn.metrics import classification_report
 from sklearn.preprocessing import LabelEncoder
 
@@ -193,6 +194,12 @@ experiments = experiments + [experiment + ('embeddings', )
                              for experiment in experiments]
 experiments += [('embeddings', )]
 
+classifiers = {
+    'lr': LogisticRegression(C=1.0),
+    'sgd': SGDClassifier(n_iter=100, shuffle=True),
+    'svm': LinearSVC(),
+}
+
 for experiment in experiments:
     print "Features: %s" % ', '.join(experiment)
     if 'embeddings' in experiment and len(experiment) > 1:
@@ -211,10 +218,7 @@ for experiment in experiments:
     y_train = le.fit_transform(y_train_docs)
     y_test = le.transform(y_test_docs)
     # initialize a classifier
-    if sys.argv[3] == 'auto':
-        clf = LogisticRegression(C=1.0, class_weight='auto')
-    else:
-        clf = LogisticRegression(C=1.0)
+    clf = classifiers[sys.argv[3]]
     clf.fit(X_train, y_train)
     preds = clf.predict(X_test)
     print classification_report(y_test, preds)

### 2014-12-04 experiment.py
add name classifiers

@@ -219,6 +219,7 @@ for experiment in experiments:
     y_test = le.transform(y_test_docs)
     # initialize a classifier
     clf = classifiers[sys.argv[3]]
+    print clf.__class__.__name__
     clf.fit(X_train, y_train)
     preds = clf.predict(X_test)
     print classification_report(y_test, preds)

### 2014-12-04 experiment.py
fix classification on nouns...

@@ -231,7 +231,7 @@ for experiment in experiments:
             if w[3] in ('noun', 'name'):
                 noun_preds.append(i + j)
         i += len(X[idx])
-    print classification_report(preds[noun_preds], y_test[noun_preds])
+    print classification_report(y_test[noun_preds], preds[noun_preds])
 
     print "Fitting a majority vote DummyClassifier"
     dummy_clf = DummyClassifier(strategy='constant', constant=1)

### 2014-12-04 experiment.py
add knn

@@ -9,6 +9,7 @@ from sklearn.cross_validation import train_test_split
 from sklearn.dummy import DummyClassifier
 from sklearn.feature_extraction import DictVectorizer
 from sklearn.linear_model import LogisticRegression, SGDClassifier
+from sklearn.neighbors import KNeighborsClassifier
 from sklearn.svm import LinearSVC
 from sklearn.metrics import classification_report
 from sklearn.preprocessing import LabelEncoder
@@ -198,6 +199,7 @@ classifiers = {
     'lr': LogisticRegression(C=1.0),
     'sgd': SGDClassifier(n_iter=100, shuffle=True),
     'svm': LinearSVC(),
+    'knn': KNeighborsClassifier(weights='distance', metric='cosine')
 }
 
 for experiment in experiments:

### 2014-12-04 experiment.py
different nn algorithm

@@ -199,7 +199,7 @@ classifiers = {
     'lr': LogisticRegression(C=1.0),
     'sgd': SGDClassifier(n_iter=100, shuffle=True),
     'svm': LinearSVC(),
-    'knn': KNeighborsClassifier(weights='distance', metric='cosine')
+    'knn': KNeighborsClassifier(algorithm='brute', weights='distance', metric='cosine')
 }
 
 for experiment in experiments:

### 2014-12-04 experiment.py
syns to l2 space

@@ -187,6 +187,7 @@ X_test_docs = [X[i] for i in X_test_idx]
 y_test_docs = [label for i in y_test_idx for label in y[i]]
 # load the desired word2vec model
 model = Word2Vec.load(sys.argv[1])
+model.init_sims(replace=True)
 
 # set up a number of experimental settings
 experiments = [('word',), ('word', 'pos'), ('word', 'pos', 'root'),

### 2014-12-05 experiment.py
updates

@@ -187,7 +187,7 @@ X_test_docs = [X[i] for i in X_test_idx]
 y_test_docs = [label for i in y_test_idx for label in y[i]]
 # load the desired word2vec model
 model = Word2Vec.load(sys.argv[1])
-model.init_sims(replace=True)
+# model.init_sims(replace=True)
 
 # set up a number of experimental settings
 experiments = [('word',), ('word', 'pos'), ('word', 'pos', 'root'),
@@ -200,7 +200,7 @@ classifiers = {
     'lr': LogisticRegression(C=1.0),
     'sgd': SGDClassifier(n_iter=100, shuffle=True),
     'svm': LinearSVC(),
-    'knn': KNeighborsClassifier(algorithm='brute', weights='distance', metric='cosine')
+    'knn': KNeighborsClassifier(weights='distance')
 }
 
 for experiment in experiments:

### 2014-12-05 experiment.py
lower all words!

@@ -28,7 +28,7 @@ def load_data(limit=None):
                 y.append([])
             else:
                 fields = line.strip().split('\t')
-                X[-1].append([field if field else None for field in fields[:-2]])
+                X[-1].append([field.lower() if field else None for field in fields[:-2]])
                 assert X[-1]
                 y[-1].append(fields[-2])
     return X, y
@@ -116,7 +116,7 @@ class WordEmbeddings(BaseEstimator):
     def transform(self, X):
         # x is a document, word[0] is the word token
         return np.vstack([self.model[word[0]] if word[0] in self.model else
-                          np.zeros(model.layer1_size) for x in X for word in x])
+                          np.zeros(self.model.layer1_size) for x in X for word in x])
 
     def fit_transform(self, X, y=None):
         self.fit(X, y)

### 2014-12-05 experiment.py
better lowering...

@@ -28,7 +28,7 @@ def load_data(limit=None):
                 y.append([])
             else:
                 fields = line.strip().split('\t')
-                X[-1].append([field.lower() if field else None for field in fields[:-2]])
+                X[-1].append([field if field else None for field in fields[:-2]])
                 assert X[-1]
                 y[-1].append(fields[-2])
     return X, y
@@ -115,7 +115,7 @@ class WordEmbeddings(BaseEstimator):
 
     def transform(self, X):
         # x is a document, word[0] is the word token
-        return np.vstack([self.model[word[0]] if word[0] in self.model else
+        return np.vstack([self.model[word[0].lower()] if word[0].lower() in self.model else
                           np.zeros(self.model.layer1_size) for x in X for word in x])
 
     def fit_transform(self, X, y=None):

### 2014-12-05 experiment.py
weighted F1-scores

@@ -1,6 +1,7 @@
 import codecs
 import sys
 
+from functools import partial
 import numpy as np
 import scipy.sparse as sp
 
@@ -11,7 +12,7 @@ from sklearn.feature_extraction import DictVectorizer
 from sklearn.linear_model import LogisticRegression, SGDClassifier
 from sklearn.neighbors import KNeighborsClassifier
 from sklearn.svm import LinearSVC
-from sklearn.metrics import classification_report
+from sklearn.metrics import classification_report, precision_recall_fscore_support
 from sklearn.preprocessing import LabelEncoder
 
 from gensim.models.word2vec import Word2Vec
@@ -225,6 +226,7 @@ for experiment in experiments:
     print clf.__class__.__name__
     clf.fit(X_train, y_train)
     preds = clf.predict(X_test)
+    classification_report = partial(precision_recall_fscore_support, average='weighted')
     print classification_report(y_test, preds)
     print "Classification report on nouns:"
     noun_preds = []

### 2014-12-05 experiment.py
add triplet embeddings

@@ -1,7 +1,6 @@
 import codecs
 import sys
 
-from functools import partial
 import numpy as np
 import scipy.sparse as sp
 
@@ -12,11 +11,11 @@ from sklearn.feature_extraction import DictVectorizer
 from sklearn.linear_model import LogisticRegression, SGDClassifier
 from sklearn.neighbors import KNeighborsClassifier
 from sklearn.svm import LinearSVC
-from sklearn.metrics import classification_report, precision_recall_fscore_support
+from sklearn.metrics import classification_report
 from sklearn.preprocessing import LabelEncoder
 
 from gensim.models.word2vec import Word2Vec
-
+from gensim import matutils
 
 def load_data(limit=None):
     X, y = [[]], [[]]
@@ -107,6 +106,37 @@ class Windower(BaseEstimator):
         return self.transform(X)
 
 
+class TripletEmbeddings(BaseEstimator):
+    def __init__(self, model, summed=True):
+        self.model = model
+        self.summed = summed
+
+    def fit(self, X, y=None):
+        return self
+
+    def transform(self, X):
+        X_ = []
+        for d, doc in enumerate(X):
+            for i, word in enumerate(doc):
+                if self.summed:
+                    stack = np.vstack
+                else:
+                    stack = np.hstack
+                embeddings = stack(
+                    [self.model[doc[k][0].lower()] if
+                     (k >= 0 and k < len(doc) and doc[k][0].lower() in self.model) else
+                     np.zeros(self.model.layer1_size) for k in (i-1, i, i+1)])
+                if self.summed:
+                    embeddings = matutils.unitvec(embeddings.mean(axis=0)).astype(np.float32)
+                X_.append(embeddings)
+        X_ = np.vstack(X_)
+        return X_
+
+    def fit_transform(self, X, y=None):
+        self.fit(X, y)
+        return self.transform(X)
+
+
 class WordEmbeddings(BaseEstimator):
     def __init__(self, model):
         self.model = model
@@ -191,11 +221,12 @@ model = Word2Vec.load(sys.argv[1])
 # model.init_sims(replace=True)
 
 # set up a number of experimental settings
-experiments = [('word',), ('word', 'pos'), ('word', 'pos', 'root'),
-               ('word', 'pos', 'root', 'rel'), tuple(FIELDNAMES)]
-experiments = experiments + [experiment + ('embeddings', )
-                             for experiment in experiments]
-experiments += [('embeddings', )]
+#experiments = [('word',), ('word', 'pos'), ('word', 'pos', 'root'),
[…]

### 2014-12-05 experiment.py
revert changes

@@ -15,7 +15,7 @@ from sklearn.metrics import classification_report
 from sklearn.preprocessing import LabelEncoder
 
 from gensim.models.word2vec import Word2Vec
-from gensim import matutils
+
 
 def load_data(limit=None):
     X, y = [[]], [[]]
@@ -106,37 +106,6 @@ class Windower(BaseEstimator):
         return self.transform(X)
 
 
-class TripletEmbeddings(BaseEstimator):
-    def __init__(self, model, summed=True):
-        self.model = model
-        self.summed = summed
-
-    def fit(self, X, y=None):
-        return self
-
-    def transform(self, X):
-        X_ = []
-        for d, doc in enumerate(X):
-            for i, word in enumerate(doc):
-                if self.summed:
-                    stack = np.vstack
-                else:
-                    stack = np.hstack
-                embeddings = stack(
-                    [self.model[doc[k][0].lower()] if
-                     (k >= 0 and k < len(doc) and doc[k][0].lower() in self.model) else
-                     np.zeros(self.model.layer1_size) for k in (i-1, i, i+1)])
-                if self.summed:
-                    embeddings = matutils.unitvec(embeddings.mean(axis=0)).astype(np.float32)
-                X_.append(embeddings)
-        X_ = np.vstack(X_)
-        return X_
-
-    def fit_transform(self, X, y=None):
-        self.fit(X, y)
-        return self.transform(X)
-
-
 class WordEmbeddings(BaseEstimator):
     def __init__(self, model):
         self.model = model
@@ -221,11 +190,10 @@ model = Word2Vec.load(sys.argv[1])
 # model.init_sims(replace=True)
 
 # set up a number of experimental settings
-#experiments = [('word',), ('word', 'pos'), ('word', 'pos', 'root'),
-#               ('word', 'pos', 'root', 'rel'), tuple(FIELDNAMES)]
-#experiments = experiments + [experiment + ('embeddings', )
-#                             for experiment in experiments]
-#experiments += [experiment + ('tripletembeddings', ) for experiment in experiments]
+experiments = [('word',), ('word', 'pos'), ('word', 'pos', 'root'),
+               ('word', 'pos', 'root', 'rel'), tuple(FIELDNAMES)]
+experiments = experiments + [experiment + ('embeddings', )
+                             for experiment in experiments]
 experiments = [('embeddings', ), ('tripletembeddings',)]
 
 classifiers = {
[…]

### 2014-12-05 learning_curve.py
add learning curve

@@ -1,0 +1,49 @@
+import sys
+
+import seaborn as sb
+import numpy as np
+import pandas as pd
+
+from sklearn.linear_model import LogisticRegression
+from sklearn.metrics import precision_recall_fscore_support
+from sklearn.cross_validation import train_test_split
+from sklearn.preprocessing import LabelEncoder
+
+from gensim.models.word2vec import Word2Vec
+
+from experiment import FeatureStacker, WordEmbeddings, Windower, load_data
+from experiment import include_features
+
+
+model = Word2Vec.load(sys.argv[1])
+X, y = load_data(sys.argv[2])
+
+X_train_idx, X_test_idx, y_train_idx, y_test_idx = train_test_split(
+    range(len(X)), range(len(X)), test_size=0.2, random_state=2014)
+X_train_docs = [X[i] for i in X_train_idx]
+y_train_docs = [label for i in y_train_idx for label in y[i]]
+X_test_docs = [X[i] for i in X_test_idx]
+y_test_docs = [label for i in y_test_idx for label in y[i]]
+
+scores = np.zeros((10, 3))
+sizes = []
+for i, train_size in enumerate(np.arange(0.1, 1.1, 0.1)):
+    size = int(len(X_train_docs) * train_size)
+    sizes.append(size)
+    X_train = include_features(X_train_docs[:size], ('word', 'pos'))
+    X_test = include_features(X_test_docs, ('word', 'pos'))
+    features = FeatureStacker(('windower', Windower(window_size=3)),
+                              ('embeddings', WordEmbeddings(model)))
+    X_train = features.fit_transform(X_train)
+    X_test = features.transform(X_test)
+    le = LabelEncoder()
+    y_train = le.fit_transform(y_train_docs[:X_train.shape[0]])
+    y_test = le.transform(y_test_docs)
+    clf = LogisticRegression(C=1.0)
+    clf.fit(X_train, y_train)
+    preds = clf.predict(X_test)
+    scores[i] = precision_recall_fscore_support(y_test, preds, average='micro')[:3]
+
+df = pd.DataFrame(scores, index=sizes, columns=['precision', 'recall', 'F-score'])
+df.plot()
+sb.plt.savefig("learning_curve.pdf")

### 2014-12-05 experiment.py
small updates

@@ -17,9 +17,9 @@ from sklearn.preprocessing import LabelEncoder
 from gensim.models.word2vec import Word2Vec
 
 
-def load_data(limit=None):
+def load_data(filename, limit=None):
     X, y = [[]], [[]]
-    with codecs.open(sys.argv[2], encoding="utf-8") as infile:
+    with codecs.open(filename, encoding="utf-8") as infile:
         for i, line in enumerate(infile):
             if limit is not None and i >= limit:
                 break
@@ -175,70 +175,72 @@ def include_features(X, features):
     return [[[field for i, field in enumerate(word) if i in excluded]
              for word in doc] for doc in X]
 
-# read the data and extract all features
-X, y = load_data(limit=None)
-# split the data into a train and test set (this is based on documents, not words!)
-X_train_idx, X_test_idx, y_train_idx, y_test_idx = train_test_split(
-    range(len(X)), range(len(X)), test_size=0.2, random_state=1)
-# get the actual data by flattening the documents
-X_train_docs = [X[i] for i in X_train_idx]
-y_train_docs = [label for i in y_train_idx for label in y[i]]
-X_test_docs = [X[i] for i in X_test_idx]
-y_test_docs = [label for i in y_test_idx for label in y[i]]
-# load the desired word2vec model
-model = Word2Vec.load(sys.argv[1])
-# model.init_sims(replace=True)
+if __name__ == '__main__':
 
-# set up a number of experimental settings
-experiments = [('word',), ('word', 'pos'), ('word', 'pos', 'root'),
-               ('word', 'pos', 'root', 'rel'), tuple(FIELDNAMES)]
-experiments = experiments + [experiment + ('embeddings', )
-                             for experiment in experiments]
-experiments = [('embeddings', ), ('tripletembeddings',)]
+    # read the data and extract all features
+    X, y = load_data(sys.argv[2], limit=None)
+    # split the data into a train and test set (this is based on documents, not words!)
+    X_train_idx, X_test_idx, y_train_idx, y_test_idx = train_test_split(
+        range(len(X)), range(len(X)), test_size=0.2, random_state=1)
+    # get the actual data by flattening the documents
+    X_train_docs = [X[i] for i in X_train_idx]
+    y_train_docs = [label for i in y_train_idx for label in y[i]]
+    X_test_docs = [X[i] for i in X_test_idx]
+    y_test_docs = [label for i in y_test_idx for label in y[i]]
+    # load the desired word2vec model
[…]

### 2014-12-05 learning_curve.py
more elaborate output

@@ -1,5 +1,6 @@
 import sys
-
+import matplotlib
+matplotlib.use('Agg')
 import seaborn as sb
 import numpy as np
 import pandas as pd
@@ -25,25 +26,41 @@ y_train_docs = [label for i in y_train_idx for label in y[i]]
 X_test_docs = [X[i] for i in X_test_idx]
 y_test_docs = [label for i in y_test_idx for label in y[i]]
 
-scores = np.zeros((10, 3))
-sizes = []
-for i, train_size in enumerate(np.arange(0.1, 1.1, 0.1)):
-    size = int(len(X_train_docs) * train_size)
-    sizes.append(size)
-    X_train = include_features(X_train_docs[:size], ('word', 'pos'))
-    X_test = include_features(X_test_docs, ('word', 'pos'))
-    features = FeatureStacker(('windower', Windower(window_size=3)),
-                              ('embeddings', WordEmbeddings(model)))
-    X_train = features.fit_transform(X_train)
-    X_test = features.transform(X_test)
-    le = LabelEncoder()
-    y_train = le.fit_transform(y_train_docs[:X_train.shape[0]])
-    y_test = le.transform(y_test_docs)
-    clf = LogisticRegression(C=1.0)
-    clf.fit(X_train, y_train)
-    preds = clf.predict(X_test)
-    scores[i] = precision_recall_fscore_support(y_test, preds, average='micro')[:3]
 
-df = pd.DataFrame(scores, index=sizes, columns=['precision', 'recall', 'F-score'])
-df.plot()
+experiments = [('word',), ('word', 'pos'), ('word', 'pos', 'root'),
+               ('word', 'pos', 'root', 'rel')]
+experiments = experiments + [experiment + ('embeddings', )
+                             for experiment in experiments]
+experiments += [('embeddings', )]
+
+scores = np.zeros((10, len(experiments)))
+sizes = []
+
+for e, experiment in enumerate(experiments):
+    print experiment
+    for i, train_size in enumerate(np.arange(0.1, 1.1, 0.1)):
+        size = int(len(X_train_docs) * train_size)
+        sizes.append(size)
+        if experiment == ('embeddings', ):
+            features = FeatureStacker(('embeddings', WordEmbeddings(model)))
+            experiment = ('word', ) + experiment
+        else:
+            features = FeatureStacker(('windower', Windower(window_size=3)),
+                                      ('embeddings', WordEmbeddings(model)))
[…]

### 2014-12-05 learning_curve.py
whoops!

@@ -37,6 +37,7 @@ scores = np.zeros((10, len(experiments)))
 sizes = []
 
 for e, experiment in enumerate(experiments):
+    sizes = []
     print experiment
     for i, train_size in enumerate(np.arange(0.1, 1.1, 0.1)):
         size = int(len(X_train_docs) * train_size)

### 2014-12-06 learning_curve.py
fix

@@ -6,7 +6,7 @@ import numpy as np
 import pandas as pd
 
 from sklearn.linear_model import LogisticRegression
-from sklearn.metrics import precision_recall_fscore_support
+from sklearn.metrics import f1_score
 from sklearn.cross_validation import train_test_split
 from sklearn.preprocessing import LabelEncoder
 
@@ -16,7 +16,7 @@ from experiment import FeatureStacker, WordEmbeddings, Windower, load_data
 from experiment import include_features
 
 
-model = Word2Vec.load(sys.argv[1])
+model = Word2Vec.load_word2vec_format(sys.argv[1], binary=True)
 X, y = load_data(sys.argv[2])
 
 X_train_idx, X_test_idx, y_train_idx, y_test_idx = train_test_split(
@@ -58,10 +58,8 @@ for e, experiment in enumerate(experiments):
         clf = LogisticRegression(C=1.0)
         clf.fit(X_train, y_train)
         preds = clf.predict(X_test)
-    scores[i, e] = precision_recall_fscore_support(y_test, preds, average='micro')[2]
+        scores[i, e] = f1_score(y_test, preds, average='micro')
 
 df = pd.DataFrame(scores, index=sizes,
                   columns=['_'.join(exp) for exp in experiments])
-fig = sb.plt.figure()
-ax = df.plot()
-sb.plt.savefig("learning_curve.pdf")
+df.to_csv("learning_curve.csv")

### 2014-12-06 paper/paper.tex
update

@@
 \author[2]{Marten van der Meulen}
-\author[3]{Antal van den Bosch}
+\author[3]{Theo Meder}
+\author[4]{Antal van den Bosch}
 \affil[1]{Meertens Institute\\
@@
   \texttt{folgert.karsdorp@meertens.knaw.nl}}
-\affil[1]{Meertens Institute\\
+\affil[2]{Meertens Institute\\
   Amsterdam, The Netherlands\\
   \texttt{marten.van.der.meulen@meertens.knaw.nl}}
-\affil[2]{Radboud University\\
+\affil[3]{Meertens Institute\\
+  Amsterdam, The Netherlands\\
+  \texttt{theo.meder@meertens.knaw.nl}}
+\affil[4]{Radboud University\\
   Nijmegen, The Netherlands\\
   \texttt{a.vandenbosch@let.ru.nl}}
-\authorrunning{F. Karsdorp and M. van der Meulen and A. van den Bosch}
-\Copyright{Folgert Karsdorp and Marten van der Meulen and Antal van den Bosch}
-\subjclass{machine learning}
-\keywords{machine learning}
+%\authorrunning{F. Karsdorp and M. van der Meulen and A. van den Bosch}
+%\Copyright{Folgert Karsdorp and Marten van der Meulen and Antal van
+%  den Bosch}
+\authorrunning{F. Karsdorp and M. van der Meulen and T. Meder and A. van den Bosch}
+\Copyright{Folgert Karsdorp and Marten van der Meulen and Theo Meder
+  and Antal van den Bosch}
+\subjclass{I.2.7 Natural Language Processing}
+\keywords{animacy detection, deep learning, folklore}
 
@@
 \begin{abstract}
-\blindtext[1]
+  This paper presents a linguistically uninformed computational model
+  for animacy classification. The model makes use of word $n$-grams in
+  combinations with lower dimensional word embedding representations
+  that were learned from a web-scale corpus. We compare the model to a
+  number of linguistically informed models that use features such as
+  dependency tags and show competitive results.
 \end{abstract}
@@
 mates. The fundamental nature of the distinction between animate and
-inanimate appears in the early acquisition of the division. Children
-of less than 6 months old are well able to make the division. Recent
-brain research shows that the distinction appears in the organization
-of the ventral temporal cortex in both humans and other primates. For
-some researchers this provides evidence for the idea that the division
+inanimate appears in the early acquisition of the division
+\cite{opfer:02}. Children of less than 6 months old are well able to
+make the division. Recent brain research shows that the distinction
+appears in the organization of the brain (e.g.~\cite{gao:12}). For some
+researchers this provides evidence for the idea that the division
 between animate and inanimate is an innate part of how we see the
@@
 animacy and aliveness are better to be treated as epistemological
-stances rather than fixed states in the world. Not ineffable qualia
-but behavioral capacity defines our stance towards objects. Depending
-on whether people \textit{think} that an object is animate, they
-utilize different cognitive strategies to explain and predict the
-actions of those objects. The following opening of a well-known story
-will make this point more clear:
+stances rather than fixed states in the world \cite{dennett:96}. Not
+ineffable qualia but behavioral capacity defines our stance towards
+objects. Depending on whether people \textit{think} that an object is
+animate, they utilize different cognitive strategies to explain and
+predict the actions of those objects. The following opening of a
+well-known story will make this point more clear:
 
@@
 useful in a variety of natural language processing tasks dealing with
-syntax, semantics, parsing, information retrieval, etc. Existing
-methods to automatically label natural language for animacy are
-usually rule based or apply machine learning techniques or make a
-combination of both. Common to most approaches is that they make use
-of semantic lexicons with information about animacy and syntactic cues
-in a text. Both feature types are relatively costly to obtain as they
-require large vocabularies or syntactic parsing systems, which are
-besides for a handful of western languages, often not available.
+e.g.\ anaphora resolution and dependency parsing
+\cite{orasan:07,lee:13,ovr:niv:07}. Existing methods to automatically
+label natural language for animacy are usually rule based or apply
+machine learning techniques or make a combination of both. Common to
+most approaches is that they make use of semantic lexicons with
+information about animacy and syntactic cues in a text. Both feature
+types are relatively costly to obtain as they require large
+vocabularies or syntactic parsing systems, which are besides for a
+handful of western languages, often not available.
 
-In this paper we present a new method to automatically label texts for
-animacy. We will show that we can mostly do away with features that
-require syntactic parsing or semantic lexicons while still yielding
-competitive performance. We will focus on labeling animacy in stories
-because stories pose some interesting problems to automatic systems of
-animacy recognition. As the example of the fleeing pancake already
-made clear, quite some entities in stories, though inanimate in the
-`real' world, exhibit all behavioral capacities to consider them to be
-animate and rational agents. Such examples make a clear case for
-developing dynamic, data driven systems that do not rely too much on
-static and fixed world knowledge.
+In this paper we present a new linguistically uninformed model to
+automatically label texts for animacy. We will show that we can (mostly)
+do away with features that require syntactic parsing or semantic
+lexicons while still yielding competitive performance. We will focus
+on labeling animacy in stories because stories pose some interesting
+problems to automatic systems of animacy recognition. As the example
+of the fleeing pancake already made clear, quite some entities in
+stories, though inanimate in the `real' world, exhibit all behavioral
+capacities to consider them to be animate and rational agents. Such
+examples make a clear case for developing dynamic, data driven systems
+that do not rely too much on static and fixed world knowledge.
 
@@
 short overview of existing techniques to automatically label texts for
-animacy (\S\ref{sec:previous-work}). After a description a the corpus
-used in our study and the annotations of the corpus have been
-established (\S\ref{sec:data}), we will give an account of the
-computational models used in the study in Section \ref{sec:models}. We
-report on the empirical results in Section \ref{sec:results}. The last
-section offers our conclusions and possible directions for future
-research.
+animacy (\S\ref{sec:previous-work}). After a description of the corpus
+used in our study and how the annotations of the corpus have been
+established (\S\ref{sec:data}), we will give an account of our
+computational models in Section \ref{sec:models}. We report on the
+empirical results in Section \ref{sec:results}. The last section
+offers our conclusions and possible directions for future research.
 
@@
 hers, himself, herself, or a combination of such pronouns
-(e.g. his/her)'' Not only are some parts of these definitions somewhat
-circular, we believe that they are both essentialistic and are not
+(e.g. his/her)''. Not only are some parts of these definitions somewhat
+circular, we believe that they are both essentialistic and not
 particularly suited for the rich and wondrous entities that live in
@@
 ``you decide to treat the object whose behavior is to be predicted as
-a rational agent'' \cite[17]{dennett:96}.
+a rational agent'' \cite[pp. 17]{dennett:96}.
 
@@
 WordNet. They label each synset in WordNet for animacy. Using a
-variety of rules to detect the head of an NP, the use the fraction of
+variety of rules to detect the head of an NP, they use the fraction of
 synsets in which a particular noun occurs to arrive at a
@@
 each lemma after which a machine learning system (decision tree or
-$k$-nearest neighbor classifier) is trained. In \cite{bowman:12}
+$k$-nearest neighbor classifier) is trained. In \cite{bowman:12} a
 similar approach is presented. In this study a maximum entropy
@@
 There are a number of problems with the approaches mentioned
-above. First, almost all of them heavily rely on costly features derived from
-lexical-semantic databases or syntactic parsing. For most languages in
-the world, however, we cannot rely on these resources. Second, animacy
-detection is often seen as a useful feature for a range of natural
-language processing techniques, such as anaphora resolution and
-syntactic parsing. The mutual dependence of these techniques and
-animacy detection, however, leads to a chicken and egg
-situation. Another problem with the approaches above is that they are
-lemma based which means that the models are generally insensitive to
-different usages of a particular word. A pancake is typically not an
-animate entity, but, as we have seen, whether it is animate or not is
-only to be determined in relation to the story in which it appears. We
-therefore need to develop a dynamic, data driven system that is token
-based and does not rely too much on static and fixed world knowledge.
+above. First, almost all of them heavily rely on costly,
+linguistically informed features derived from lexical-semantic
+databases or syntactic parsing. For most languages in the world,
+however, we cannot rely on these resources. Second, animacy detection
+is often seen as a useful feature for a range of natural language
+processing techniques, such as anaphora resolution and syntactic
+parsing. The mutual dependence of these techniques and animacy
+detection, however, leads to a chicken and egg situation. Another
+problem with the approaches above is that they are lemma based which
+means that the models are generally insensitive to different usages of
+a particular word. A pancake is typically not an animate entity, but,
+as we have seen, whether it is animate or not is only to be determined
+in relation to the story in which it appears. We therefore need to
+develop a dynamic, data driven system that is token based and does not
+rely too much on static and fixed world knowledge.
 
@@
 \cite{sinninghe:78}. The collection contains retellings of popular
-stories from the Netherlands and Flanders, such as \textit{The
-  Bremen Town Musicians} (ATU 130) and \textit{The Table, the Ass, and
-  the Stick } (ATU 563). Some of the less well-know stories are
+stories from the Netherlands and Flanders, such as \textit{The Bremen
+  Town Musicians} (ATU 130) and \textit{The Table, the Ass, and the
+  Stick } (ATU 563). Some of the less well-know stories are
 \textit{Cock, Hen, Duck, Pin, and Needle on a Journey} (AT 0210) and
 (ATU 780) \textit{The Singing Bone}. A digital version of the
-collection available in the Dutch Folktale Database from the Meertens
-Institute (corpus SINVSUNV.20E).\footnote{See
+collection is available in the Dutch Folktale Database from the
+Meertens Institute (corpus SINVSUNV.20E).\footnote{See
   \url{http://www.verhalenbank.nl}} We chose this collection because
@@
 
+We formulate the problem of animacy detection as a classification
+problem where the goal is to assign to each word in a document a label
+indicating whether it represents an animate or an inanimate word.
+
+\subsection{Evaluation}
+
+Animate words are far outnumbered by inanimate words in our
+collection. Reporting accuracy scores would therefore provide skewed
+results too much in favor of the majority category. The rarity of
+animate words makes evaluation measures such as the well-known $F1$-score
+more appropriate. For all experiments we report on the precision,
+recall and $F1$-score \cite{rijsbergen:79}. In most of the literature on
+animacy detection results are only given for the classification of
+nouns or noun phrases. We will report the results for all words in a
+text and on those specifically for nouns separately.
+
+In real world applications, an animacy detection system will most
+likely be faced with completely new texts instead of single words. It
+is therefore important to construct a training and test procedure in
+such a way that it mimics this situation as closely as possible. If we
+would make a random split of say 80\% of the data for training and
+20\% for testing on the word level, we run the risk of mixing training
+data with test data making it too easy for a system to rely on words
+it has seen from the same text. \cite{bowman:12} fall into this trap
+by making a random split in their data on the sentence level. It is
+very likely that in their setup sentences from the same document are
+present in both the training data and the test data which makes their
+evaluation unrealistic. To circumvent this problem, we train our model
+on 80\% of the stories (n=59) and test on the remaining 15 stories.
+
 \subsection{Features}
@@
 
-\subparagraph*{Lexical features}
+\subsubsection{Lexical features}
 We take a sliding window approach where for each focus word (i.e.\ the
@@
 word itself. In all experiments we set $n$ to 3. In addition to the
-word forms, we extract for each word in a window the lemmata
-corresponding to the words.
+word forms, for each word in a window we extract its lemma.
 
-\subparagraph*{Morphological Features}
+\subsubsection{Morphological Features}
 For each word we extract its Part-of-Speech tag. For reasons of
-comparability for choose to use the tags as provided by the
-output of the syntactic parser, instead of a more specialized
+comparability we choose to use the tags as provided by the output of
+the syntactic parser Alpino, instead of a more specialized
 Part-of-Speech tagger. Again, we take a sliding window approach and
-extract the Part-of-Speech tags for $n$ word left of the focus word
-and $n$ word right of the focus word as well as the tag of the focus
-word itself.
+extract the Part-of-Speech tags for three word left of the focus word,
+three word right of the focus word and the tag of the focus word
+itself.
 
-\subparagraph*{Syntactic Features}
+\subsubsection{Syntactic Features}
 We extract the dependency tag for each word and its $n$ neighbors to
 the right and to the left as provided by the syntactic parser
-Alpino. Animate entities tend to take subject or object positions in a
-sentences which is why this feature is expected and has proven to
-perform rather well.
+Alpino. Animate entities tend to take the position of subject or
+object in a sentence which is why this feature is expected and has
+proven to perform rather well.
 
-\subparagraph*{Semantic Features}
-For each focus word we add a 300 dimension vector representation that
-was learned by applying word2vec on the Dutch version of CoW. Missing
-words were given a vector were all dimensions are set to zero.
+\subsubsection{Semantic Features} The most innovative feature we have
+included in our model is concerned with semantic similarity. In his
+\textit{Philophische Untersuchungen} Wittgenstein already suggests
+that ``Die Bedeutung eines Wortes ist sein Gebrauch in der Sprache''
+(PI 43). The well-known insight in computational linguistics that the
+meaning of words can be approximated by comparing the linguistic
+contexts in which words appear, is an implementation of this very
+idea. The idea is very simple, and equally strong: words that often
+co-appear with the same set of words, will have a more similar
+meaning. Recently, there has been a lot of interest in procedures that
+can automatically induce so-called `word-embeddings' from large,
+unannotated collections of texts. These models typically attempt to
+learn a lower-dimensional vector for each word in the vocabulary which
+captures the typical co-occurrence patterns of the word in the
+corpus. The similarity between words can then be approximated by
+applying arithmetics to these vectors of word embeddings.
+
+We have trained word embeddings with 300 dimensions using the popular
+skip-gram architecture \cite{mikolov:13} on the Dutch corpus of COW
+(COrpora from the Web).\footnote{The word embedding vectors have been
+  published online at XXX.} COW is a collection of linguistically
+processed web corpora for English, Dutch, Spanish, French, Swedish and
+German \cite{schaefer:12}. The 2014 Dutch corpus contains 6.8 billion
+word tokens. The idea behind using the word embeddings is that
+similarities between animate words can be estimated by inspecting the
+context in which they occur. The hypothesis is that the word
+embeddings of an animate word are more similar to those of other
+animate words than than to those of inanimate words. To give an
+example, the 5 nearest neighbors of \textit{Assepoester}
+(`Cinderella') are: \textit{Sneeuwwitje} (`Snow White') (sim = 0.807),
+\textit{Roodkapje} (`Little Red Riding Hood') (sim = 0.747),
+\textit{Doornroosje} (`Sleeping Beauty') (sim = 0.715),
+\textit{Repelsteeltje} (`Rumpelstiltskin') (sim = 0.669) and
+\textit{Alladin} (0.645). In contrast, the 5 most similar words to
+\textit{boek} (`book') are: \textit{boekje} (`book-\textsc{dim}')
+(sim=0.8), \textit{kinderboek} (`children's book') (0.757),
+\textit{stripboek} (`comic') (0.713), \textit{kortverhaal} (`short
+story') (0.709), \textit{jeugdboek} (`youth book') (0.706).
+
+We add the word embedding vectors to our animacy model as follows. We
+augment each word in the data with its corresponding 300 dimension
+word embeddings vector. Out-of-vocabulary words were given a vector were all
+dimensions are set to zero.
 
 \subsection{Models}
-We make use of a Logistic Regression classifier also know as a Maximum
-Entropy classifier with L2 regularization as implemented in
-\cite{sklearn}. In all experiments, we set the regularization strength
-parameter $C$ to 1.
+We employ a Maximum Entropy classifier with L2 regularization as
+implemented in \cite{sklearn}. In all experiments, we set the
+regularization strength parameter $C$ to 1.
 
@@
 
-\subsection{Evaluation}
-
-Animate words are far outnumbered by the inanimate words in our
-collection. Reporting accuracy scores would therefore provide
-distorted results to much in favor of the majority category. The
-rarity of animate words make evaluation measures such as the
-well-known F1-score more appropriate. For all experiments we report on
-the precision, recall and F1-score \cite{rijsbergen:79}. In most of
-the literature on animacy detection results are only given for the
-classification of nouns. We will report the results for all words in a
-text and on those specifically for nouns separately.
-
-Story-based training!
-
 \section{Results}\label{sec:results}
 
-\begin{table*}
+\begin{table}
 \centering
@@
 \toprule
-features                                                 &           & precision & recall & $F1$ \\ \midrule
-\multirow{2}{*}{word}                                    & animate   & 0.93      & 0.80   & 0.86 \\
-                                                         & inanimate & 0.96      & 0.99   & 0.98 \\
-\multirow{2}{*}{word + PoS}                              & animate   & 0.92      & 0.88   & 0.90 \\
-                                                         & inanimate & 0.98      & 0.99   & 0.98 \\
-\multirow{2}{*}{word + PoS + lemma}                      & animate   & 0.92      & 0.88   & 0.90 \\
-                                                         & inanimate & 0.98      & 0.99   & 0.98 \\
-\multirow{2}{*}{word + PoS + lemma + dep}                & animate   & 0.93      & 0.88   & 0.90 \\
-                                                         & inanimate & 0.98      & 0.99   & 0.98 \\
-\multirow{2}{*}{word + embeddings}                       & animate   & 0.93      & 0.89   & 0.91 \\
-                                                         & inanimate & 0.98      & 0.99   & 0.98 \\
-\rowcolor{Gray}                                          & animate   & 0.93      & 0.93   & 0.93 \\
-\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings} & inanimate & 0.99      & 0.99   & 0.99 \\
-\multirow{2}{*}{word + PoS + lemma + embeddings}         & animate   & 0.93      & 0.93   & 0.93 \\
-                                                         & inanimate & 0.99      & 0.99   & 0.99 \\
-\multirow{2}{*}{word + PoS + lemma + dep + embeddings}   & animate   & 0.92      & 0.93   & 0.93 \\
-                                                         & inanimate & 0.99      & 0.99   & 0.99 \\
-\multirow{2}{*}{embeddings}                              & animate   & 0.90      & 0.76   & 0.83 \\
-                                                         & inanimate & 0.96      & 0.98   & 0.97 \\
+features                                                 &           & precision & recall & $F1$ & avg $F1$              \\ \midrule
+\multirow{2}{*}{word}                                    & animate   & 0.93      & 0.80   & 0.86 & \multirow{2}{*}{0.96} \\
+                                                         & inanimate & 0.96      & 0.99   & 0.98 &                       \\
+\multirow{2}{*}{word + PoS}                              & animate   & 0.92      & 0.88   & 0.90 & \multirow{2}{*}{0.97} \\
+                                                         & inanimate & 0.98      & 0.99   & 0.98 &                       \\
+\multirow{2}{*}{word + PoS + lemma}                      & animate   & 0.93      & 0.89   & 0.91 & \multirow{2}{*}{0.97} \\
+                                                         & inanimate & 0.98      & 0.99   & 0.98 &                       \\
+\multirow{2}{*}{word + PoS + lemma + dep}                & animate   & 0.93      & 0.89   & 0.91 & \multirow{2}{*}{0.97} \\
+                                                         & inanimate & 0.98      & 0.99   & 0.98 &                       \\
+\multirow{2}{*}{word + embeddings}                       & animate   & 0.92      & 0.93   & 0.92 & \multirow{2}{*}{0.98} \\
+                                                         & inanimate & 0.99      & 0.99   & 0.99 &                       \\
+\rowcolor{Gray}                                          & animate   & 0.92      & 0.93   & 0.93 &                       \\
+\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings} & inanimate & 0.99      & 0.99   & 0.99 & \multirow{-2}{*}{0.98}\\
+\multirow{2}{*}{word + PoS + lemma + embeddings}         & animate   & 0.93      & 0.93   & 0.93 & \multirow{2}{*}{0.98} \\
+                                                         & inanimate & 0.99      & 0.99   & 0.99 &                       \\
+\multirow{2}{*}{word + PoS + lemma + dep + embeddings}   & animate   & 0.92      & 0.93   & 0.93 & \multirow{2}{*}{0.98} \\
+                                                         & inanimate & 0.99      & 0.99   & 0.99 &                       \\
+\multirow{2}{*}{word embeddings}                         & animate   & 0.90      & 0.92   & 0.91 & \multirow{2}{*}{0.97} \\
+                                                         & inanimate & 0.99      & 0.98   & 0.98 &                       \\
 
@@
 \label{tab:results-all}
-\end{table*}
+\end{table}
 
-\begin{table*}
+In table \ref{tab:results-all} we present the results for all nine
+models on the complete data set. For each model we report the
+precision, recall and $F1$-score for the animate words and the
+inanimate words. All models perform well on classifying inanimate
+words. However, since this is the majority class, it is far more
+interesting to compare the performance of the models on the animate
+instances. It is interesting to observe that the `simple' $n$-gram
+word model already performs rather well. Adding more features, like
+Part-of-Speech, lemmata, etcetera, only has a positive impact on the
+recall of the model leaving the precision untouched. As can be
+observed from the table, the rather expensive dependency features show
+barely any improvement. The model that only uses word embedding
+features, is one of the best performing models. This was slightly
+unexpected, since this is the model in which features are context
+insensitive. However, the model that combines the word embeddings with
+the word $n$-gram features outperforms the other linguistically
+informed models that make use of lemmata, Part-of-Speech tags or
+dependency information. The best performance is achieved by the model
+that combines the word features, Part-of-Speech tags and the word
+embeddings with an $F1$-score of 0.93 for animate words and 0.99 on
+inanimate words. Adding more features does not result in any more
+performance gain.
+
+\begin{table}
 \centering
@@
 \toprule
-features                                               &           & precision & recall & $F1$ \\ \midrule
-\multirow{2}{*}{word}                                  & animate   & 0.64      & 0.96   & 0.77 \\
-                                                       & inanimate & 0.98      & 0.81   & 0.89 \\
-\multirow{2}{*}{word + PoS}                            & animate   & 0.80      & 0.92   & 0.86 \\
-                                                       & inanimate & 0.95      & 0.89   & 0.92 \\
-\multirow{2}{*}{word + PoS + lemma}                    & animate   & 0.80      & 0.92   & 0.86 \\
-                                                       & inanimate & 0.95      & 0.89   & 0.92 \\
-\multirow{2}{*}{word + PoS + lemma + dep}              & animate   & 0.80      & 0.93   & 0.86 \\
-                                                       & inanimate & 0.96      & 0.88   & 0.92 \\
-\multirow{2}{*}{word + embeddings}                     & animate   & 0.83      & 0.93   & 0.88 \\
-                                                       & inanimate & 0.96      & 0.90   & 0.93 \\
-\rowcolor{Gray}                                        & animate   & 0.92      & 0.92   & 0.92 \\
-\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings} & inanimate & 0.95    & 0.95   & 0.95 \\
-\multirow{2}{*}{word + PoS + lemma + embeddings}       & animate   & 0.92      & 0.92   & 0.92 \\
-                                                       & inanimate & 0.95      & 0.95   & 0.95 \\
-\multirow{2}{*}{word + PoS + lemma + dep + embeddings} & animate   & 0.91      & 0.92   & 0.92 \\
-                                                       & inanimate & 0.95      & 0.95   & 0.95 \\
-\multirow{2}{*}{embeddings}                            & animate   & 0.66      & 0.88   & 0.75 \\
-                                                       & inanimate & 0.94      & 0.81   & 0.87 \\
+features                                               &           & precision & recall & $F1$ & avg $F1$              \\ \midrule
+\multirow{2}{*}{word}                                  & animate   & 0.96      & 0.64   & 0.77 & \multirow{2}{*}{0.84} \\
+                                                       & inanimate & 0.81      & 0.98   & 0.89 &                       \\
+\multirow{2}{*}{word + PoS}                            & animate   & 0.92      & 0.80   & 0.86 & \multirow{2}{*}{0.90} \\
+                                                       & inanimate & 0.89      & 0.95   & 0.92 &                       \\
+\multirow{2}{*}{word + PoS + lemma}                    & animate   & 0.93      & 0.82   & 0.87 & \multirow{2}{*}{0.90} \\
+                                                       & inanimate & 0.89      & 0.96   & 0.92 &                       \\
+\multirow{2}{*}{word + PoS + lemma + dep}              & animate   & 0.93      & 0.82   & 0.87 & \multirow{2}{*}{0.91} \\
+                                                       & inanimate & 0.90      & 0.96   & 0.93 &                       \\
+\multirow{2}{*}{word + embeddings}                     & animate   & 0.92      & 0.91   & 0.92 & \multirow{2}{*}{0.94} \\
+                                                       & inanimate & 0.95      & 0.95   & 0.95 &                       \\
+\rowcolor{Gray}                                        & animate   & 0.92      & 0.92   & 0.92 &                       \\
+\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings} & inanimate & 0.95    & 0.95   & 0.95 & \multirow{-2}{*}{0.94}\\
+\multirow{2}{*}{word + PoS + lemma + embeddings}       & animate   & 0.92      & 0.91   & 0.92 & \multirow{2}{*}{0.94} \\
+                                                       & inanimate & 0.95      & 0.95   & 0.95 &                       \\
+\multirow{2}{*}{word + PoS + lemma + dep + embeddings} & animate   & 0.92      & 0.91   & 0.92 & \multirow{2}{*}{0.94} \\
+                                                       & inanimate & 0.94      & 0.95   & 0.95 &                       \\
+\multirow{2}{*}{word embeddings}                       & animate   & 0.88      & 0.90   & 0.89 & \multirow{2}{*}{0.92} \\
+                                                       & inanimate & 0.94      & 0.92   & 0.93 &                       \\
 
@@
 \label{tab:results-noun}
-\end{table*}
+\end{table}
+
+Table \ref{tab:results-noun} presents the results for all nouns and
+names in the data set. The best performance is again achieved by the
+model that combines the word features with the Part-of-Speech tags and
+word embeddings, resulting in an $F1$-score of 0.92 for animate
+instances and 0.95 for inanimate instances.
 
 \section{Concluding Remarks}
+
+The approach taken in this paper to create a model for animacy
+classification using linguistically uninformed features proves to be
+successful. We compared the performance of linguistically informed
+models (using features such as Part-of-Speech and dependency tags) to
+models that make use of lower dimensional representations of the
+data. With the exception of the model that solely makes use of these
+representations, all models benefit from adding these features. The
+model that requires the least linguistic information (word $n$-grams
+plus word embeddings) outperforms all linguistically informed
+models. The best results are reported by the model that combines word
+$n$-grams with Part-of-Speech $n$-grams and word embeddings.
+
+We have the following recommendation for future research. Natural
+language processing models such as co-reference resolution or
+linguistic parsing could benefit from a module that filters animate
+from inanimate candidate words. Since these models typically depend
+heavily on linguistic features, it is important that additional
+features, such as animacy, are not dependent on these features as
+well. Our linguistically uninformed model for animacy detection
+provides such a module.
 

### 2014-12-06 learning_curve.py
load

@@ -16,7 +16,7 @@ from experiment import FeatureStacker, WordEmbeddings, Windower, load_data
 from experiment import include_features
 
 
-model = Word2Vec.load_word2vec_format(sys.argv[1], binary=True)
+model = Word2Vec.load(sys.argv[1])
 X, y = load_data(sys.argv[2])
 
 X_train_idx, X_test_idx, y_train_idx, y_test_idx = train_test_split(

### 2014-12-06 crf.py
add crf

@@ -1,0 +1,93 @@
+import codecs
+import sys
+from sklearn.cross_validation import train_test_split
+from gensim.models.word2vec import Word2Vec
+from sklearn.metrics import classification_report
+import pycrfsuite
+
+def load_data(filename):
+    X, y = [[[]]], [[[]]]
+    with codecs.open(filename, encoding='utf-8') as infile:
+        for i, line in enumerate(infile):
+            if line.startswith('<FB/>'):
+                X.append([[]])
+                y.append([[]])
+            else:
+                fields = line.strip().split('\t')
+                X[-1][-1].append([field if field else None for field in fields[:-2]])
+                y[-1][-1].append(fields[-2])
+                if fields[0] in '?.!': # sentence breaks TODO: make this more robust
+                    X[-1].append([])
+                    y[-1].append([])
+    return X, y
+
+model = Word2Vec.load(sys.argv[1])
+X, y = load_data(sys.argv[2])
+
+X_train_idx, X_test_idx, y_train_idx, y_test_idx = train_test_split(
+    range(len(X)), range(len(X)), test_size=0.2, random_state=1)
+
+X_train = [sent for i in X_train_idx for sent in X[i]]
+y_train = [label for i in y_train_idx for label in y[i]]
+X_test = [sent for i in X_test_idx for sent in X[i]]
+y_test = [label for i in y_test_idx for label in y[i]]
+
+def word2features(sent, i, features):
+    FIELDNAMES = ['word', 'root', 'lcat', 'pos','rel', 'sense', 'frame',
+                  'special','noun_det', 'noun_countable', 'noun_number',
+                  'verb_auxiliary', 'verb_tense', 'verb_complements',
+                  'animate', 'reference']
+    header = {w: k for k, w in enumerate(FIELDNAMES)}
+    included = map(header.get, features)
+    features = {}
+    for j in range(i-2, i+3):
+        if j >=0 and j < len(sent):
+            for k, field in enumerate(sent[j]):
+                if k in included:
+                    features['%s%s-%s' % ('-' if j < i else '+' if j > i else '',
+                                             abs(i-j), k)] = field.lower()
+    if 'embeddings' in features:
+        word = sent[i][0].lower()
+        if word in model:
+            features.update({'w2v:%s' % i: v for i, v in enumerate(model[word])})
+    return features
+
+def sent2features(sent, features):
[…]

### 2014-12-06 experiment.py

@@ -114,7 +114,7 @@ class WordEmbeddings(BaseEstimator):
         return self
 
     def transform(self, X):
-        # x is a document, word[0] is the word token
+        # x is a document, word[0] is the word tokenq
         return np.vstack([self.model[word[0].lower()] if word[0].lower() in self.model else
                           np.zeros(self.model.layer1_size) for x in X for word in x])
 

### 2014-12-06 cow.py
update cow with vvb

@@ -1,9 +1,9 @@
 import codecs
 import logging
 import os
-import re
 
 from gensim.models.word2vec import Word2Vec
+from gensim.utils import tokenize
 
 logging.basicConfig(format='%(asctime)s : %(levelname)s : %(message)s',
                     level=logging.INFO)
@@ -12,6 +12,7 @@ class CowReader(object):
     root = '/vol/bigdata/corpora/COW/NLCOW'
     dirs = ['nlcow14ax01', 'nlcow14ax02', 'nlcow14ax03', 'nlcow14ax04',
             'nlcow14ax05', 'nlcow14ax06', 'nlcow14ax07']
+    vvb = '/vol/tensusers/fkarsdorp/vvb.tokenized.txt'
 
     def __iter__(self):
         for directory in CowReader.dirs:
@@ -29,7 +30,10 @@ class CowReader(object):
                         word, pos, lemma = line.strip().split('\t')
                         if pos not in ('$.', 'punc'):
                             sentence.append(word.lower())
+        with codecs.open(CowReader.vvb, encoding='utf-8') as vvb:
+            for sentence in vvb:
+                yield list(tokenize(sentence, lowercase=True))
 
 sentences = CowReader()
 model = Word2Vec(sentences, size=300, window=10, min_count=10, workers=20)
-model.save("/vol/tensusers/fkarsdorp/cow.w2v")
+model.save("/vol/tensusers/fkarsdorp/cow-vvb.w2v")

### 2014-12-06 experiment.py

@@ -114,7 +114,7 @@ class WordEmbeddings(BaseEstimator):
         return self
 
     def transform(self, X):
-        # x is a document, word[0] is the word tokenq
+        # x is a document, word[0] is the word token
         return np.vstack([self.model[word[0].lower()] if word[0].lower() in self.model else
                           np.zeros(self.model.layer1_size) for x in X for word in x])
 

### 2014-12-10 paper/paper.tex
updates

@@
 mates. The fundamental nature of the distinction between animate and
-inanimate appears in the early acquisition of the division
+inanimate is reflected in the fact that the division is acquired very early in life.
 \cite{opfer:02}. Children of less than 6 months old are well able to
-make the division. Recent brain research shows that the distinction
+make the division. Moreover, recent brain research shows that the distinction
 appears in the organization of the brain (e.g.~\cite{gao:12}). For some
@@
 world.
-
+%hoe sluit dit aan op de eerdere alinea? lijkt alsof je jezelf tegenspreekt namelijk: als het een standpunt is, waarom zien we dat dan niet in de hersens?%
 A common view supported by philosophers such as Daniel Dennett is that
 animacy and aliveness are better to be treated as epistemological
-stances rather than fixed states in the world \cite{dennett:96}. Not
+stances rather than fixed states in the world \cite{dennett:96}: not
 ineffable qualia but behavioral capacity defines our stance towards
@@
 animate, they utilize different cognitive strategies to explain and
-predict the actions of those objects. The following opening of a
+predict the actions of those objects. Th following opening of a
 well-known story will make this point more clear:
@@
 \cite{orasan:07,lee:13,ovr:niv:07}. Existing methods to automatically
-label natural language for animacy are usually rule based or apply
-machine learning techniques or make a combination of both. Common to
+label natural language for animacy are usually either rule based or apply
+machine learning techniques, or make a combination of both. Common to
 most approaches is that they make use of semantic lexicons with
@@
 types are relatively costly to obtain as they require large
-vocabularies or syntactic parsing systems, which are besides for a
-handful of western languages, often not available.
+vocabularies or syntactic parsing systems, which, with the exception of a few languages, are not readily available.
 
@@
 
-The rest of this paper is structured as follows. We will start with an
-short overview of existing techniques to automatically label texts for
-animacy (\S\ref{sec:previous-work}). After a description of the corpus
+The rest of this paper is structured as follows. We will start with a
+short overview of existing techniques for automatically labeling animacy in texts (\S\ref{sec:previous-work}). After a description of the corpus
 used in our study and how the annotations of the corpus have been
@@
 
-In most of the literature on automatic animacy detection, people make
-a binary distinction between animate and inanimate. \cite{bowman:12}
-for example, define objects to be animate if they are alive and have
+%ik wil eigenlijk meer uitleg in deze paragraaf: 
+In most of the literature on automatic animacy detection, a static, binary distinction is made
+between animate and inanimate. \cite{bowman:12}
+for example, define objects as animate if they are alive and have
 the ability to move under their own will. \cite{orasan:07} define
-animacy in the context of anaphora resolution: ``if its referent can
+animacy in the context of anaphora resolution: something is animate ``if its referent can
 also be referred to using one of the pronouns he, she, him, her, his,
 hers, himself, herself, or a combination of such pronouns
-(e.g. his/her)''. Not only are some parts of these definitions somewhat
-circular, we believe that they are both essentialistic and not
-particularly suited for the rich and wondrous entities that live in
-the realm of stories. We choose to define animacy in terms of
-Dennett's intentional stance which comes down to the question whether
+(e.g. his/her)''. However, some parts of these definitions are somewhat
+circular and essentialistic. More importantly, they are not
+particularly applicable to the rich and wondrous entities that live in
+the realm of stories. As our example above showed, it is possible for an otherwise as inanimate labeled object to attain animacy. We therefore choose to define animacy in terms of
+Dennett's intentional stance, which is more dynamic, and which ultimately comes down to the question whether
 ``you decide to treat the object whose behavior is to be predicted as
@@
 a number of animacy classifiers that make use of syntactic and
-morphological features. Some of the features used are: how often has a
-noun been analyzed as `subject' or `object'? How often does a noun
-occur in a passive by-phrase? How often does a noun occur as a
-subject, immediately followed by animate personal pronouns or
-inanimate personal pronouns?  These features are then aggregated for
+morphological features.
+These features include the frequency of analysis of the noun as `subject' or `object', the frequency of the occurence of a noun in a passive $by$-phrase, and the frequency of the noun as a subject followed by either animate personal pronouns or
+inanimate personal pronouns. These features are then aggregated for
 each lemma after which a machine learning system (decision tree or
 $k$-nearest neighbor classifier) is trained. In \cite{bowman:12} a
-similar approach is presented. In this study a maximum entropy
+similar approach is presented. In this study a Maximum Entropy
 classifier is trained on the basis of three feature types: (1)
@@
 % commentary on above approaches
-There are a number of problems with the approaches mentioned
-above. First, almost all of them heavily rely on costly,
+The approaches mentioned above present us with a number of problems. First, almost all of them heavily rely on costly,
 linguistically informed features derived from lexical-semantic
 databases or syntactic parsing. For most languages in the world,
-however, we cannot rely on these resources. Second, animacy detection
-is often seen as a useful feature for a range of natural language
+however, we cannot rely on these resources, either because they do not exist, or because %syntactic parsers have too low results%. 
+Second, animacy detection is often seen as a useful feature for a range of natural language
 processing techniques, such as anaphora resolution and syntactic
@@
 detection, however, leads to a chicken and egg situation. Another
-problem with the approaches above is that they are lemma based which
+problem with the approaches above is that they are lemma-based, which
 means that the models are generally insensitive to different usages of
 a particular word. A pancake is typically not an animate entity, but,
-as we have seen, whether it is animate or not is only to be determined
-in relation to the story in which it appears. We therefore need to
+as we have seen, whether it is animate or not may depend on the story in which it appears, and even within the story the animacy may change. We therefore need to
 develop a dynamic, data driven system that is token based and does not
@@
 
-Our corpus consists of 74 Dutch stories from the collection
+To develop this dynamic, data driven system we used  a corpus of Dutch folktales. The corpus consists of 74 Dutch stories from the collection
 \textit{Volkssprookjes uit Nederland en Vlaanderen} compiled by
-\cite{sinninghe:78}. The collection contains retellings of popular
-stories from the Netherlands and Flanders, such as \textit{The Bremen
+\cite{sinninghe:78}. The collection is made up of Dutch retellings of popular stories, including such tales as \textit{The Bremen
   Town Musicians} (ATU 130) and \textit{The Table, the Ass, and the
-  Stick } (ATU 563). Some of the less well-know stories are
-\textit{Cock, Hen, Duck, Pin, and Needle on a Journey} (AT 0210) and
-(ATU 780) \textit{The Singing Bone}. A digital version of the
+  Stick } (ATU 563), as well as lesser-known stories such as
+\textit{The Singing Bone} (ATU 780) and \textit{Cock, Hen, Duck, Pin, and Needle on a Journey} (AT 0210).
+This last story is again a clear example of otherwise inanimate objects being animate in a folk story, as is concerns the adventures of several household items, such as a pin, a hackle, an egg, and a whetstone.
+A digital version of the
 collection is available in the Dutch Folktale Database from the
 Meertens Institute (corpus SINVSUNV.20E).\footnote{See
-  \url{http://www.verhalenbank.nl}} We chose this collection because
-of its homogeneity while it still displays enough diversity: (1) all
-tales are written in standard Dutch, (2) the stories are edited by the
-same editor, (3) the tales have a comparable length and (4) the
-collection contains tales from a number of different genres, such as
-fairy tales, legends, riddles etc.
+\url{http://www.verhalenbank.nl}} Using a single collection for our corpus presents us with a helpful homogeneity with regard to the editor, length and language use.
+On the other hand, the collection displays diversity as well in the choice of the stories, which entail fairy tales, legends, and sagas.
+%getallen nakijken%
+%moet je de labels hier al wel noemen? want dat lijkt me eerder een resultaat dan een pre-processing stap toch?%
 
-We assigned to each word in the collection a binary label `animate' or
-`inanimate'. 11542 words were labeled as animate and 62926 words as
-inanimate. The vocabulary size is 5549 with 743 unique animate words
-and 5011 unique inanimate words. The annotations consist of 4627
-animate words that have been tagged as noun or name against 6878
-inanimate words.
+The corpus consists of 74504 words, from 5549 unique words.
+All words or tokens were assigned a binary label `animate' or `inanimate', where all words that were not animate (e.g.
+determiners, adjectives, etc.) were labelled as inanimate.
+This resulted in 11542 animate tokens of 743 types, and 62926 inanimate tokens from 5011 unique inanimate words.
+A selection of words, such as the abovementioned pancake and egg, was labeled as both animate and inanimate, depending on the context.
+The distribution of the nouns is striking: of the animate tokens, 4627 (40,1\%) are nound, while only 6878 of the inanimate tokens (11 \%) are nouns.
+This shows that being a noun is already somewhat of an indication for animacy.
+
+% voorbeeld met gloss 4 regels (NL, ani/inani/Engels/vrij vertaling)
 
@@
 state of the art syntactic parser for Dutch, Alpino \cite{bouma:01}.
+
+% gaat nu de volgorde niet fout? Eerst tokenizeren en parsen, en dan animate/inanimate toekennen?%
 
@@
 We formulate the problem of animacy detection as a classification
-problem where the goal is to assign to each word in a document a label
+problem where the goal is to assign a label at word level, rather than at lemma level,
 indicating whether it represents an animate or an inanimate word.
@@
 Animate words are far outnumbered by inanimate words in our
-collection. Reporting accuracy scores would therefore provide skewed
-results too much in favor of the majority category. The rarity of
+collection (see (\S\ref{sec:data})). Reporting accuracy scores would therefore provide skewed
+results, strongly favoring the majority category. The relative rarity of
 animate words makes evaluation measures such as the well-known $F1$-score
 more appropriate. For all experiments we report on the precision,
-recall and $F1$-score \cite{rijsbergen:79}. In most of the literature on
-animacy detection results are only given for the classification of
-nouns or noun phrases. We will report the results for all words in a
-text and on those specifically for nouns separately.
+recall and $F1$-score \cite{rijsbergen:79}.
+Also, while in most of the literature on animacy detection results are only presented for the classification of nouns or noun phrases, we will present these and also report on the results for all words in a
+text.
 
@@
 such a way that it mimics this situation as closely as possible. If we
-would make a random split of say 80\% of the data for training and
+would, for example, make a random split of 80\% of the data for training and
 20\% for testing on the word level, we run the risk of mixing training
-data with test data making it too easy for a system to rely on words
+data with test data, thereby making it too easy for a system to rely on words
 it has seen from the same text. \cite{bowman:12} fall into this trap
@@
 very likely that in their setup sentences from the same document are
-present in both the training data and the test data which makes their
-evaluation unrealistic. To circumvent this problem, we train our model
-on 80\% of the stories (n=59) and test on the remaining 15 stories.
+present in both the training data and the test data, making their
+evaluation unrealistic. To circumvent this problem, we split at the story level, using 80\% of the stories (n=59) as training material and the remaining 15 stories as test.
+%dat is de innovatie right? dat het niet gaat om woord of zin maar om de hele tekst%
 
@@
 \subsubsection{Morphological Features}
-For each word we extract its Part-of-Speech tag. For reasons of
+For each word we extract its part-of-speech tag. For reasons of
 comparability we choose to use the tags as provided by the output of
 the syntactic parser Alpino, instead of a more specialized
-Part-of-Speech tagger. Again, we take a sliding window approach and
-extract the Part-of-Speech tags for three word left of the focus word,
-three word right of the focus word and the tag of the focus word
-itself.
+part-of-speech tagger. Again, we take a sliding window approach and
+extract the part-of-speech tags for three words left and right of the focus word, as well as the tag of the focus word itself.
 
@@
 proven to perform rather well.
+
+%is n hier weer 3? Je specificeert dat niet%
 
@@
 
+%je moet een of meerdere bronnen noemen hier, by "recently there has been interest"%
+
 We have trained word embeddings with 300 dimensions using the popular
@@
 similarities between animate words can be estimated by inspecting the
-context in which they occur. The hypothesis is that the word
+context in which they occur. From this follows for example that the word
 embeddings of an animate word are more similar to those of other
-animate words than than to those of inanimate words. To give an
+animate words, as opposed to the embeddings of inanimate words. To give an
 example, the 5 nearest neighbors of \textit{Assepoester}
@@
 augment each word in the data with its corresponding 300 dimension
-word embeddings vector. Out-of-vocabulary words were given a vector were all
-dimensions are set to zero.
+word embeddings vector. Out-of-vocabulary words were given a vector were all dimensions are set to zero.
 
@@
 \section{Results}\label{sec:results}
+
+In table \ref{tab:results-all} we present the results for all nine
+models on the complete data set. For each model we report the
+precision, recall and $F1$-score for the animate words and the
+inanimate words.
 
@@
 
-In table \ref{tab:results-all} we present the results for all nine
-models on the complete data set. For each model we report the
-precision, recall and $F1$-score for the animate words and the
-inanimate words. All models perform well on classifying inanimate
+All models perform well on classifying inanimate
 words. However, since this is the majority class, it is far more
@@
 Part-of-Speech, lemmata, etcetera, only has a positive impact on the
-recall of the model leaving the precision untouched. As can be
-observed from the table, the rather expensive dependency features show
-barely any improvement. The model that only uses word embedding
-features, is one of the best performing models. This was slightly
+recall of the model, while leaving the precision untouched. As can be
+observed from the table, employing the rather expensive dependency features show
+barely any improvement. Unexpectedly, the model that only uses word embedding
+features is one of the best performing models. This was slightly
 unexpected, since this is the model in which features are context
@@
 the word $n$-gram features outperforms the other linguistically
-informed models that make use of lemmata, Part-of-Speech tags or
+informed models that make use of lemmata, part-of-speech tags or
 dependency information. The best performance is achieved by the model
-that combines the word features, Part-of-Speech tags and the word
+that combines the word features, part-of-speech tags and the word
 embeddings with an $F1$-score of 0.93 for animate words and 0.99 on
@@
 performance gain.
+
+Table \ref{tab:results-noun} presents the results for all nouns and
+names in the data set. The best performance is again achieved by the
+model that combines the word features with the Part-of-Speech tags and
+word embeddings, resulting in an $F1$-score of 0.92 for animate
+instances and 0.95 for inanimate instances.
 
@@
 
-Table \ref{tab:results-noun} presents the results for all nouns and
-names in the data set. The best performance is again achieved by the
-model that combines the word features with the Part-of-Speech tags and
-word embeddings, resulting in an $F1$-score of 0.92 for animate
-instances and 0.95 for inanimate instances.
-
 \section{Concluding Remarks}
@@
 successful. We compared the performance of linguistically informed
-models (using features such as Part-of-Speech and dependency tags) to
+models (using features such as part-of-speech and dependency tags) to
 models that make use of lower dimensional representations of the

### 2014-12-10 kappa.py

@@ -1,0 +1,64 @@
+import codecs
+import sys
+import glob
+import os
+sys.path.append("/Users/folgert/local/brat/server/src")
+
+from tokenise import gtb_token_boundary_gen
+import numpy as np
+
+from sklearn.metrics import confusion_matrix
+
+
+annotators = folgert, marten = 0, 1
+
+def read_annotation(filename):
+    for line in codecs.open(filename, encoding='utf-8'):
+        _, entity, _ = line.strip().split('\t')
+        _, start, end = entity.split()
+        yield filename.split('/')[-1][1:], int(start), int(end)
+
+def word_positions(filename):
+    with codecs.open(filename) as infile:
+        text = infile.read()
+    return [o for o in gtb_token_boundary_gen(text)]
+
+annotations = {}
+root = "/Users/folgert/Dropbox/stories/data/sinninghe-kappa"
+for filename in glob.glob(os.path.join(root, '*.ann')):
+    filename_str = filename.split('/')[-1]
+    for start, end in word_positions(filename.replace(".ann", ".txt")):
+        found = False
+        annotations[filename_str[1:], start, end] = [0, 0]
+
+for filename in glob.glob(os.path.join(root, '*.ann')):
+    filename_str = filename.split('/')[-1]
+    for annotation, start, end in read_annotation(filename):
+        if (annotation, start, end) in annotations:
+            annotations[annotation, start, end][
+                folgert if filename_str.startswith('F') else marten] = 1
+        else:
+            for (f, s, e) in annotations:
+                if f == annotation and (s >= start and e <= end):
+                    annotations[annotation, s, e][
+                        folgert if filename_str.startswith('F') else marten] = 1
+
+folgert_ann, marten_ann = zip(*annotations.values())
+cm = confusion_matrix(folgert_ann, marten_ann)
+
+def kappa(table):
+    table = np.array(table, dtype=np.float)
+    n_annotations = float(table.sum())
+    p_a = (table[1, 1] + table[0, 0]) / n_annotations
+    p_1 = table[1,:].sum() / n_annotations * table[:,1].sum() / n_annotations
+    p_0 = table[0,:].sum() / n_annotations * table[:,0].sum() / n_annotations
+    p_e = p_1 + p_0
+    return (p_a - p_e) / (1 - p_e)
+
+
+def test():
+    cm = np.array([[20., 5.], [10., 15.]])
+    assert abs(kappa(cm) - 0.4) < 1e-5
+    print 'test passed'
+
+test()

### 2014-12-10 experiment.py
add auc scores

@@ -11,7 +11,7 @@ from sklearn.feature_extraction import DictVectorizer
 from sklearn.linear_model import LogisticRegression, SGDClassifier
 from sklearn.neighbors import KNeighborsClassifier
 from sklearn.svm import LinearSVC
-from sklearn.metrics import classification_report
+from sklearn.metrics import classification_report, roc_auc_score
 from sklearn.preprocessing import LabelEncoder
 
 from gensim.models.word2vec import Word2Vec
@@ -228,6 +228,8 @@ if __name__ == '__main__':
         clf.fit(X_train, y_train)
         preds = clf.predict(X_test)
         print classification_report(y_test, preds)
+        pred_probs = clf.predict_proba(X_test)
+        print 'AUC', roc_auc_score(y_test, pred_probs)
         print "Classification report on nouns:"
         noun_preds = []
         i = 0
@@ -238,9 +240,9 @@ if __name__ == '__main__':
             i += len(X[idx])
         print classification_report(y_test[noun_preds], preds[noun_preds])
 
-        print "Fitting a majority vote DummyClassifier"
-        dummy_clf = DummyClassifier(strategy='constant', constant=1)
-        dummy_clf.fit(X_train, y_train)
-        preds = dummy_clf.predict(X_test)
-        print "Classification report for Dummy Classifier:"
-        print classification_report(y_test, preds)
+        # print "Fitting a majority vote DummyClassifier"
+        # dummy_clf = DummyClassifier(strategy='constant', constant=1)
+        # dummy_clf.fit(X_train, y_train)
+        # preds = dummy_clf.predict(X_test)
+        # print "Classification report for Dummy Classifier:"
+        # print classification_report(y_test, preds)

### 2014-12-10 experiment.py
cross validation support

@@ -3,15 +3,17 @@ import sys
 
 import numpy as np
 import scipy.sparse as sp
+import pandas as pd
+import seaborn as sb
 
 from sklearn.base import BaseEstimator
-from sklearn.cross_validation import train_test_split
-from sklearn.dummy import DummyClassifier
+from sklearn.cross_validation import KFold
 from sklearn.feature_extraction import DictVectorizer
 from sklearn.linear_model import LogisticRegression, SGDClassifier
 from sklearn.neighbors import KNeighborsClassifier
 from sklearn.svm import LinearSVC
-from sklearn.metrics import classification_report, roc_auc_score
+from sklearn.metrics import classification_report, average_precision_score
+from sklearn.metrics import precision_recall_curve, precision_recall_fscore_support
 from sklearn.preprocessing import LabelEncoder
 
 from gensim.models.word2vec import Word2Vec
@@ -180,17 +182,11 @@ if __name__ == '__main__':
     # read the data and extract all features
     X, y = load_data(sys.argv[2], limit=None)
     # split the data into a train and test set (this is based on documents, not words!)
-    X_train_idx, X_test_idx, y_train_idx, y_test_idx = train_test_split(
-        range(len(X)), range(len(X)), test_size=0.2, random_state=1)
-    # get the actual data by flattening the documents
-    X_train_docs = [X[i] for i in X_train_idx]
-    y_train_docs = [label for i in y_train_idx for label in y[i]]
-    X_test_docs = [X[i] for i in X_test_idx]
-    y_test_docs = [label for i in y_test_idx for label in y[i]]
-    # load the desired word2vec model
+    scores = pd.DataFrame(
+        columns=['experiment', 'fold', 'class', 'precision', 'recall', 'Fscore', 'AUC'])
+    noun_scores = pd.DataFrame(
+        columns=['experiment', 'fold', 'class', 'precision', 'recall', 'Fscore', 'AUC'])
     model = Word2Vec.load(sys.argv[1])
-    # model.init_sims(replace=True)
-
     # set up a number of experimental settings
     experiments = [('word',), ('word', 'pos'), ('word', 'pos', 'root'),
                    ('word', 'pos', 'root', 'rel'), tuple(FIELDNAMES)]
@@ -205,40 +201,59 @@ if __name__ == '__main__':
         'knn': KNeighborsClassifier(weights='distance')
     }
 
-    for experiment in experiments:
-        print "Features: %s" % ', '.join(experiment)
[…]

### 2014-12-12 experiment.py
add evaluation of ambiguous words

@@ -186,6 +186,8 @@ if __name__ == '__main__':
         columns=['experiment', 'fold', 'class', 'precision', 'recall', 'Fscore', 'AUC'])
     noun_scores = pd.DataFrame(
         columns=['experiment', 'fold', 'class', 'precision', 'recall', 'Fscore', 'AUC'])
+    ambiguous_scores = pd.DataFrame(
+        columns=['experiment', 'fold', 'class', 'precision', 'recall', 'Fscore', 'AUC'])
     model = Word2Vec.load(sys.argv[1])
     # set up a number of experimental settings
     experiments = [('word',), ('word', 'pos'), ('word', 'pos', 'root'),
@@ -200,6 +202,9 @@ if __name__ == '__main__':
         'svm': LinearSVC(),
         'knn': KNeighborsClassifier(weights='distance')
     }
+
+    ambiguous_words = set(line.strip() for line in codecs.open(
+        'data/ambiguous.txt', encoding='utf-8'))
 
     n_experiments = 0
     for k, (train_index, test_index) in enumerate(KFold(
@@ -253,6 +258,22 @@ if __name__ == '__main__':
             ap = average_precision_score(y_test[noun_preds], pred_probs[noun_preds][:,1])
             noun_scores.loc[n_experiments] = np.array([exp_name, k, 0, p[0], r[0], f[0], ap])
             noun_scores.loc[n_experiments+1] = np.array([exp_name, k, 1, p[1], r[1], f[1], ap])
+
+            ambiguous_preds = []
+            i = 0
+            for idx in test_index:
+                for j, w in enumerate(X[idx]):
+                    if w[0] in ambiguous_words:
+                        ambiguous_preds.append(i + j)
+                i += len(X[idx])
+            print "Classification report on ambigous words:"
+            print classification_report(y_test[ambiguous_preds], preds[ambiguous_preds])
+            p, r, f, _ = precision_recall_fscore_support(
+                y_test[ambiguous_preds], preds[ambiguous_preds])
+            ap = average_precision_score(y_test[ambiguous_preds], pred_probs[ambiguous_preds][:,1])
+            ambiguous_scores.loc[n_experiments] = np.array([exp_name, k, 0, p[0], r[0], f[0], ap])
+            ambiguous_scores.loc[n_experiments+1] = np.array([exp_name, k, 1, p[1], r[1], f[1], ap])
+
             n_experiments += 2
 
         # print "Fitting a majority vote DummyClassifier"

### 2014-12-12 paper/paper.tex
kappa scores and new results

@@
 
-\usepackage{blindtext}
+\usepackage{gb4e}
 \bibliographystyle{plain}
@@
 mates. The fundamental nature of the distinction between animate and
-inanimate is reflected in the fact that the division is acquired very early in life.
-\cite{opfer:02}. Children of less than 6 months old are well able to
-make the division. Moreover, recent brain research shows that the distinction
-appears in the organization of the brain (e.g.~\cite{gao:12}). For some
-researchers this provides evidence for the idea that the division
-between animate and inanimate is an innate part of how we see the
-world.
-%hoe sluit dit aan op de eerdere alinea? lijkt alsof je jezelf tegenspreekt namelijk: als het een standpunt is, waarom zien we dat dan niet in de hersens?%
+inanimate is reflected in the fact that the division is acquired very
+early in life~\cite{opfer:02}. Children of less than 6 months old
+are well able to make the division. Moreover, recent brain research
+shows that the distinction appears in the organization of the brain
+(e.g.~\cite{gao:12}). For some researchers this provides evidence for
+the idea that the division between animate and inanimate is an innate
+part of how we see the world.
+
 A common view supported by philosophers such as Daniel Dennett is that
@@
 \cite{orasan:07,lee:13,ovr:niv:07}. Existing methods to automatically
-label natural language for animacy are usually either rule based or apply
-machine learning techniques, or make a combination of both. Common to
-most approaches is that they make use of semantic lexicons with
-information about animacy and syntactic cues in a text. Both feature
-types are relatively costly to obtain as they require large
-vocabularies or syntactic parsing systems, which, with the exception of a few languages, are not readily available.
+label natural language for animacy are usually either rule based or
+apply machine learning techniques, or make a combination of
+both. Common to most approaches is that they make use of semantic
+lexicons with information about animacy and syntactic cues in a
+text. Both feature types are relatively costly to obtain as they
+require large vocabularies or syntactic parsing systems, which, with
+the exception of a few languages, are not readily available.
 
@@
 The rest of this paper is structured as follows. We will start with a
-short overview of existing techniques for automatically labeling animacy in texts (\S\ref{sec:previous-work}). After a description of the corpus
-used in our study and how the annotations of the corpus have been
-established (\S\ref{sec:data}), we will give an account of our
-computational models in Section \ref{sec:models}. We report on the
+short overview of existing techniques for automatically labeling
+animacy in texts (\S\ref{sec:previous-work}). After a description of
+the corpus used in our study and how the annotations of the corpus
+have been established (\S\ref{sec:data}), we will give an account of
+our computational models in Section \ref{sec:models}. We report on the
 empirical results in Section \ref{sec:results}. The last section
@@
 %ik wil eigenlijk meer uitleg in deze paragraaf:
-In most of the literature on automatic animacy detection, a static, binary distinction is made
-between animate and inanimate. \cite{bowman:12}
-for example, define objects as animate if they are alive and have
-the ability to move under their own will. \cite{orasan:07} define
-animacy in the context of anaphora resolution: something is animate ``if its referent can
-also be referred to using one of the pronouns he, she, him, her, his,
-hers, himself, herself, or a combination of such pronouns
-(e.g. his/her)''. However, some parts of these definitions are somewhat
-circular and essentialistic. More importantly, they are not
+In most of the literature on automatic animacy detection, a static,
+binary distinction is made between animate and
+inanimate. \cite{bowman:12} for example, define objects as animate if
+they are alive and have the ability to move under their own
+will. \cite{orasan:07} define animacy in the context of anaphora
+resolution: something is animate ``if its referent can also be
+referred to using one of the pronouns he, she, him, her, his, hers,
+himself, herself, or a combination of such pronouns
+(e.g. his/her)''. However, some parts of these definitions are
+somewhat circular and essentialistic. More importantly, they are not
 particularly applicable to the rich and wondrous entities that live in
-the realm of stories. As our example above showed, it is possible for an otherwise as inanimate labeled object to attain animacy. We therefore choose to define animacy in terms of
-Dennett's intentional stance, which is more dynamic, and which ultimately comes down to the question whether
-``you decide to treat the object whose behavior is to be predicted as
-a rational agent'' \cite[pp. 17]{dennett:96}.
+the realm of stories. As our example above showed, it is possible for
+an otherwise as inanimate labeled object to attain animacy. We
+therefore choose to define animacy in terms of Dennett's intentional
+stance, which is more dynamic, and which ultimately comes down to the
+question whether ``you decide to treat the object whose behavior is to
+be predicted as a rational agent'' \cite[pp. 17]{dennett:96}.
 
@@
 a number of animacy classifiers that make use of syntactic and
-morphological features.
-These features include the frequency of analysis of the noun as `subject' or `object', the frequency of the occurence of a noun in a passive $by$-phrase, and the frequency of the noun as a subject followed by either animate personal pronouns or
+morphological features. These features include the frequency of
+analysis of the noun as `subject' or `object', the frequency of the
+occurence of a noun in a passive $by$-phrase, and the frequency of the
+noun as a subject followed by either animate personal pronouns or
 inanimate personal pronouns. These features are then aggregated for
@@
 
-\cite{moore:13} present an interesting approach that combines a number
-of animacy classifiers in a voting scheme and aims at an interpretable
-and correctable model of animacy classification. They combine a
-variety of classifiers such as the WordNet-based approach of
-\cite{evans:00}, dictionary sources and systems for named entity
-recognition.
+\cite{moore:13} present an approach that combines a number of animacy
+classifiers in a voting scheme and aims at an interpretable and
+correctable model of animacy classification. They combine a variety of
+classifiers such as the WordNet-based approach of \cite{evans:00},
+dictionary sources and systems for named entity recognition.
 
 % commentary on above approaches
-The approaches mentioned above present us with a number of problems. First, almost all of them heavily rely on costly,
+The approaches mentioned above present us with a number of
+problems. First, almost all of them heavily rely on costly,
 linguistically informed features derived from lexical-semantic
 databases or syntactic parsing. For most languages in the world,
-however, we cannot rely on these resources, either because they do not exist, or because %syntactic parsers have too low results%. 
-Second, animacy detection is often seen as a useful feature for a range of natural language
-processing techniques, such as anaphora resolution and syntactic
-parsing. The mutual dependence of these techniques and animacy
-detection, however, leads to a chicken and egg situation. Another
-problem with the approaches above is that they are lemma-based, which
-means that the models are generally insensitive to different usages of
-a particular word. A pancake is typically not an animate entity, but,
-as we have seen, whether it is animate or not may depend on the story in which it appears, and even within the story the animacy may change. We therefore need to
-develop a dynamic, data driven system that is token based and does not
-rely too much on static and fixed world knowledge.
+however, we cannot rely on these resources, either because they do not
+exist, or because their performance is insufficient.  Second, animacy
+detection is often seen as a useful feature for a range of natural
+language processing techniques, such as anaphora resolution and
+syntactic parsing. The mutual dependence of these techniques and
+animacy detection, however, leads to a chicken and egg
+situation. Another problem with the approaches above is that they are
+lemma-based, which means that the models are generally insensitive to
+different usages of a particular word. A pancake is typically not an
+animate entity, but, as we have seen, whether it is animate or not may
+depend on the story in which it appears, and even within the story the
+animacy may change. We therefore need to develop a dynamic, data
+driven system that is token based and does not rely too much on static
+and fixed world knowledge.
 
@@
 
-To develop this dynamic, data driven system we used  a corpus of Dutch folktales. The corpus consists of 74 Dutch stories from the collection
+To develop this dynamic, data driven system we used a corpus of Dutch
+folktales. The corpus consists of 74 Dutch stories from the collection
 \textit{Volkssprookjes uit Nederland en Vlaanderen} compiled by
-\cite{sinninghe:78}. The collection is made up of Dutch retellings of popular stories, including such tales as \textit{The Bremen
-  Town Musicians} (ATU 130) and \textit{The Table, the Ass, and the
-  Stick } (ATU 563), as well as lesser-known stories such as
-\textit{The Singing Bone} (ATU 780) and \textit{Cock, Hen, Duck, Pin, and Needle on a Journey} (AT 0210).
-This last story is again a clear example of otherwise inanimate objects being animate in a folk story, as is concerns the adventures of several household items, such as a pin, a hackle, an egg, and a whetstone.
-A digital version of the
-collection is available in the Dutch Folktale Database from the
-Meertens Institute (corpus SINVSUNV.20E).\footnote{See
-\url{http://www.verhalenbank.nl}} Using a single collection for our corpus presents us with a helpful homogeneity with regard to the editor, length and language use.
-On the other hand, the collection displays diversity as well in the choice of the stories, which entail fairy tales, legends, and sagas.
+\cite{sinninghe:78}. The collection is made up of Dutch retellings of
+popular stories, including such tales as \textit{The Bremen Town
+  Musicians} (ATU 130) and \textit{The Table, the Ass, and the Stick }
+(ATU 563), as well as lesser-known stories such as \textit{The Singing
+  Bone} (ATU 780) and \textit{Cock, Hen, Duck, Pin, and Needle on a
+  Journey} (ATU 210). This last story is again a clear example of
+otherwise inanimate objects being animate in a folk story, as it
+concerns the adventures of several household items, such as a
+\textit{pin}, a \textit{hackle}, an \textit{egg}, and a
+\textit{whetstone}. A digital version of the collection is available
+in the Dutch Folktale Database from the Meertens Institute (corpus
+SINVSUNV.20E).\footnote{See \url{http://www.verhalenbank.nl}} Using a
+single collection for our corpus presents us with a helpful
+homogeneity with regard to the editor, length and language use. On the
+other hand, the collection displays diversity as well in the choice of
+the stories, which entail fairy tales, legends, and sagas.
 %getallen nakijken%
@@
 
-The corpus consists of 74504 words, from 5549 unique words.
-All words or tokens were assigned a binary label `animate' or `inanimate', where all words that were not animate (e.g.
-determiners, adjectives, etc.) were labelled as inanimate.
-This resulted in 11542 animate tokens of 743 types, and 62926 inanimate tokens from 5011 unique inanimate words.
-A selection of words, such as the abovementioned pancake and egg, was labeled as both animate and inanimate, depending on the context.
-The distribution of the nouns is striking: of the animate tokens, 4627 (40,1\%) are nound, while only 6878 of the inanimate tokens (11 \%) are nouns.
-This shows that being a noun is already somewhat of an indication for animacy.
-
-% voorbeeld met gloss 4 regels (NL, ani/inani/Engels/vrij vertaling)
+The data were labeled by two annotators. We computed an
+inter-annotator agreement score (Cohen's Kappa) of $K=0.95$ on the
+basis of five stories that were annotated by both annotators. The
+following sentence provides an example annotation. All unlabeled words
+are implicitly considered to be inanimate.
+\begin{exe}
+\ex
+\gll Jij smid, jij bent de sterkste; hou je vast aan de bovenste
+takken, en dan ga jij kleermaker aan zijn benen hangen en zo gaan we maar door\\
+    \textsc{anim} \textsc{anim} \textsc{anim} {} {} {} {} \textsc{anim} {} {} {} {} {} {} {} {} \textsc{anim} \textsc{anim}
+    {} \textsc{anim} {} {} {} {} {} \textsc{anim} {} {} \\
+\trans `You, blacksmith, you are the strongest; hold on to the upper
+branches and then, you, tailor, will grab his legs and so we go on\ldots'
+\end{exe}
+The corpus consists of 74504 words, from 5549 unique words. All word
+tokens were assigned a binary label `animate' or `inanimate', where
+all words that were not animate (e.g. determiners, adjectives, etc.)
+were labelled as inanimate. This resulted in 11542 animate tokens of
+743 types, and 62926 inanimate tokens from 5011 unique inanimate
+words. A selection of words, such as the abovementioned
+\textit{pancake} and \textit{egg}, was labeled as both animate and
+inanimate, depending on the context. The distribution of the nouns is
+telling: of the animate tokens, 4627 (40,1\%) are nound, while only
+6878 of the inanimate tokens (11 \%) are nouns. This shows that being
+a noun is already somewhat of an indication for animacy.
 
@@
 We formulate the problem of animacy detection as a classification
-problem where the goal is to assign a label at word level, rather than at lemma level,
-indicating whether it represents an animate or an inanimate word.
+problem where the goal is to assign a label at word level, rather than
+at lemma level, indicating whether it represents an animate or an
+inanimate word.
 
@@
 
-Animate words are far outnumbered by inanimate words in our
-collection (see (\S\ref{sec:data})). Reporting accuracy scores would therefore provide skewed
-results, strongly favoring the majority category. The relative rarity of
-animate words makes evaluation measures such as the well-known $F1$-score
-more appropriate. For all experiments we report on the precision,
-recall and $F1$-score \cite{rijsbergen:79}.
-Also, while in most of the literature on animacy detection results are only presented for the classification of nouns or noun phrases, we will present these and also report on the results for all words in a
-text.
+Animate words are far outnumbered by inanimate words in our collection
+(see (\S\ref{sec:data})). Reporting accuracy scores would therefore
+provide skewed results, strongly favoring the majority category. The
+relative rarity of animate words makes evaluation measures such as the
+well-known $F1$-score more appropriate. For all experiments we report
+on the precision, recall and $F1$-score \cite{rijsbergen:79}. Also,
+while in most of the literature on animacy detection results are only
+presented for the classification of nouns or noun phrases, we will
+present these and also report on the results for all words in a text.
 
@@
 such a way that it mimics this situation as closely as possible. If we
-would, for example, make a random split of 80\% of the data for training and
-20\% for testing on the word level, we run the risk of mixing training
-data with test data, thereby making it too easy for a system to rely on words
-it has seen from the same text. \cite{bowman:12} fall into this trap
-by making a random split in their data on the sentence level. It is
-very likely that in their setup sentences from the same document are
-present in both the training data and the test data, making their
-evaluation unrealistic. To circumvent this problem, we split at the story level, using 80\% of the stories (n=59) as training material and the remaining 15 stories as test.
-%dat is de innovatie right? dat het niet gaat om woord of zin maar om de hele tekst%
+would, for example, make a random split of 80\% of the data for
+training and 20\% for testing on the word level, we run the risk of
+mixing training data with test data, thereby making it too easy for a
+system to rely on words it has seen from the same
+text. \cite{bowman:12} fall into this trap by making a random split in
+their data on the sentence level. It is very likely that in their
+setup sentences from the same document are present in both the
+training data and the test data, making their evaluation
+unrealistic. To circumvent this problem, we split the data at the
+story level. We make use of 10-fold cross-validation, where we
+randomly select 10\% of the stories for testing and train on the
+remaining 90\%.
 
@@
 models on the complete data set. For each model we report the
-precision, recall and $F1$-score for the animate words and the
+precision, recall, $F1$-score and Area Under the Precision-Recall
+Curve (AUC) for the animate words and the
 inanimate words.
@@
 \toprule
-features                                                 &           & precision & recall & $F1$ & avg $F1$              \\ \midrule
-\multirow{2}{*}{word}                                    & animate   & 0.93      & 0.80   & 0.86 & \multirow{2}{*}{0.96} \\
-                                                         & inanimate & 0.96      & 0.99   & 0.98 &                       \\
-\multirow{2}{*}{word + PoS}                              & animate   & 0.92      & 0.88   & 0.90 & \multirow{2}{*}{0.97} \\
-                                                         & inanimate & 0.98      & 0.99   & 0.98 &                       \\
-\multirow{2}{*}{word + PoS + lemma}                      & animate   & 0.93      & 0.89   & 0.91 & \multirow{2}{*}{0.97} \\
-                                                         & inanimate & 0.98      & 0.99   & 0.98 &                       \\
-\multirow{2}{*}{word + PoS + lemma + dep}                & animate   & 0.93      & 0.89   & 0.91 & \multirow{2}{*}{0.97} \\
-                                                         & inanimate & 0.98      & 0.99   & 0.98 &                       \\
-\multirow{2}{*}{word + embeddings}                       & animate   & 0.92      & 0.93   & 0.92 & \multirow{2}{*}{0.98} \\
-                                                         & inanimate & 0.99      & 0.99   & 0.99 &                       \\
-\rowcolor{Gray}                                          & animate   & 0.92      & 0.93   & 0.93 &                       \\
-\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings} & inanimate & 0.99      & 0.99   & 0.99 & \multirow{-2}{*}{0.98}\\
-\multirow{2}{*}{word + PoS + lemma + embeddings}         & animate   & 0.93      & 0.93   & 0.93 & \multirow{2}{*}{0.98} \\
-                                                         & inanimate & 0.99      & 0.99   & 0.99 &                       \\
-\multirow{2}{*}{word + PoS + lemma + dep + embeddings}   & animate   & 0.92      & 0.93   & 0.93 & \multirow{2}{*}{0.98} \\
-                                                         & inanimate & 0.99      & 0.99   & 0.99 &                       \\
-\multirow{2}{*}{word embeddings}                         & animate   & 0.90      & 0.92   & 0.91 & \multirow{2}{*}{0.97} \\
-                                                         & inanimate & 0.99      & 0.98   & 0.98 &                       \\
-
+features & class & precision &  recall & $F1$ &  AUC \\
+\midrule
+\multirow{2}{*}{embeddings}                            & inanimate     &       0.98 &    0.99 &    0.98 & \multirow{2}{*}{0.93} \\
+                                                       & animate     &       0.92 &    0.89 &    0.91 &  \\
+\multirow{2}{*}{word}                                  & inanimate     &       0.96 &    0.99 &    0.98 & \multirow{2}{*}{0.94} \\
+                                                       & animate     &       0.94 &    0.78 &    0.85 &  \\
+\multirow{2}{*}{word + embeddings}                     & inanimate     &       0.98 &    0.99 &    0.98 & \multirow{2}{*}{0.96} \\
+                                                       & animate     &       0.93 &    0.90 &    0.91 &  \\
+\multirow{2}{*}{word + PoS}                            & inanimate     &       0.97 &    0.99 &    0.98 & \multirow{2}{*}{0.96} \\
+                                                       & animate     &       0.94 &    0.86 &    0.90 &  \\
+\rowcolor{Gray}                                        & inanimate     &       0.98 &    0.99 &    0.99 &  \\
+\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings}   & animate     &       0.94 &    0.92 &    0.93 & \multirow{-2}{*}{0.97} \\
+\multirow{2}{*}{word + PoS + lemma}                    & inanimate    &       0.98 &    0.99 &    0.98 & \multirow{2}{*}{0.96} \\
+                                                       & animate     &       0.94 &    0.87 &    0.90 &  \\
+\multirow{2}{*}{word + PoS + lemma + embeddings}       & inanimate     &       0.98 &    0.99 &    0.99 & \multirow{2}{*}{0.97} \\
+                                                       & animate     &       0.94 &    0.92 &    0.93 &  \\
+\multirow{2}{*}{word + PoS + lemma + dep}              & inanimate     &       0.98 &    0.99 &    0.98 & \multirow{2}{*}{0.96} \\
+                                                       & animate     &       0.94 &    0.88 &    0.91 &  \\
+\multirow{2}{*}{word + PoS + lemma + dep + embeddings} & inanimate     &       0.98 &    0.99 &    0.99 & \multirow{2}{*}{0.97} \\
+                                                       & animate     &       0.94 &    0.92 &    0.93 &  \\
 \bottomrule
 \end{tabular}
-\caption{Precision, recall and $F1$ scores for animate and inanimate classes per feature settings for all words.}
+\caption{Precision, recall, $F1$ and Area Under the Curve (AUC) scores for animate and inanimate classes per feature setting for all words.}
 \label{tab:results-all}
@@
 
-All models perform well on classifying inanimate
-words. However, since this is the majority class, it is far more
-interesting to compare the performance of the models on the animate
-instances. It is interesting to observe that the `simple' $n$-gram
-word model already performs rather well. Adding more features, like
-Part-of-Speech, lemmata, etcetera, only has a positive impact on the
-recall of the model, while leaving the precision untouched. As can be
-observed from the table, employing the rather expensive dependency features show
-barely any improvement. Unexpectedly, the model that only uses word embedding
-features is one of the best performing models. This was slightly
-unexpected, since this is the model in which features are context
-insensitive. However, the model that combines the word embeddings with
-the word $n$-gram features outperforms the other linguistically
-informed models that make use of lemmata, part-of-speech tags or
-dependency information. The best performance is achieved by the model
-that combines the word features, part-of-speech tags and the word
-embeddings with an $F1$-score of 0.93 for animate words and 0.99 on
-inanimate words. Adding more features does not result in any more
-performance gain.
+All models perform well on classifying inanimate words. However, since
+this is the majority class, it is of course far more interesting to
+compare the performance of the models on the animate instances. It is
+interesting to observe that the `simple' $n$-gram word model already
+performs rather well. Adding more features, like Part-of-Speech,
+lemmata, etcetera, only has a positive impact on the recall of the
+model, while leaving the precision untouched. As can be observed from
+the table, employing the rather expensive dependency features shows
+barely any improvement.
+
+The model that only uses word embedding features is one of the best
+performing models. This is a context insensitive model that operates
+on the level of the vocabulary, which means that it will predict the
+same outcome for each token of a particular word type. The high
+precision and high recall show us that this model has acquired
+knowledge about which words \textit{typically} group with animate
+words and which with inanimate words. However, the models that combines
+the word embeddings with the context sensitive features, such as word
+$n$-grams or Part-of-Speech tags, outperform the context insensitive
+model.
+
+The best performance is
+achieved by the model that combines the word features, part-of-speech
+tags and the word embeddings with an $F1$-score of 0.93 for animate
+words and 0.99 on inanimate words. Adding more features does not
+result in any more performance gain.
 
@@
 word embeddings, resulting in an $F1$-score of 0.92 for animate
-instances and 0.95 for inanimate instances.
+instances and 0.94 for inanimate instances.
 
@@
 \toprule
-features                                               &           & precision & recall & $F1$ & avg $F1$              \\ \midrule
-\multirow{2}{*}{word}                                  & animate   & 0.96      & 0.64   & 0.77 & \multirow{2}{*}{0.84} \\
-                                                       & inanimate & 0.81      & 0.98   & 0.89 &                       \\
-\multirow{2}{*}{word + PoS}                            & animate   & 0.92      & 0.80   & 0.86 & \multirow{2}{*}{0.90} \\
-                                                       & inanimate & 0.89      & 0.95   & 0.92 &                       \\
-\multirow{2}{*}{word + PoS + lemma}                    & animate   & 0.93      & 0.82   & 0.87 & \multirow{2}{*}{0.90} \\
-                                                       & inanimate & 0.89      & 0.96   & 0.92 &                       \\
-\multirow{2}{*}{word + PoS + lemma + dep}              & animate   & 0.93      & 0.82   & 0.87 & \multirow{2}{*}{0.91} \\
-                                                       & inanimate & 0.90      & 0.96   & 0.93 &                       \\
-\multirow{2}{*}{word + embeddings}                     & animate   & 0.92      & 0.91   & 0.92 & \multirow{2}{*}{0.94} \\
-                                                       & inanimate & 0.95      & 0.95   & 0.95 &                       \\
-\rowcolor{Gray}                                        & animate   & 0.92      & 0.92   & 0.92 &                       \\
-\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings} & inanimate & 0.95    & 0.95   & 0.95 & \multirow{-2}{*}{0.94}\\
-\multirow{2}{*}{word + PoS + lemma + embeddings}       & animate   & 0.92      & 0.91   & 0.92 & \multirow{2}{*}{0.94} \\
-                                                       & inanimate & 0.95      & 0.95   & 0.95 &                       \\
-\multirow{2}{*}{word + PoS + lemma + dep + embeddings} & animate   & 0.92      & 0.91   & 0.92 & \multirow{2}{*}{0.94} \\
-                                                       & inanimate & 0.94      & 0.95   & 0.95 &                       \\
-\multirow{2}{*}{word embeddings}                       & animate   & 0.88      & 0.90   & 0.89 & \multirow{2}{*}{0.92} \\
-                                                       & inanimate & 0.94      & 0.92   & 0.93 &                       \\
-
+features & class &  precision &  recall & $F1$ &  AUC \\
+\midrule
+\multirow{2}{*}{embeddings}                            & inanimate     &       0.89 &    0.94 &    0.92 & \multirow{2}{*}{0.92} \\
+                                                       & animate       &       0.91 &    0.85 &    0.88 &  \\
+\multirow{2}{*}{word}                                  & inanimate     &       0.78 &    0.98 &    0.87 & \multirow{2}{*}{0.91} \\
+                                                       & animate       &       0.96 &    0.60 &    0.74 &  \\
+\multirow{2}{*}{word + embeddings}                     & inanimate     &       0.90 &    0.96 &    0.93 & \multirow{2}{*}{0.96} \\
+                                                       & animate       &       0.94 &    0.86 &    0.89 &  \\
+\multirow{2}{*}{word + pos}                            & inanimate     &       0.86 &    0.96 &    0.90 & \multirow{2}{*}{0.94} \\
+                                                       & animate       &       0.92 &    0.78 &    0.84 &  \\
+\rowcolor{Gray}                                        & inanimate     &       0.93 &    0.96 &    0.94 &  \\
+\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings} & animate       &       0.94 &    0.89 &    0.92 &\multirow{-2}{*}{0.97}  \\
+\multirow{2}{*}{word + PoS + lemma}                    & inanimate     &       0.87 &    0.96 &    0.91 & \multirow{2}{*}{0.95} \\
+                                                       & animate       &       0.94 &    0.80 &    0.86 &  \\
+\multirow{2}{*}{word + PoS + lemma + embeddings}       & inanimate     &       0.93 &    0.96 &    0.94 & \multirow{2}{*}{0.97} \\
+                                                       & animate       &       0.94 &    0.89 &    0.92 &  \\
+\multirow{2}{*}{word + PoS + lemma + dep}              & inanimate     &       0.87 &    0.96 &    0.91 & \multirow{2}{*}{0.95} \\
+                                                       & animate       &       0.93 &    0.80 &    0.86 &  \\
+\multirow{2}{*}{word + PoS + lemma + dep + embeddings} & inanimate     &       0.93 &    0.96 &    0.94 & \multirow{2}{*}{0.97} \\
+                                                       & animate       &       0.94 &    0.90 &    0.92 &  \\
 \bottomrule
 \end{tabular}
-\caption{Precision, recall and $F1$ scores for animate and inanimate classes
+\caption{Precision, recall, $F1$ and Area Under the Curve (AUC) scores for animate and inanimate classes
   per feature settings for all words tagged as noun.}

### 2014-12-16 animacy.py
updates and frog data

@@ -1,0 +1,53 @@
+import codecs
+import os
+import sys
+import glob
+from collections import Counter
+from sklearn.linear_model import LogisticRegression
+from sklearn.preprocessing import LabelEncoder
+from gensim.models.word2vec import Word2Vec
+from experiment import Windower, WordEmbeddings, FeatureStacker
+
+from frog import Frog
+
+def load_data(filename, limit=None):
+    X, y = [[]], [[]]
+    with codecs.open(filename, encoding="utf-8") as infile:
+        for i, line in enumerate(infile):
+            if limit is not None and i >= limit:
+                break
+            if line.startswith("<FB/>"):
+                X.append([])
+                y.append([])
+            else:
+                fields = line.strip().split('\t')
+                X[-1].append([field if field else None for field in fields[:-1]])
+                assert X[-1]
+                y[-1].append('yes' if fields[-1] == 'animate' else 'no')
+    return X, y
+
+X, y = load_data(sys.argv[1], limit=None)
+model = Word2Vec.load_word2vec_format(sys.argv[2], binary=True)
+feature_vectorizer = FeatureStacker(('windower', Windower(window_size=3)),
+                                    ('embeddings', WordEmbeddings(model)))
+
+X = feature_vectorizer.fit_transform([[word for word in doc] for doc in X])
+y = LabelEncoder().fit_transform([l for labels in y for l in labels])
+
+clf = LogisticRegression().fit(X, y)
+frogger = Frog(int(sys.argv[3]))
+
+for filename in glob.glob(os.path.join(sys.argv[4], "*")):
+    print filename
+    characters = Counter()
+    with codecs.open(filename, encoding='utf-8') as infile:
+        doc = infile.read()
+        document = frogger.tag(doc)
+        document = [[f.decode('utf-8') for f in w[:-1]]
+                    for sentence in document for w in sentence]
+    X_test = feature_vectorizer.transform([document])
+    preds = clf.predict(X_test)
+    for i, pred in enumerate(preds):
+        if pred == 1 and document[i][2] in ('N', 'SPEC'):
+            characters[document[i][0]] += 1
+    print ', '.join(sorted(characters, key=characters.__getitem__, reverse=True))

### 2014-12-16 experiment.py

@@ -4,7 +4,6 @@ import sys
 import numpy as np
 import scipy.sparse as sp
 import pandas as pd
-import seaborn as sb
 
 from sklearn.base import BaseEstimator
 from sklearn.cross_validation import KFold
@@ -13,7 +12,7 @@ from sklearn.linear_model import LogisticRegression, SGDClassifier
 from sklearn.neighbors import KNeighborsClassifier
 from sklearn.svm import LinearSVC
 from sklearn.metrics import classification_report, average_precision_score
-from sklearn.metrics import precision_recall_curve, precision_recall_fscore_support
+from sklearn.metrics import precision_recall_fscore_support
 from sklearn.preprocessing import LabelEncoder
 
 from gensim.models.word2vec import Word2Vec
@@ -181,7 +180,6 @@ if __name__ == '__main__':
 
     # read the data and extract all features
     X, y = load_data(sys.argv[2], limit=None)
-    # split the data into a train and test set (this is based on documents, not words!)
     scores = pd.DataFrame(
         columns=['experiment', 'fold', 'class', 'precision', 'recall', 'Fscore', 'AUC'])
     noun_scores = pd.DataFrame(

### 2014-12-16 frog.py

@@ -1,0 +1,43 @@
+import re
+import socket
+
+from collections import namedtuple
+Word = namedtuple("Word", ["word", "lemma", "pos", "animate"])
+pos_re = re.compile('([A-Z]+)\((.*?)\)')
+
+
+class Frog(object):
+    def __init__(self, port):
+        self.BUFSIZE = 4096
+        self.port = port
+
+    def tag(self, text):
+        self.socket = socket.socket(socket.AF_INET,socket.SOCK_STREAM)
+        self.socket.settimeout(120.0)
+        self.socket.connect(('localhost', self.port))
+
+        text = text.strip(' \t\n')
+        text = text.encode('utf-8') + b'\r\nEOT\r\n'
+        self.socket.sendall(text)
+        done = False
+        output = []
+        while not done:
+            data = b""
+            while not data or data[-1] != b'\n':
+                more = self.socket.recv(self.BUFSIZE)
+                if not more: break
+                data += more
+            for line in data.strip(' \t\r\n').split('\n'):
+                line = line.strip()
+                if line == 'READY':
+                    done = True
+                    break
+                elif line:
+                    index, word, lemma, _, pos, _ = line.split('\t')
+                    if index == '1':
+                        output.append([])
+                    pos = pos_re.search(pos)
+                    head = pos.group(1)
+                    output[-1].append(Word(word, lemma, head, 'inanimate'))
+        self.socket.close()
+        return output

### 2014-12-16 frogdataset.py

@@ -1,0 +1,48 @@
+from storypy import Story
+import glob
+import codecs
+import os
+import sys
+
+from frog import Frog, Word
+
+frogger = Frog(int(sys.argv[1]))
+
+def token_boundaries(tokens, original_text):
+    curr_pos = 0
+    for tok in tokens:
+        start_pos = original_text.index(tok, curr_pos)
+        # TODO: Check if we fail to find the token!
+        end_pos = start_pos + len(tok)
+        yield (start_pos, end_pos)
+        curr_pos = end_pos
+
+for filename in glob.glob(os.path.join(sys.argv[2], "SINVS*.ann")):
+    if 'anomalies' in filename:
+        continue
+    story = Story.load(filename)
+    if sys.argv[3] == 'chars':
+        characters = {(start, end): (id, name)
+                      for character in story.characters
+                      for id, name, start, end in character.chain}
+    else:
+        characters = {(start, end): (id, name)
+                      for location in story.locations
+                      for id, name, start, end in location.chain}
+
+    with codecs.open(filename.replace(".ann", ".txt"), encoding='utf-8') as f:
+        orig_text = f.read()
+        tokens = [word for sent in frogger.tag(orig_text) for word in sent]
+        offsets = list(token_boundaries([t[0].decode('utf-8') for t in tokens], orig_text))
+        found_characters = {(start, end): False for start, end in characters}
+        for char_start, char_end in characters:
+            start_found = False
+            for i, (start, end) in enumerate(offsets):
+                if (start == char_start or
+                    (start < char_start < end) or
+                    (char_start < start < char_end)):
+                    start_found = True
+                    tokens[i] = Word(*(tokens[i][:-1] + ('animate',)))
+        for token in tokens:
+            print '\t'.join(token)
+        print '<FB/>'

### 2014-12-17 paper/paper.tex
update paper

@@
 A common view supported by philosophers such as Daniel Dennett is that
-animacy and aliveness are better to be treated as epistemological
+animacy and aliveness are to be treated as epistemological
 stances rather than fixed states in the world \cite{dennett:96}: not
@@
 e.g.\ anaphora resolution and dependency parsing
-\cite{orasan:07,lee:13,ovr:niv:07}. Existing methods to automatically
-label natural language for animacy are usually either rule based or
-apply machine learning techniques, or make a combination of
-both. Common to most approaches is that they make use of semantic
-lexicons with information about animacy and syntactic cues in a
-text. Both feature types are relatively costly to obtain as they
-require large vocabularies or syntactic parsing systems, which, with
-the exception of a few languages, are not readily available.
+\cite{orasan:07,lee:13,ovr:niv:07}. Existing methods for the automatic
+labeling of text for animacy are usually either rule based or apply
+machine learning techniques, or make a combination of both. Common to
+most approaches is that they make use of semantic lexicons with
+information about animacy and syntactic cues in a text. Both feature
+types are relatively costly to obtain as they require large
+vocabularies or syntactic parsing systems, which, with the exception
+of a few languages, are not readily available.
 
 In this paper we present a new linguistically uninformed model to
-automatically label texts for animacy. We will show that we can (mostly)
+automatically label texts for animacy. We show that we can (mostly)
 do away with features that require syntactic parsing or semantic
-lexicons while still yielding competitive performance. We will focus
+lexicons while still yielding competitive performance. We focus
 on labeling animacy in stories because stories pose some interesting
@@
 stories, though inanimate in the `real' world, exhibit all behavioral
-capacities to consider them to be animate and rational agents. Such
+capacities to consider them animate and rational agents. Such
 examples make a clear case for developing dynamic, data driven systems
@@
 the only study that makes use of a fully labeled corpus for animacy.
-Not so much concerned with animacy detection, \cite{karsdorp:12}
+Partially related to animacy detection, \cite{karsdorp:12}
 attempt to extract the cast (i.e.\ all characters) from a
@@
 language processing techniques, such as anaphora resolution and
-syntactic parsing. The mutual dependence of these techniques and
+syntactic parsing. The mutual dependence between these techniques and
 animacy detection, however, leads to a chicken and egg
@@
 contexts in which words appear, is an implementation of this very
-idea. The idea is very simple, and equally strong: words that often
+idea. The idea is very simple, yet powerful: words that often
 co-appear with the same set of words, will have a more similar
 meaning. Recently, there has been a lot of interest in procedures that
-can automatically induce so-called `word-embeddings' from large,
+can automatically induce so-called `word embeddings' from large,
 unannotated collections of texts. These models typically attempt to
-learn a lower-dimensional vector for each word in the vocabulary which
+learn a lower dimensional vector for each word in the vocabulary which
 captures the typical co-occurrence patterns of the word in the
@@
 augment each word in the data with its corresponding 300 dimension
-word embeddings vector. Out-of-vocabulary words were given a vector were all dimensions are set to zero.
+word embeddings vector. Out-of-vocabulary words were given a vector in
+which all dimensions are set to zero.
 
@@
 
-In table \ref{tab:results-all} we present the results for all nine
+In Table \ref{tab:results-all} we present the results for all nine
 models on the complete data set. For each model we report the
@@
 \multirow{2}{*}{embeddings}                            & inanimate     &       0.98 &    0.99 &    0.98 & \multirow{2}{*}{0.93} \\
-                                                       & animate     &       0.92 &    0.89 &    0.91 &  \\
-\multirow{2}{*}{word}                                  & inanimate     &       0.96 &    0.99 &    0.98 & \multirow{2}{*}{0.94} \\
+                                                       & animate     &       0.92 &    0.89 &    0.90 &  \\
+\multirow{2}{*}{word}                                  & inanimate     &       0.96 &    0.99 &    0.98 & \multirow{2}{*}{0.93} \\
                                                        & animate     &       0.94 &    0.78 &    0.85 &  \\
@@
                                                        & animate     &       0.93 &    0.90 &    0.91 &  \\
-\multirow{2}{*}{word + PoS}                            & inanimate     &       0.97 &    0.99 &    0.98 & \multirow{2}{*}{0.96} \\
+\multirow{2}{*}{word + PoS}                            & inanimate     &       0.97 &    0.99 &    0.98 & \multirow{2}{*}{0.95} \\
+                                                       & animate     &       0.94 &    0.86 &    0.89 &  \\
+\rowcolor{Gray}                                        & inanimate     &       0.98 &    0.99 &    0.99 &  \\
+\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings}   & animate     &       0.94 &    0.91 &    0.93 & \multirow{-2}{*}{0.97} \\
+\multirow{2}{*}{word + PoS + lemma}                    & inanimate    &       0.97 &    0.99 &    0.98 & \multirow{2}{*}{0.96} \\
                                                        & animate     &       0.94 &    0.86 &    0.90 &  \\
-\rowcolor{Gray}                                        & inanimate     &       0.98 &    0.99 &    0.99 &  \\
-\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings}   & animate     &       0.94 &    0.92 &    0.93 & \multirow{-2}{*}{0.97} \\
-\multirow{2}{*}{word + PoS + lemma}                    & inanimate    &       0.98 &    0.99 &    0.98 & \multirow{2}{*}{0.96} \\
-                                                       & animate     &       0.94 &    0.87 &    0.90 &  \\
 \multirow{2}{*}{word + PoS + lemma + embeddings}       & inanimate     &       0.98 &    0.99 &    0.99 & \multirow{2}{*}{0.97} \\
-                                                       & animate     &       0.94 &    0.92 &    0.93 &  \\
-\multirow{2}{*}{word + PoS + lemma + dep}              & inanimate     &       0.98 &    0.99 &    0.98 & \multirow{2}{*}{0.96} \\
-                                                       & animate     &       0.94 &    0.88 &    0.91 &  \\
+                                                       & animate     &       0.94 &    0.91 &    0.93 &  \\
+\multirow{2}{*}{word + PoS + lemma + dep}              & inanimate     &       0.97 &    0.99 &    0.98 & \multirow{2}{*}{0.96} \\
+                                                       & animate     &       0.94 &    0.86 &    0.90 &  \\
 \multirow{2}{*}{word + PoS + lemma + dep + embeddings} & inanimate     &       0.98 &    0.99 &    0.99 & \multirow{2}{*}{0.97} \\
@@
 interesting to observe that the `simple' $n$-gram word model already
-performs rather well. Adding more features, like Part-of-Speech,
+performs rather well. Adding more features, such as Part-of-Speech,
 lemmata, etcetera, only has a positive impact on the recall of the
@@
 \multirow{2}{*}{embeddings}                            & inanimate     &       0.89 &    0.94 &    0.92 & \multirow{2}{*}{0.92} \\
-                                                       & animate       &       0.91 &    0.85 &    0.88 &  \\
+                                                       & animate       &       0.92 &    0.84 &    0.88 &  \\
 \multirow{2}{*}{word}                                  & inanimate     &       0.78 &    0.98 &    0.87 & \multirow{2}{*}{0.91} \\
@@
 \multirow{2}{*}{word + pos}                            & inanimate     &       0.86 &    0.96 &    0.90 & \multirow{2}{*}{0.94} \\
-                                                       & animate       &       0.92 &    0.78 &    0.84 &  \\
+                                                       & animate       &       0.93 &    0.78 &    0.84 &  \\
 \rowcolor{Gray}                                        & inanimate     &       0.93 &    0.96 &    0.94 &  \\
@@
                                                        & animate       &       0.94 &    0.80 &    0.86 &  \\
-\multirow{2}{*}{word + PoS + lemma + embeddings}       & inanimate     &       0.93 &    0.96 &    0.94 & \multirow{2}{*}{0.97} \\
+\multirow{2}{*}{word + PoS + lemma + embeddings}       & inanimate     &       0.92 &    0.96 &    0.94 & \multirow{2}{*}{0.97} \\
                                                        & animate       &       0.94 &    0.89 &    0.92 &  \\
@@
 \multirow{2}{*}{word + PoS + lemma + dep + embeddings} & inanimate     &       0.93 &    0.96 &    0.94 & \multirow{2}{*}{0.97} \\
-                                                       & animate       &       0.94 &    0.90 &    0.92 &  \\
+                                                       & animate       &       0.94 &    0.89 &    0.92 &  \\
 \bottomrule

### 2014-12-22 experiment.py
add backoff model

@@ -186,16 +186,16 @@ if __name__ == '__main__':
         columns=['experiment', 'fold', 'class', 'precision', 'recall', 'Fscore', 'AUC'])
     ambiguous_scores = pd.DataFrame(
         columns=['experiment', 'fold', 'class', 'precision', 'recall', 'Fscore', 'AUC'])
-    model = Word2Vec.load(sys.argv[1])
+    model = Word2Vec.load_word2vec_format(sys.argv[1], binary=True)
     # set up a number of experimental settings
     experiments = [('word',), ('word', 'pos'), ('word', 'pos', 'root'),
-                   ('word', 'pos', 'root', 'rel'), tuple(FIELDNAMES)]
+                   ('word', 'pos', 'root', 'rel')] # tuple(FIELDNAMES)
     experiments = experiments + [experiment + ('embeddings', )
                                  for experiment in experiments]
     experiments += [('embeddings', )]
 
     classifiers = {
-        'lr': LogisticRegression(C=1.0),
+        'lr': LogisticRegression,
         'sgd': SGDClassifier(n_iter=100, shuffle=True),
         'svm': LinearSVC(),
         'knn': KNeighborsClassifier(weights='distance')
@@ -211,33 +211,76 @@ if __name__ == '__main__':
         X_train_docs = [X[i] for i in train_index]
         y_train_docs = [label for i in train_index for label in y[i]]
         X_test_docs = [X[i] for i in test_index]
+        test_words = [word[0] for i in test_index for word in X[i]]
         y_test_docs = [label for i in test_index for label in y[i]]
 
         for experiment in experiments:
+            backoff = False
             print "Features: %s" % ', '.join(experiment)
             exp_name = '_'.join(experiment)
             if 'embeddings' in experiment and len(experiment) > 1:
                 features = FeatureStacker(('windower', Windower(window_size=3)),
                                           ('embeddings', WordEmbeddings(model)))
+                backoff_features = Windower(window_size=3)
+                backoff = True
             elif 'embeddings' in experiment:
                 features = WordEmbeddings(model)
                 experiment = ('word', ) + experiment # needed to extract the vectors
             else:
                 features = Windower(window_size=3)
+
             X_train = include_features(X_train_docs, experiment)
             X_test = include_features(X_test_docs, experiment)
             X_train = features.fit_transform(X_train)
             X_test = features.transform(X_test)
+
+            if backoff:
+                X_train_backoff = include_features(
[…]

### 2014-12-22 experiment.py
verbosity

@@ -246,8 +246,8 @@ if __name__ == '__main__':
             y_train = le.fit_transform(y_train_docs)
             y_test = le.transform(y_test_docs)
             # initialize a classifier
-            clf = classifiers[sys.argv[3]]()
-            backoff_clf = classifiers[sys.argv[3]]()
+            clf = classifiers[sys.argv[3]](C=1.0, random_state=1)
+            backoff_clf = classifiers[sys.argv[3]](C=1.0, random_state=1)
             print clf.__class__.__name__
             clf.fit(X_train, y_train)
             if backoff:
@@ -257,6 +257,7 @@ if __name__ == '__main__':
                 preds, pred_probs = [], []
                 for i, word in enumerate(X_test):
                     if test_words[i].lower() not in model:
+                        print "Backoff prediction for", test_words[i]
                         preds.append(backoff_clf.predict(X_test_backoff[i])[0])
                         pred_probs.append(backoff_clf.predict_proba(X_test_backoff[i]))
                     else:
@@ -269,6 +270,7 @@ if __name__ == '__main__':
                 preds, pred_probs = [], []
                 for i, word in enumerate(X_test):
                     if test_words[i].lower() not in model:
+                        print "Default prediction for", test_words[i]
                         preds.append(0)
                         pred_probs.append(np.array([1.0, 0.0]))
                     else:

### 2014-12-22 experiment.py
loading gensim format

@@ -186,7 +186,7 @@ if __name__ == '__main__':
         columns=['experiment', 'fold', 'class', 'precision', 'recall', 'Fscore', 'AUC'])
     ambiguous_scores = pd.DataFrame(
         columns=['experiment', 'fold', 'class', 'precision', 'recall', 'Fscore', 'AUC'])
-    model = Word2Vec.load_word2vec_format(sys.argv[1], binary=True)
+    model = Word2Vec.load(sys.argv[1])
     # set up a number of experimental settings
     experiments = [('word',), ('word', 'pos'), ('word', 'pos', 'root'),
                    ('word', 'pos', 'root', 'rel')] # tuple(FIELDNAMES)

### 2014-12-24 paper/paper.tex
update paper

@@
 \begin{abstract}
-  This paper presents a linguistically uninformed computational model
-  for animacy classification. The model makes use of word $n$-grams in
+  Existing animacy detection systems, while scoring well, seem to be
+  out of sync with current (psycho)linguistic research, which takes a
+  more dynamic approach to animacy. This paper presents a
+  linguistically uninformed computational model for animacy
+  classification. The model makes use of word $n$-grams in
   combinations with lower dimensional word embedding representations
@@
 survival. Those objects could be prey, for example, or predators or
-mates. The fundamental nature of the distinction between animate and
-inanimate is reflected in the fact that the division is acquired very
-early in life~\cite{opfer:02}. Children of less than 6 months old
-are well able to make the division. Moreover, recent brain research
-shows that the distinction appears in the organization of the brain
-(e.g.~\cite{gao:12}). For some researchers this provides evidence for
-the idea that the division between animate and inanimate is an innate
-part of how we see the world.
+mates. The fundamental nature that the distinction between animate and
+inanimate has for humans is reflected in the fact that this division
+is acquired very early in life \cite{opfer:02}: children of less than
+6 months old are well able to distinguish the two categories from one
+another. Moreover, recent brain research shows that the distinction
+appears in the organization of the brain (e.g.~\cite{gao:12}). For
+some researchers, this provides evidence for the idea that the
+division between animate and inanimate is an innate part of how we see
+the world.
 
-A common view supported by philosophers such as Daniel Dennett is that
-animacy and aliveness are to be treated as epistemological
-stances rather than fixed states in the world \cite{dennett:96}: not
-ineffable qualia but behavioral capacity defines our stance towards
-objects. Depending on whether people \textit{think} that an object is
-animate, they utilize different cognitive strategies to explain and
-predict the actions of those objects. Th following opening of a
-well-known story will make this point more clear:
+Traditionally, even though animacy may be a scalar rather than a
+strictly categorical distinction (see e.g. the animacy hierarchy in
+\cite{comrie:89} and research such as \cite{rosenbach:08}), the
+animate/inanimate distinction is taken as fixed with regard to lexical
+items. This standpoint has been challenged, however, by researchers
+from different fields. Firstly, it has long been established in
+linguistic typology that not all languages award animacy to the same
+entities in different grammatical categories: as \cite{comrie:89} for
+example notes, English, and many other languages, distinguishes
+between human and not-human in the choice of pronouns, whereas Russian
+distinguishes between animate, entailing humans and animals, versus
+non-animate, in their interrogative pronouns. This indicates different
+subdivisions of animacy in the respective languages. Secondly,
+philosophers such as Daniel Dennett support the view that animacy and
+aliveness are to be treated as epistemological stances rather than
+fixed states in the world: not ineffable qualia but behavioral
+capacity defines our stance towards objects \cite{dennett:96}. In
+other words, depending on whether people \textit{think} that an object
+is animate, they utilize different cognitive strategies to explain and
+predict the actions of those objects. Finally, evidence from
+psycholinguistic research has accumulated to support the view of
+animacy as a cognitive viewpoint rather than an extra-perceptive
+absolute. \cite{nieuwland:05} for example show that test subjects
+readily accept animate behavior from inanimate objects within the
+proper contexts, and \cite{vogels:13} moreover emphasize the relation
+between animacy and motion, showing that factors such as
+self-propelment play a crucial role in recognizing and/or awarding
+animacy to certain objects. This is exemplified in the opening of this
+well-known story:
 
@@
 
-As readers (or listeners) of a story, we choose to view participating
-objects at varying levels of abstraction in order to predict their
-behavior. Dennett \cite{dennett:96} defines three levels of
-abstraction: (1) the \textit{physical stance}, (2) the \textit{design
-  stance} and (3) the \textit{intentional stance}. The physical stance
-deals with predictions about objects given their physical
-properties. The design stance is slightly more abstract and deals with
-concepts like purpose, function or design. Finally the intentional
-stance is concerned with belief, thinking and intentions. These are
-all cognitive strategies we use to predict and explain the actions of
-objects in our environment. Interestingly, in the process of reading
-the opening of the story about the fleeing pancake, readers and
-listeners experience the transition from one strategy to the next
-quite clearly. Initially, the pancake is interpreted from a physical
-stance or perhaps the more abstract design stance in terms of the
-purpose (i.e. to stave off hunger). It is only at the last adverbial
-phrase `as hard as he could' that we start to wonder whether we should
-adopt to the yet more abstract intentional stance and consider the
-pancake to be a rational agent.
+Although initially, based on their knowlegde of the world, readers
+will regard the pancake as inamimate, the self-propelled motion verb
+`rolled' initiates our shift towards an animate interpretation of the
+pancake. As readers (or listeners) of a story, we choose to view
+participating objects at varying levels of abstraction in order to
+predict their behavior. Dennett \cite{dennett:96} defines three levels
+of abstraction: (1) the \textit{physical stance}, (2) the
+\textit{design stance} and (3) the \textit{intentional stance}. The
+physical stance deals with predictions about objects given their
+physical properties. The design stance is slightly more abstract and
+deals with concepts like purpose, function or design. Finally the
+intentional stance is concerned with belief, thinking and
+intentions. These are all cognitive strategies we use to predict and
+explain the actions of objects in our environment. Interestingly, in
+the process of reading the opening of the story about the fleeing
+pancake, readers and listeners experience the transition from one
+strategy to the next quite clearly. Initially, the pancake is
+interpreted from a physical stance, or perhaps the more abstract
+design stance in terms of the purpose (i.e. to stave off hunger). It
+is only at the last adverbial phrase `as hard as he could' that we
+start to wonder whether we should adopt to the yet more abstract
+intentional stance and consider the pancake to be a rational agent.
 
@@
 machine learning techniques, or make a combination of both. Common to
-most approaches is that they make use of semantic lexicons with
-information about animacy and syntactic cues in a text. Both feature
-types are relatively costly to obtain as they require large
-vocabularies or syntactic parsing systems, which, with the exception
-of a few languages, are not readily available.
+most approaches is the fact that they make use of semantic lexicons
+with information about animacy, as well as syntactic cues in a
+text. Both feature types are relatively costly to obtain as they
+require large vocabularies or syntactic parsing systems, which, with
+the exception of a few languages, are not readily available.
 
 In this paper we present a new linguistically uninformed model to
-automatically label texts for animacy. We show that we can (mostly)
-do away with features that require syntactic parsing or semantic
-lexicons while still yielding competitive performance. We focus
-on labeling animacy in stories because stories pose some interesting
-problems to automatic systems of animacy recognition. As the example
-of the fleeing pancake already made clear, quite some entities in
-stories, though inanimate in the `real' world, exhibit all behavioral
-capacities to consider them animate and rational agents. Such
-examples make a clear case for developing dynamic, data driven systems
-that do not rely too much on static and fixed world knowledge.
+automatically label texts for animacy. We show that we can (mostly) do
+away with features that require syntactic parsing or semantic lexicons
+while still yielding competitive performance. We focus on labeling
+animacy in stories because stories pose some interesting problems to
+automatic systems of animacy recognition. As the example of the
+fleeing pancake already made clear, quite some entities in stories,
+though inanimate in the `real' world, exhibit all behavioral
+capacities to consider them animate and rational agents. Another
+example is the \textit{Sorcerer's Apprentice} sequence in Walt
+Disney's famous \textit{Fantasia}, in which brooms display the ability
+to collect buckets of water. Such examples, in which otherwise as
+inanimatedly labeled entities such as pancakes and brooms act, make a
+clear case for developing dynamic, data driven systems that do not
+rely too much on static and fixed world knowledge, but rather on
+immediate context.
 
@@
 short overview of existing techniques for automatically labeling
-animacy in texts (\S\ref{sec:previous-work}). After a description of
-the corpus used in our study and how the annotations of the corpus
-have been established (\S\ref{sec:data}), we will give an account of
-our computational models in Section \ref{sec:models}. We report on the
+animacy in texts, including the definitions of animacy used in these
+papers (\S\ref{sec:previous-work}). After a description of the corpus
+used in our study and how the annotations of the corpus have been
+established (\S\ref{sec:data}), we will give an account of our
+computational models in Section \ref{sec:models}. We report on the
 empirical results in Section \ref{sec:results}. The last section
@@
 \section{Previous Work}\label{sec:previous-work}
-
-%ik wil eigenlijk meer uitleg in deze paragraaf:
-In most of the literature on automatic animacy detection, a static,
-binary distinction is made between animate and
-inanimate. \cite{bowman:12} for example, define objects as animate if
-they are alive and have the ability to move under their own
-will. \cite{orasan:07} define animacy in the context of anaphora
-resolution: something is animate ``if its referent can also be
-referred to using one of the pronouns he, she, him, her, his, hers,
-himself, herself, or a combination of such pronouns
-(e.g. his/her)''. However, some parts of these definitions are
-somewhat circular and essentialistic. More importantly, they are not
-particularly applicable to the rich and wondrous entities that live in
-the realm of stories. As our example above showed, it is possible for
-an otherwise as inanimate labeled object to attain animacy. We
-therefore choose to define animacy in terms of Dennett's intentional
-stance, which is more dynamic, and which ultimately comes down to the
-question whether ``you decide to treat the object whose behavior is to
-be predicted as a rational agent'' \cite[pp. 17]{dennett:96}.
 
@@
 
-% commentary on above approaches
 The approaches mentioned above present us with a number of
-problems. First, almost all of them heavily rely on costly,
+problems. Firstly, almost all of them heavily rely on costly,
 linguistically informed features derived from lexical-semantic
@@
 animacy detection, however, leads to a chicken and egg
-situation. Another problem with the approaches above is that they are
-lemma-based, which means that the models are generally insensitive to
-different usages of a particular word. A pancake is typically not an
-animate entity, but, as we have seen, whether it is animate or not may
-depend on the story in which it appears, and even within the story the
-animacy may change. We therefore need to develop a dynamic, data
-driven system that is token based and does not rely too much on static
-and fixed world knowledge.
+situation.
+
+Another major problem with the approaches above is, as was said above,
+that they are lemma-based, which means that the models are generally
+insensitive to different usages of a particular word in particular
+contexts. In other words, in most of the literature on automatic
+animacy detection, a static, binary distinction is made between
+animate and inanimate. \cite{bowman:12} for example, define objects as
+animate if they are alive and have the ability to move under their own
+will. \cite{orasan:07} define animacy in the context of anaphora
+resolution: something is animate ``if its referent can also be
+referred to using one of the pronouns he, she, him, her, his, hers,
+himself, herself, or a combination of such pronouns
+(e.g. his/her)''. However, as was explained above, these definitions
+are not necessarily in line with current linguistic and neurological
+research, and seemingly ignore findings from research such as
+\cite{nieuwland:05}. Similarly, they are not particularly applicable
+to the rich and wondrous entities that live in the realm of
+stories. As was shown above, although a pancake is typically not an
+animate entity, its animacy depends on the story in which it appears,
+and even within the story the animacy may change. To accomodate this
+possibility, we therefore choose to define animacy in terms of
+Dennett's intentional stance, which is more dynamic, and which
+ultimately comes down to the question whether ``you decide to treat
+the object whose behavior is to be predicted as a rational agent''
+\cite[pp. 17]{dennett:96}. Our system for animacy detection therefore
+needs to be dynamic, data driven, and token based. It cannot rely too
+heavily on static and fixed world knowledge.
 
@@
 To develop this dynamic, data driven system we used a corpus of Dutch
-folktales. The corpus consists of 74 Dutch stories from the collection
-\textit{Volkssprookjes uit Nederland en Vlaanderen} compiled by
+folktales. One of the reasons to use folk tales was that, as
+\cite{vogels:13} note, `In cartoons or fairy tales [\ldots] inanimate
+entities or animals are often anthropomorphized', which means that the
+material could yield interesting cases of unexpected animacy, as is
+the case with the pancake in the abovementioned story. Our corpus
+consists of 74 Dutch stories from the collection
+\textit{Volkssprookjes uit Nederland en Vlaanderen}, compiled by
 \cite{sinninghe:78}. The collection is made up of Dutch retellings of
-popular stories, including such tales as \textit{The Bremen Town
-  Musicians} (ATU 130) and \textit{The Table, the Ass, and the Stick }
-(ATU 563), as well as lesser-known stories such as \textit{The Singing
-  Bone} (ATU 780) and \textit{Cock, Hen, Duck, Pin, and Needle on a
-  Journey} (ATU 210). This last story is again a clear example of
-otherwise inanimate objects being animate in a folk story, as it
-concerns the adventures of several household items, such as a
-\textit{pin}, a \textit{hackle}, an \textit{egg}, and a
+popular and widespread stories, including such tales as \textit{The
+  Bremen Town Musicians} (ATU 130) and \textit{The Table, the Ass, and
+  the Stick } (ATU 563), as well as lesser-known stories such as
+\textit{The Singing Bone} (ATU 780) and \textit{Cock, Hen, Duck, Pin,
+  and Needle on a Journey} (ATU 210). This last story is again a clear
+example where otherwise inanimate objects are animated in a folk
+story, as it concerns the adventures of several household items, such
+as a \textit{pin}, a \textit{hackle}, an \textit{egg}, and a
 \textit{whetstone}. A digital version of the collection is available
@@
 single collection for our corpus presents us with a helpful
-homogeneity with regard to the editor, length and language use. On the
-other hand, the collection displays diversity as well in the choice of
-the stories, which entail fairy tales, legends, and sagas.
-%getallen nakijken%
-%moet je de labels hier al wel noemen? want dat lijkt me eerder een resultaat dan een pre-processing stap toch?%
+homogeneity with regard to the editor, length of the stories (which is
+between X and Y words) and language use. On the other hand, the
+collection displays diversity as well in the choice of the stories,
+which entail fairy tales, legends, and sagas.
 
-The data were labeled by two annotators. We computed an
-inter-annotator agreement score (Cohen's Kappa) of $K=0.95$ on the
-basis of five stories that were annotated by both annotators. The
-following sentence provides an example annotation. All unlabeled words
-are implicitly considered to be inanimate.
+All together, the corpus consists of 74504 words, from 5549 unique
+words. Using the annotation tool brat (brat rapid annotation tool), an
+online environment for collaborative editing\footnote{Available here:
+  \url{http://brat.nlplab.org}}, two annotators labeled words for
+animacy, within the context of the story.\footnote{On the basis of
+  five stories that were annotated by both annotators we computed an
+  inter-annotator agreement score (Cohen's Kappa) of $K=0.95$.} All
+unlabeled words were implicitly considered to be inanimate. The
+following sentence provides an example annotation.
 \begin{exe}
@@
 takken, en dan ga jij kleermaker aan zijn benen hangen en zo gaan we maar door\\
-    \textsc{anim} \textsc{anim} \textsc{anim} {} {} {} {} \textsc{anim} {} {} {} {} {} {} {} {} \textsc{anim} \textsc{anim}
-    {} \textsc{anim} {} {} {} {} {} \textsc{anim} {} {} \\
+    \textsc{animate} \textsc{animate} \textsc{animate} {} {} {} {} \textsc{animate} {} {} {} {} {} {} {} {} \textsc{animate} \textsc{animate}
+    {} \textsc{animate} {} {} {} {} {} \textsc{animate} {} {} \\
 \trans `You, blacksmith, you are the strongest; hold on to the upper
@@
 \end{exe}
-The corpus consists of 74504 words, from 5549 unique words. All word
-tokens were assigned a binary label `animate' or `inanimate', where
-all words that were not animate (e.g. determiners, adjectives, etc.)
-were labelled as inanimate. This resulted in 11542 animate tokens of
-743 types, and 62926 inanimate tokens from 5011 unique inanimate
-words. A selection of words, such as the abovementioned
-\textit{pancake} and \textit{egg}, was labeled as both animate and
-inanimate, depending on the context. The distribution of the nouns is
-telling: of the animate tokens, 4627 (40,1\%) are nound, while only
-6878 of the inanimate tokens (11 \%) are nouns. This shows that being
-a noun is already somewhat of an indication for animacy.
+
+Because we interpreted animacy within the context of the story, the
+same lexical item could be labeled differently in different stories.
+For example, in the abovementioned example of the pancake, which
+occurs in SINVS076 in our corpus, the pancake is tagged consistently
+as `animate'. In another story, SINVS042, where at one point a soldier
+is baking pancakes, the pancakes do not act, and are thus not labeled
+as `animate'. The following sentences show how this was employed in
+practice.
+\begin{exe}
+\ex
+\gll Terwijl hij de pannekoek bakte, keek hij naar het ding, dat uit de schouw gevallen was \\
+    {} \textsc{animate} {} {} {} {} \textsc{animate} {} {} {} {} {} {} {} {} {} {}\\
+\trans `While he was baking the pancake, he looked at the thing, which had fallen from the hearth\ldots'
+\end{exe}
+\begin{exe}
+\ex
+\gll Toevallig stond de deur van de keuken open en de pannekoek rolde naar buiten, het veld in, zo hard hij maar kon.\\
+    {} {} {} {} {} {} {} {} {} {} \textsc{animate} {} {} {} {} {} {} {} {} {}\textsc{animate} {} {} \\
+\trans `Coincidentally the door of the kitchen was open and the pancake rolled outside, into the field, as fast as it could'
+\end{exe}
+
+This annotation resulted in 11542 animate tokens of 743 types, while
+implicitly yielding 62926 inanimate tokens from 5011 unique inanimate
+words. Because of our context-dependent approach, some words, such as
+\textit{pancake} and \textit{egg}, occurred in both animate types as
+inanimate types, because they were labeled as both animate and
+inanimate in some cases in our corpus. It is telling that of the
+animate tokens 4627 (40,1\%) were nouns, while only 6878 of the
+inanimate tokens (11 \%) are nouns. This shows that being a noun is
+already somewhat of an indication for animacy.
 
@@
 state of the art syntactic parser for Dutch, Alpino \cite{bouma:01}.
-
-% gaat nu de volgorde niet fout? Eerst tokenizeren en parsen, en dan animate/inanimate toekennen?%
+%dit kan iets uitgebreider eventueel%
 
@@
 This section describes our experimental setup including the features
-used, the machine learning models we applied and our methods of
+used, the machine learning models we applied, and our methods of
 evaluation.\footnote{The data set and the code to perform the
@@
 \subsection{Task description}
-
+%misschien deze subsection invoegen in bovenstaande inleiding. FK, nee
+%ik vind dit altijd een erg prettige plaats (eventueel een
+%soortgelijke zin in de inleiding, maar hier mag-ie niet weg :-))
 We formulate the problem of animacy detection as a classification
 problem where the goal is to assign a label at word level, rather than
-at lemma level, indicating whether it represents an animate or an
-inanimate word.
+at lemma level. This label indicates whether the word is classified as
+animate or inanimate.
 
@@
 Animate words are far outnumbered by inanimate words in our collection
-(see (\S\ref{sec:data})). Reporting accuracy scores would therefore
+(see \S\ref{sec:data}). Reporting accuracy scores would therefore
 provide skewed results, strongly favoring the majority category. The
@@
 well-known $F1$-score more appropriate. For all experiments we report
-on the precision, recall and $F1$-score \cite{rijsbergen:79}. Also,
-while in most of the literature on animacy detection results are only
-presented for the classification of nouns or noun phrases, we will
-present these and also report on the results for all words in a text.
+on the precision, recall and $F1$-score~\cite{rijsbergen:79}. As a
+second measure of evaluation, we compute the Area Under the Precision
+Recall Curve (AUC_{PR}) which also takes into account possible class
+imbalance. Also, while in most of the literature on animacy detection
+results are only presented for the classification of nouns or noun
+phrases, we will present these while furthermore reporting on the
+results for all words in a text.
+%bovenstaande zin loopt nog niet helemaal lekker%
 
@@
 text. \cite{bowman:12} fall into this trap by making a random split in
-their data on the sentence level. It is very likely that in their
-setup sentences from the same document are present in both the
-training data and the test data, making their evaluation
-unrealistic. To circumvent this problem, we split the data at the
-story level. We make use of 10-fold cross-validation, where we
-randomly select 10\% of the stories for testing and train on the
-remaining 90\%.
+their data on the sentence level. In such a setup, it is highly likely
+that sentences from the same document are present in both the training
+data and the test data, making their evaluation unrealistic. To
+circumvent this problem, we split the data at the story level. We make
+use of 10-fold cross-validation, where we randomly select 10\% of the
+stories for testing and train on the remaining 90\%.
 
@@
 \subsubsection{Lexical features}
-We take a sliding window approach where for each focus word (i.e.\ the
+We take a sliding window approach where for each focus word (i.e. the
 word for which we want to predict whether it is animate or not) we
-extract $n$ words to the left, $n$ words to the right and the focus
-word itself. In all experiments we set $n$ to 3. In addition to the
-word forms, for each word in a window we extract its lemma.
+extract both $n$ words to the left and $n$ words to the right, as well
+as the focus word itself. In all experiments we set $n$ to 3. In
+addition to the word forms, for each word in a window we also extract
+its lemma.
 
@@
 part-of-speech tagger. Again, we take a sliding window approach and
-extract the part-of-speech tags for three words left and right of the focus word, as well as the tag of the focus word itself.
-
+extract the part-of-speech tags for three words left and right of the
+focus word, as well as the tag of the focus word itself.
+%waarom Alpino voor comparibility, omdat we dat ook gebruikten in het taggen van het corpus?%
 \subsubsection{Syntactic Features}
-We extract the dependency tag for each word and its $n$ neighbors to
+We extract the dependency tag for each word and its $n=3$ neighbors to
 the right and to the left as provided by the syntactic parser
@@
 proven to perform rather well.
-
-%is n hier weer 3? Je specificeert dat niet%
 
@@
 can automatically induce so-called `word embeddings' from large,
-unannotated collections of texts. These models typically attempt to
-learn a lower dimensional vector for each word in the vocabulary which
-captures the typical co-occurrence patterns of the word in the
+unannotated collections of texts
+(e.g.~\cite{mikolov:13,pennington:14}). These models typically attempt
+to learn a lower dimensional vector for each word in the vocabulary
+which captures the typical co-occurrence patterns of the word in the
 corpus. The similarity between words can then be approximated by
-applying arithmetics to these vectors of word embeddings.
-
-%je moet een of meerdere bronnen noemen hier, by "recently there has been interest"%
+applying similarity metrics, such as the cosine metric, to these
+vectors of word embeddings.
 
@@
 
-We add the word embedding vectors to our animacy model as follows. We
-augment each word in the data with its corresponding 300 dimension
-word embeddings vector. Out-of-vocabulary words were given a vector in
-which all dimensions are set to zero.
-
 \subsection{Models}
@@
 
+Although our background corpus is sufficiently large to cover most
+words in an unseen text, there will always be rare words for which we
+do not have learned word embeddings. Therefore, in order to
+effectively make use of the word embedding vectors, we need a way to
+deal with out-of-vocabulary items. We adopt a simple strategy were we
+make use of a primary classifier and a back-off classifier. For models
+6 to 9, we augment each word with its corresponding 300 dimension word
+embeddings vector. In the case of out-of-vocabulary words, we resort
+to a back-off model that contains all features except the word
+embeddings. For example, a model that makes use of words and word
+embeddings, will make a prediction on the basis of the word features
+alone. In case of the model that solely uses the embeddings (model 5),
+the back-off classifier is represented by a majority vote classifier,
+which classifies unseen words as inanimate.
+
 \section{Results}\label{sec:results}
@@
 precision, recall, $F1$-score and Area Under the Precision-Recall
-Curve (AUC) for the animate words and the
+Curve (AUC_{PR}) for the animate words and the
 inanimate words.
@@
 \toprule
-features & class & precision &  recall & $F1$ &  AUC \\
+features & class & Precision &  Recall & $F1$ &  AUC_{PR} \\
 \midrule
 \multirow{2}{*}{embeddings}                            & inanimate     &       0.98 &    0.99 &    0.98 & \multirow{2}{*}{0.93} \\
-                                                       & animate     &       0.92 &    0.89 &    0.90 &  \\
+                                                       & animate     &       0.93 &    0.89 &    0.91 &  \\
 \multirow{2}{*}{word}                                  & inanimate     &       0.96 &    0.99 &    0.98 & \multirow{2}{*}{0.93} \\
@@
 \multirow{2}{*}{word + embeddings}                     & inanimate     &       0.98 &    0.99 &    0.98 & \multirow{2}{*}{0.96} \\
-                                                       & animate     &       0.93 &    0.90 &    0.91 &  \\
+                                                       & animate     &       0.94 &    0.90 &    0.91 &  \\
 \multirow{2}{*}{word + PoS}                            & inanimate     &       0.97 &    0.99 &    0.98 & \multirow{2}{*}{0.95} \\
@@
 \end{tabular}
-\caption{Precision, recall, $F1$ and Area Under the Curve (AUC) scores for animate and inanimate classes per feature setting for all words.}
+\caption{Precision, Recall, $F1$-score and Area Under the Precision Recall Curve (AUC_{PR}) scores for animate and inanimate classes per feature setting for all words.}
 \label{tab:results-all}
@@
 
-The best performance is
-achieved by the model that combines the word features, part-of-speech
-tags and the word embeddings with an $F1$-score of 0.93 for animate
-words and 0.99 on inanimate words. Adding more features does not
-result in any more performance gain.
-
+The best performance is achieved by the model that combines the word
+features, part-of-speech tags and the word embeddings, which has an
+$F1$-score of 0.93 on animate words and 0.99 on inanimate
+words. Adding more features does not result in any more performance
+gain.
+%misschien hier dan nog een voorbeeld noemen, of verwijzen naar relevant cijfer in de tabel%
 Table \ref{tab:results-noun} presents the results for all nouns and
 names in the data set. The best performance is again achieved by the
-model that combines the word features with the Part-of-Speech tags and
+model that combines the word features with the part-of-speech tags and
 word embeddings, resulting in an $F1$-score of 0.92 for animate
-instances and 0.94 for inanimate instances.
+instances and 0.95 for inanimate instances.
 
@@
 \toprule
-features & class &  precision &  recall & $F1$ &  AUC \\
+features & class &  Precision &  Recall & $F1$ &  AUC_{PR} \\
 \midrule
-\multirow{2}{*}{embeddings}                            & inanimate     &       0.89 &    0.94 &    0.92 & \multirow{2}{*}{0.92} \\
-                                                       & animate       &       0.92 &    0.84 &    0.88 &  \\
+\multirow{2}{*}{embeddings}                            & inanimate     &       0.90 &    0.96 &    0.92 & \multirow{2}{*}{0.91} \\
+                                                       & animate       &       0.93 &    0.85 &    0.89 &  \\
 \multirow{2}{*}{word}                                  & inanimate     &       0.78 &    0.98 &    0.87 & \multirow{2}{*}{0.91} \\
                                                        & animate       &       0.96 &    0.60 &    0.74 &  \\
-\multirow{2}{*}{word + embeddings}                     & inanimate     &       0.90 &    0.96 &    0.93 & \multirow{2}{*}{0.96} \\
-                                                       & animate       &       0.94 &    0.86 &    0.89 &  \\
-\multirow{2}{*}{word + pos}                            & inanimate     &       0.86 &    0.96 &    0.90 & \multirow{2}{*}{0.94} \\
+\multirow{2}{*}{word + embeddings}                     & inanimate     &       0.90 &    0.97 &    0.93 & \multirow{2}{*}{0.96} \\
+                                                       & animate       &       0.95 &    0.85 &    0.90 &  \\
+\multirow{2}{*}{word + PoS}                            & inanimate     &       0.86 &    0.96 &    0.90 & \multirow{2}{*}{0.94} \\
                                                        & animate       &       0.93 &    0.78 &    0.84 &  \\
-\rowcolor{Gray}                                        & inanimate     &       0.93 &    0.96 &    0.94 &  \\
-\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings} & animate       &       0.94 &    0.89 &    0.92 &\multirow{-2}{*}{0.97}  \\
+\rowcolor{Gray}                                        & inanimate     &       0.93 &    0.96 &    0.95 &  \\
+\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings} & animate       &       0.95 &    0.90 &    0.92 &\multirow{-2}{*}{0.97}  \\
 \multirow{2}{*}{word + PoS + lemma}                    & inanimate     &       0.87 &    0.96 &    0.91 & \multirow{2}{*}{0.95} \\
                                                        & animate       &       0.94 &    0.80 &    0.86 &  \\
-\multirow{2}{*}{word + PoS + lemma + embeddings}       & inanimate     &       0.92 &    0.96 &    0.94 & \multirow{2}{*}{0.97} \\
-                                                       & animate       &       0.94 &    0.89 &    0.92 &  \\
+\multirow{2}{*}{word + PoS + lemma + embeddings}       & inanimate     &       0.93 &    0.96 &    0.94 & \multirow{2}{*}{0.97} \\
+                                                       & animate       &       0.95 &    0.89 &    0.92 &  \\
 \multirow{2}{*}{word + PoS + lemma + dep}              & inanimate     &       0.87 &    0.96 &    0.91 & \multirow{2}{*}{0.95} \\
                                                        & animate       &       0.93 &    0.80 &    0.86 &  \\
-\multirow{2}{*}{word + PoS + lemma + dep + embeddings} & inanimate     &       0.93 &    0.96 &    0.94 & \multirow{2}{*}{0.97} \\
-                                                       & animate       &       0.94 &    0.89 &    0.92 &  \\
+\multirow{2}{*}{word + PoS + lemma + dep + embeddings} & inanimate     &       0.93 &    0.96 &    0.95 & \multirow{2}{*}{0.97} \\
+                                                       & animate       &       0.95 &    0.90 &    0.92 &  \\
 \bottomrule
 \end{tabular}
-\caption{Precision, recall, $F1$ and Area Under the Curve (AUC) scores for animate and inanimate classes
+\caption{Precision, Recall, $F1$ score and Area Under the Precision Recall Curve (AUC_{PR}) scores for animate and inanimate classes
   per feature settings for all words tagged as noun.}
@@
 \end{table}
+
+\section{Animacy Detection in the Wild}
+
+In a classical evaluation setup -- as with our approach -- general
+practice is to divide an annotated data set into a set for training a
+computational system of which the performance is evaluated on a
+held-out test set. Even if we include a separate development set, this
+setup always runs the risk over overfitting the data. Our annotated
+corpus contains a reasonably diverse set of stories, yet it is fairly
+small and stylistically homogeneous. Even though we performed a
+cross-validation experiment, the chances of overfitting are
+considerably high. In this section we investigate how well our animacy
+detection model performs on a collection of unannotated texts.
+
+The Dutch Folktale Database\footnote{\url{http://www.verhalenbank.nl}}
+is a collection of about 42,000 folktales~\cite{meder:10}. The
+collection contains stories from various genres (e.g.~fairytales,
+legends, urban legends, jokes, personal narratives) in a number of
+variants of Dutch and in Frisian. Every entry in the database contains
+meta-data about the story, including language, collector, place and
+date of narration, keywords, names, and subgenre. We make use of a
+subcollection of the Dutch Folktale Database which comprises 16253
+stories written in standard Dutch.
 

### 2015-01-20 paper/paper.tex
updates

@@
 
-\usepackage{gb4e}
+\usepackage{gb4e, rotating}
 \bibliographystyle{plain}
@@
   Amsterdam, The Netherlands\\
-  \texttt{folgert.karsdorp@meertens.knaw.nl}}
+  \texttt{fbkarsdorp@fastmail.nl}}
 \affil[2]{Meertens Institute\\
@@
   linguistically uninformed computational model for animacy
-  classification. The model makes use of word $n$-grams in
+  classification. This model makes use of word $n$-grams in
   combinations with lower dimensional word embedding representations
@@
 \end{abstract}
-
+%aanpassen op basis van embedding verhalenbank-evaluatie%
 \section{Introduction}
@@
 
-Traditionally, even though animacy may be a scalar rather than a
+Although animacy may be a scalar rather than a
 strictly categorical distinction (see e.g. the animacy hierarchy in
 \cite{comrie:89} and research such as \cite{rosenbach:08}), the
-animate/inanimate distinction is taken as fixed with regard to lexical
-items. This standpoint has been challenged, however, by researchers
+animate/inanimate distinction is traditonally taken as fixed with regard to lexical
+items: something is either animate (e.g. a human) or not (e.g. a chair). This standpoint has been challenged, however, by researchers
 from different fields. Firstly, it has long been established in
 linguistic typology that not all languages award animacy to the same
-entities in different grammatical categories: as \cite{comrie:89} for
-example notes, English, and many other languages, distinguishes
+entities in different grammatical categories: as \cite{comrie:89} notes, for
+example, English, and many other languages, distinguishes
 between human and not-human in the choice of pronouns, whereas Russian
-distinguishes between animate, entailing humans and animals, versus
-non-animate, in their interrogative pronouns. This indicates different
+distinguishes between animate (entailing humans and animals) versus
+non-animate (entailing everything else) in its interrogative pronouns. This indicates different
 subdivisions of animacy in the respective languages. Secondly,
@@
 predict the actions of those objects. Finally, evidence from
-psycholinguistic research has accumulated to support the view of
+psycholinguistic research has accumulated to support this view of
 animacy as a cognitive viewpoint rather than an extra-perceptive
@@
 animacy to certain objects. This is exemplified in the opening of this
-well-known story:
+well-known storyhttp://www.verhalenbank.nl/items/show/9636:
 
@@
 automatic systems of animacy recognition. As the example of the
-fleeing pancake already made clear, quite some entities in stories,
-though inanimate in the `real' world, exhibit all behavioral
-capacities to consider them animate and rational agents. Another
+fleeing pancake already made clear, in stories quite some entities can exhibit all behavioral
+capacities to consider them animate and rational agents, even though they are inanimate in the `real' world. Another
 example is the \textit{Sorcerer's Apprentice} sequence in Walt
@@
 to collect buckets of water. Such examples, in which otherwise as
-inanimatedly labeled entities such as pancakes and brooms act, make a
+inanimatedly labeled entities such as pancakes and brooms act as animate beings, make a
 clear case for developing dynamic, data driven systems that do not
@@
 computational models in Section \ref{sec:models}. We report on the
-empirical results in Section \ref{sec:results}. The last section
-offers our conclusions and possible directions for future research.
+empirical results in Section \ref{sec:results}. Next, we provide an evaluation on a larger dataset, while also showing a real world application of our animacy detection system \ref{sec:listing}.
+The last section offers our conclusions and possible directions for future research.
 
@@
 
-To develop this dynamic, data driven system we used a corpus of Dutch
-folktales. One of the reasons to use folk tales was that, as
+To develop this dynamic, data driven system we use a corpus of Dutch
+folktales. One of the reasons to use folk tales is that, as
 \cite{vogels:13} note, `In cartoons or fairy tales [\ldots] inanimate
@@
 material could yield interesting cases of unexpected animacy, as is
-the case with the pancake in the abovementioned story. Our corpus
-consists of 74 Dutch stories from the collection
+the case with the pancake in the abovementioned story.
+
+Our initial corpus consists of 74 Dutch stories from the collection
 \textit{Volkssprookjes uit Nederland en Vlaanderen}, compiled by
@@
 already somewhat of an indication for animacy.
-
+%hoeveel woorden hebben we eigenlijk die dubbel voorkomen, zoals pannekoek en ei? Leuk om te melden%
 After tokenization with the tokenization module of the Python software
@@
 \subsection{Task description}
-%misschien deze subsection invoegen in bovenstaande inleiding. FK, nee
-%ik vind dit altijd een erg prettige plaats (eventueel een
-%soortgelijke zin in de inleiding, maar hier mag-ie niet weg :-))
+
 We formulate the problem of animacy detection as a classification
@@
 relative rarity of animate words makes evaluation measures such as the
-well-known $F1$-score more appropriate. For all experiments we report
-on the precision, recall and $F1$-score~\cite{rijsbergen:79}. As a
+well-known $F1$-score more appropriate. For this reason, we report
+on the precision, recall and $F1$-score~\cite{rijsbergen:79} for all experiments. As a
 second measure of evaluation, we compute the Area Under the Precision
@@
 results are only presented for the classification of nouns or noun
-phrases, we will present these while furthermore reporting on the
+phrases, we will, while reporting on nouns and noun phrases as well, additionally report on the
 results for all words in a text.
-%bovenstaande zin loopt nog niet helemaal lekker%
 
@@
 
-\subsubsection{Semantic Features} The most innovative feature we have
+\subsubsection{Semantic Features} 
+The most innovative feature we have
 included in our model is concerned with semantic similarity. In his
@@
 that ``Die Bedeutung eines Wortes ist sein Gebrauch in der Sprache''
-(PI 43). The well-known insight in computational linguistics that the
+(PI 43). This is reflected by the well-known insight in computational linguistics that the
 meaning of words can be approximated by comparing the linguistic
@@
 similarities between animate words can be estimated by inspecting the
-context in which they occur. From this follows for example that the word
+context in which they occur. From this follows, for example, that the word
 embeddings of an animate word are more similar to those of other
-animate words, as opposed to the embeddings of inanimate words. To give an
-example, the 5 nearest neighbors of \textit{Assepoester}
-(`Cinderella') are: \textit{Sneeuwwitje} (`Snow White') (sim = 0.807),
-\textit{Roodkapje} (`Little Red Riding Hood') (sim = 0.747),
-\textit{Doornroosje} (`Sleeping Beauty') (sim = 0.715),
-\textit{Repelsteeltje} (`Rumpelstiltskin') (sim = 0.669) and
-\textit{Alladin} (0.645). In contrast, the 5 most similar words to
-\textit{boek} (`book') are: \textit{boekje} (`book-\textsc{dim}')
-(sim=0.8), \textit{kinderboek} (`children's book') (0.757),
-\textit{stripboek} (`comic') (0.713), \textit{kortverhaal} (`short
-story') (0.709), \textit{jeugdboek} (`youth book') (0.706).
+animate words, as opposed to the embeddings of inanimate words.
 
+\begin{figure}[t]
+\centering
+\includegraphics[width=\textwidth]{pca.pdf}
+\caption{Two-dimensional PCA projection of the 300 dimensional word embedding
+vectors for a number of animate and inanimate words. The horizontal line
+illustrates the separabilty between the two classes in the first dimension.}
+\label{fig:pca}
+\end{figure}
+
+To give an illustration of this idea, in Figure \ref{fig:pca} we depict a
+two-dimensional Principle Component Analysis (PCA) projection of the 300 dimensional
+word embedding vectors for a number of typically animate and typically inanimate
+ words. The horizontal grey line in the plot illustrates the separability of
+ the animate and inanimate words in the first dimension of the PCA projection.
+ It is interesting to observe that the less animate entity \emph{ghost} is the
+ one closest to all other inanimate entities. Likewise, words such as \emph{castle},
+ \emph{house} or \emph{car} are often used in figurative language (metonomy),
+ for example to refer to the people owning or living in the castle. Perhaps this
+ ambiguous animacy position is reponsible for their position in the first dimension
+ close to real animate entities.
+%dit stukje hierboven kan nog iets worden uitgewerkt, eventueel met voorbeelden%
 \subsection{Models}
@@
 effectively make use of the word embedding vectors, we need a way to
-deal with out-of-vocabulary items. We adopt a simple strategy were we
+deal with out-of-vocabulary items. We adopt a simple strategy where we
 make use of a primary classifier and a back-off classifier. For models
@@
 inanimate words.
-
+%AUC PR scores staan nu niet in de tabel, en worden schijnbaar niet gerapporteerd%
+%positie van de tabellen is niet optimaal%
 \begin{table}
 \centering
-\begin{tabular}{llrrrr}
+\begin{tabular}{llrrr}
 \toprule
-features & class & Precision &  Recall & $F1$ &  AUC_{PR} \\
+features & class & Precision &  Recall & $F1$\\
 \midrule
-\multirow{2}{*}{embeddings}                            & inanimate     &       0.98 &    0.99 &    0.98 & \multirow{2}{*}{0.93} \\
-                                                       & animate     &       0.93 &    0.89 &    0.91 &  \\
-\multirow{2}{*}{word}                                  & inanimate     &       0.96 &    0.99 &    0.98 & \multirow{2}{*}{0.93} \\
-                                                       & animate     &       0.94 &    0.78 &    0.85 &  \\
-\multirow{2}{*}{word + embeddings}                     & inanimate     &       0.98 &    0.99 &    0.98 & \multirow{2}{*}{0.96} \\
-                                                       & animate     &       0.94 &    0.90 &    0.91 &  \\
-\multirow{2}{*}{word + PoS}                            & inanimate     &       0.97 &    0.99 &    0.98 & \multirow{2}{*}{0.95} \\
-                                                       & animate     &       0.94 &    0.86 &    0.89 &  \\
-\rowcolor{Gray}                                        & inanimate     &       0.98 &    0.99 &    0.99 &  \\
-\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings}   & animate     &       0.94 &    0.91 &    0.93 & \multirow{-2}{*}{0.97} \\
-\multirow{2}{*}{word + PoS + lemma}                    & inanimate    &       0.97 &    0.99 &    0.98 & \multirow{2}{*}{0.96} \\
-                                                       & animate     &       0.94 &    0.86 &    0.90 &  \\
-\multirow{2}{*}{word + PoS + lemma + embeddings}       & inanimate     &       0.98 &    0.99 &    0.99 & \multirow{2}{*}{0.97} \\
-                                                       & animate     &       0.94 &    0.91 &    0.93 &  \\
-\multirow{2}{*}{word + PoS + lemma + dep}              & inanimate     &       0.97 &    0.99 &    0.98 & \multirow{2}{*}{0.96} \\
-                                                       & animate     &       0.94 &    0.86 &    0.90 &  \\
-\multirow{2}{*}{word + PoS + lemma + dep + embeddings} & inanimate     &       0.98 &    0.99 &    0.99 & \multirow{2}{*}{0.97} \\
-                                                       & animate     &       0.94 &    0.92 &    0.93 &  \\
+\multirow{2}{*}{embeddings}                            & inanimate     &       0.98 &    0.99 &    0.98 \\
+                                                       & animate     &       0.93 &    0.89 &    0.91   \\
+\multirow{2}{*}{word}                                  & inanimate     &       0.96 &    0.99 &    0.98 \\
+                                                       & animate     &       0.94 &    0.78 &    0.85   \\
+\multirow{2}{*}{word + embeddings}                     & inanimate     &       0.98 &    0.99 &    0.98 \\
+                                                       & animate     &       0.94 &    0.90 &    0.91   \\
+\multirow{2}{*}{word + PoS}                            & inanimate     &       0.97 &    0.99 &    0.98 \\
+                                                       & animate     &       0.94 &    0.86 &    0.89   \\
+\rowcolor{Gray}                                        & inanimate     &       0.98 &    0.99 &    0.99 \\
+\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings}   & animate     &       0.94 &    0.91 &    0.93 \\
+\multirow{2}{*}{word + PoS + lemma}                    & inanimate    &       0.97 &    0.99 &    0.98  \\
+                                                       & animate     &       0.94 &    0.86 &    0.90   \\
+\multirow{2}{*}{word + PoS + lemma + embeddings}       & inanimate     &       0.98 &    0.99 &    0.99 \\
+                                                       & animate     &       0.94 &    0.91 &    0.93   \\
+\multirow{2}{*}{word + PoS + lemma + dep}              & inanimate     &       0.97 &    0.99 &    0.98 \\
+                                                       & animate     &       0.94 &    0.86 &    0.90   \\
+\multirow{2}{*}{word + PoS + lemma + dep + embeddings} & inanimate     &       0.98 &    0.99 &    0.99 \\
+                                                       & animate     &       0.94 &    0.92 &    0.93   \\
 \bottomrule
 \end{tabular}
-\caption{Precision, Recall, $F1$-score and Area Under the Precision Recall Curve (AUC_{PR}) scores for animate and inanimate classes per feature setting for all words.}
+\caption{Precision, Recall and $F1$-score for animate and inanimate classes per feature setting for all words.}
 \label{tab:results-all}
@@
 precision and high recall show us that this model has acquired
-knowledge about which words \textit{typically} group with animate
+knowledge about which words \emph{typically} group with animate
 words and which with inanimate words. However, the models that combines
@@
 instances and 0.95 for inanimate instances.
+%moet in deze tabel ook AUC pr?%
 
@@
 \centering
-\begin{tabular}{llrrrr}
+\begin{tabular}{llrrr}
 \toprule
-features & class &  Precision &  Recall & $F1$ &  AUC_{PR} \\
+features & class &  Precision &  Recall & $F1$ \\
 \midrule
-\multirow{2}{*}{embeddings}                            & inanimate     &       0.90 &    0.96 &    0.92 & \multirow{2}{*}{0.91} \\
-                                                       & animate       &       0.93 &    0.85 &    0.89 &  \\
-\multirow{2}{*}{word}                                  & inanimate     &       0.78 &    0.98 &    0.87 & \multirow{2}{*}{0.91} \\
-                                                       & animate       &       0.96 &    0.60 &    0.74 &  \\
-\multirow{2}{*}{word + embeddings}                     & inanimate     &       0.90 &    0.97 &    0.93 & \multirow{2}{*}{0.96} \\
-                                                       & animate       &       0.95 &    0.85 &    0.90 &  \\
-\multirow{2}{*}{word + PoS}                            & inanimate     &       0.86 &    0.96 &    0.90 & \multirow{2}{*}{0.94} \\
-                                                       & animate       &       0.93 &    0.78 &    0.84 &  \\
-\rowcolor{Gray}                                        & inanimate     &       0.93 &    0.96 &    0.95 &  \\
-\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings} & animate       &       0.95 &    0.90 &    0.92 &\multirow{-2}{*}{0.97}  \\
-\multirow{2}{*}{word + PoS + lemma}                    & inanimate     &       0.87 &    0.96 &    0.91 & \multirow{2}{*}{0.95} \\
-                                                       & animate       &       0.94 &    0.80 &    0.86 &  \\
-\multirow{2}{*}{word + PoS + lemma + embeddings}       & inanimate     &       0.93 &    0.96 &    0.94 & \multirow{2}{*}{0.97} \\
-                                                       & animate       &       0.95 &    0.89 &    0.92 &  \\
-\multirow{2}{*}{word + PoS + lemma + dep}              & inanimate     &       0.87 &    0.96 &    0.91 & \multirow{2}{*}{0.95} \\
-                                                       & animate       &       0.93 &    0.80 &    0.86 &  \\
-\multirow{2}{*}{word + PoS + lemma + dep + embeddings} & inanimate     &       0.93 &    0.96 &    0.95 & \multirow{2}{*}{0.97} \\
-                                                       & animate       &       0.95 &    0.90 &    0.92 &  \\
+\multirow{2}{*}{embeddings}                            & inanimate     &       0.90 &    0.96 &    0.92  \\
+                                                       & animate       &       0.93 &    0.85 &    0.89  \\
+\multirow{2}{*}{word}                                  & inanimate     &       0.78 &    0.98 &    0.87  \\
+                                                       & animate       &       0.96 &    0.60 &    0.74  \\
+\multirow{2}{*}{word + embeddings}                     & inanimate     &       0.90 &    0.97 &    0.93  \\
+                                                       & animate       &       0.95 &    0.85 &    0.90  \\
+\multirow{2}{*}{word + PoS}                            & inanimate     &       0.86 &    0.96 &    0.90  \\
+                                                       & animate       &       0.93 &    0.78 &    0.84  \\
+\rowcolor{Gray}                                        & inanimate     &       0.93 &    0.96 &    0.95  \\
+\rowcolor{Gray}\multirow{-2}{*}{word + PoS + embeddings} & animate       &       0.95 &    0.90 &    0.92\\
+\multirow{2}{*}{word + PoS + lemma}                    & inanimate     &       0.87 &    0.96 &    0.91  \\
+                                                       & animate       &       0.94 &    0.80 &    0.86  \\
+\multirow{2}{*}{word + PoS + lemma + embeddings}       & inanimate     &       0.93 &    0.96 &    0.94  \\
+                                                       & animate       &       0.95 &    0.89 &    0.92  \\
+\multirow{2}{*}{word + PoS + lemma + dep}              & inanimate     &       0.87 &    0.96 &    0.91  \\
+                                                       & animate       &       0.93 &    0.80 &    0.86  \\
+\multirow{2}{*}{word + PoS + lemma + dep + embeddings} & inanimate     &       0.93 &    0.96 &    0.95  \\
+                                                       & animate       &       0.95 &    0.90 &    0.92  \\
 \bottomrule
 \end{tabular}
-\caption{Precision, Recall, $F1$ score and Area Under the Precision Recall Curve (AUC_{PR}) scores for animate and inanimate classes
+\caption{Precision, Recall, and $F1$ score for animate and inanimate classes
   per feature settings for all words tagged as noun.}
@@
 
-\section{Animacy Detection in the Wild}
+\section{Listing Characters in Stories}\label{sec:listing}
 
 In a classical evaluation setup -- as with our approach -- general
-practice is to divide an annotated data set into a set for training a
-computational system of which the performance is evaluated on a
-held-out test set. Even if we include a separate development set, this
-setup always runs the risk over overfitting the data. Our annotated
-corpus contains a reasonably diverse set of stories, yet it is fairly
-small and stylistically homogeneous. Even though we performed a
+practice is to train a computational system on a training data set of
+which the performance is evaluated on a held-out test set. Even if we
+include a separate development set, this setup could run the risk
+of overfitting the data. Our annotated corpus contains a reasonably
+diverse set of stories in terms of genre, yet it is fairly small and
+stylistically rather homogeneous. Even though we performed a
 cross-validation experiment, the chances of overfitting are
 considerably high. In this section we investigate how well our animacy
-detection model performs on a collection of unannotated texts.
+detection model performs on a large collection of unannotated texts.
+%ik zou misschien het stuk over evaluatie hier achteraan zetten: het is meer methodologie%
+\subsection{Data}
+For our evaluation we again make use of a collection of folk tales, consisting of a subsection of the Dutch Folktale Database\footnote{\url{http://www.verhalenbank.nl}}.
+The complete collection consists of about 42,000 folktales~\cite{meder:10}, and contains stories from various genres (e.g.~fairytales, legends, urban legends, jokes, personal narratives) in standard Dutch and Frisian, as well as a number of dialectal variants.
+Every entry in the database contains
+meta-data about the story, including language, collector, place and
+date of narration, keywords, names, and subgenre. For our paper we make use of a
+subcollection comprising 16294 stories written in standard Dutch. The distribution of genres in the subcollection is the following: urban legends ($n=2795$), legends
+($n=299$), jokes ($n=3986$), personal narratives ($n=693$), riddles
+($n=1626$), sagas ($n=6045$) and fairy tales ($n=832$). We evaluate a
+random sample of this subcollection ($n=212$) in which this genre
+distribution is taken into account.
 
-The Dutch Folktale Database\footnote{\url{http://www.verhalenbank.nl}}
-is a collection of about 42,000 folktales~\cite{meder:10}. The
-collection contains stories from various genres (e.g.~fairytales,
-legends, urban legends, jokes, personal narratives) in a number of
-variants of Dutch and in Frisian. Every entry in the database contains
-meta-data about the story, including language, collector, place and
-date of narration, keywords, names, and subgenre. We make use of a
-subcollection of the Dutch Folktale Database which comprises 16253
-stories written in standard Dutch.
+\subsection{Evaluation}
+Our definition of animacy allows us to utilize our animacy detection
+system to extract all characters from a story in a similar vain as
+\cite{karsdorp:12}. In a real application, the system labels each noun
+and name in a text for animacy. After removing duplicate words, this
+produces a set of words that comprises the cast of a story. Without
+golden standard annotations, however, we can only evaluate these
+character sets for precision and not for recall. An alternative
+approach is to produce a ranking of all words in a story where the
+goal is to allocate the highest ranks to animate entities. This allows
+us to evaluate the rankings using \emph{Average Precision} which
+computes the average over precision scores at increasing points of
+recall. Naturally, with this evaluation method, we still need to
+manually evaluate the rankings. By using a rank cutoff and evaluating
+a sample of all automatically annotated stories, we reduce the costly
+manual labor to a minimum. We order all nouns and names in a story
+using the output of the probabilistic decision function of the Maximum
+Entropy classifier. After removing duplicate words, this produces a
+final ranking. The rankings are evaluated using \emph{Average
+  Precision} with a rank cutoff at 50.
+
+\subsection{Results}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{precision-recall-curve}
+\caption{Precision-Recall Curves and Average Precision scores per
+genre.}
+\label{fig:prec-recall-curve}
+\end{figure}
+
+We present the results in Figure \ref{fig:prec-recall-curve} in which
+we show the Precision-Recall curve as well as the Average Precision
+score for each genre. The system performs well, especially on fairy
+tales (AP=0.97) and jokes (AP=0.94). The lowest performance is
+measured on riddles (AP=0.85). This lower score is mainly the result
+of the system's inability to position the word \emph{blondje}
+(`dumb blonde' with a pejorative connotation, which is prevalent in the so-called `dumb blonde jokes') high up the ranking.
+
+%misschien nog verschillende scores benoemen?%
+%misschien ook nog iets zeggen over de karakter ranking? Of was dat juist het onderdeel dat we er niet meer in gingen betrekken?%
+\begin{sidewaysfigure}
+\includegraphics[width=\linewidth]{map}
+\caption{Visualization of characters in the Dutch Folktale Database based on their embeddings using t-SNE.}
+\label{fig:map}
+\end{sidewaysfigure}
 

### 2015-01-20 experiment.py

@@ -211,8 +211,8 @@ class FeatureStacker(BaseEstimator):
                     out['%s__%s' % (name, key)] = value
             return out
 
-FIELDNAMES = ['word', 'root', 'lcat', 'pos','rel', 'sense', 'frame',
-              'special','noun_det', 'noun_countable', 'noun_number',
+FIELDNAMES = ['word', 'root', 'lcat', 'pos', 'rel', 'sense', 'frame',
+              'special', 'noun_det', 'noun_countable', 'noun_number',
               'verb_auxiliary', 'verb_tense', 'verb_complements', 'cluster',
               'animate', 'reference']
 

### 2015-01-20 rnn_animacy.py

@@ -29,7 +29,7 @@ def minibatch(l, bs):
     '''
     out  = [l[:i] for i in xrange(1, min(bs,len(l)+1) )]
     out += [l[i-bs:i] for i in xrange(bs,len(l)+1) ]
-    assert len(l) == len(out)
+    assert len(l) == len(out) # 
     return out
 
 def contextwin(l, win):
@@ -179,7 +179,7 @@ if __name__ == '__main__':
     embeddings = pretrained_embeddings.syn0
     de = embeddings.shape[0]
     vocab = pretrained_embeddings.vocab
-    learning_rate = 0.0627142536696559
+    learning_rate = float(sys.argv[7]) if len(sys.argv) == 8 else 0.0627142536696559
     verbose = 1 if sys.argv[5] == 'verbose' else 0
     bs = 9
     n_hidden = int(sys.argv[6])
